### 2023-08-20 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class

\PassOptionsToPackage{table}{xcolor}

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage{todonotes}

% Change the section names
\renewcommand{\sectionautorefname}{\S\!}%
\renewcommand{\subsectionautorefname}{\S\!\!}%
\renewcommand{\subsubsectionautorefname}{\S\!}%
\renewcommand{\equationautorefname}{Eq.}

\newcommand{\package}[1]{\textsc{#1}}
\newcommand{\stream}[1]{#1}
\newcommand{\dataarchive}[1]{\texttt{#1}}
\newcommand{\Gaia}{\dataarchive{Gaia}}
\newcommand{\gaia}{\dataarchive{gaia}}

% Math fonts
\newcommand{\mrm}[1]{\mathrm{#1}}
\newcommand{\mbs}[1]{\boldsymbol{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mfk}[1]{\mathfrak{#1}}
\newcommand{\mcal}[1]{\mathcal{#1}}

% Math symbols
\newcommand{\Exp}[1]{e^{#1}}
\newcommand{\Erf}[1]{\mathrm{erf}\left(#1\right)}
\NewDocumentCommand \dif {o m} {
	\IfNoValueTF{#1}
		{ \mathrm{d}{#2} }
		{ \mathrm{d}^{#1}{\!#2} }
}

% Statistics
\newcommand{\pdf}{\mcal{P}}
\newcommand{\cdf}{\mfk{P}}
\newcommand{\likelihood}{\mcal{L}}
\newcommand{\prior}{\mcal{\pi}}
\newcommand{\posterior}{p}

% Labels
\newcommand{\nth}[1]{{#1}_{\mrm{n}}}  % measurement
\newcommand{\fth}[1]{{#1}_{\mrm{f}}}  % feature
\newcommand{\qth}[1]{{#1}_{\mrm{q}}}  % model index
\newcommand{\unit}[1]{[\text{#1}]}

\newcommand{\smallcomponent}[2]{#2^{\scriptscriptstyle (#1)}}
\newcommand{\cmp}[2]{\smallcomponent{#1}{#2}}
\newcommand{\Scmp}[1]{\cmp{S}{#1}}
\newcommand{\Bcmp}[1]{\cmp{B}{#1}}
\newcommand{\astroM}[1]{{#1}_w}
\newcommand{\photoM}[1]{\smallcomponent{m}{#1}}
\newcommand{\chemoM}[1]{\smallcomponent{z}{#1}}

\newcommand{\Spdf}{\Scmp{\pdf}}
\newcommand{\Bpdf}{\Bcmp{\pdf}}

\newcommand{\parallax}{\varpi}
\newcommand{\sigobs}{{\sigma_*}}

\newcommand{\TODO}[1]{{\textcolor{red}{#1}}}
\newcommand{\JN}[1]{\TODO{#1}}

% Begin!
\begin{document}

% Title
\title{Characterizing Stellar Streams by ML}

% Author list
\author{Nathaniel Starkman \& Jacob Nibauer}
\author{Jo Bovy}
\author{Jeremy Webb}
\author{Ana Bonaca}

% Abstract with filler text
\begin{abstract}
    Stellar stream membership likelihoods with ML. It works!
\end{abstract}

\section{Introduction} \label{sec:intro}
    Stellar streams are the disrupted remnants of globular clusters and satellite galaxies, and their census has grown extensively in the Milky Way thanks to all-sky astrometric missions like {\it Gaia} and low-surface brightness photometric surveys (\JN{cite}). The formation of streams can be traced back to a progenitor cluster (i.e., a satellite galaxy or globular cluster), whose stars are gradually lost to the tidal field of the host galaxy and extend along a series of similar orbits producing tidal tails, which can sometimes span many tens of degrees across the sky. This orbital proximity of stars belonging to a stream makes them sensitive tracers of the mass distribution of galaxies, sourced by both the baryonic and dark matter components.

    The average properties of a stream (e.g., the mean position and velocity track of the tidal tails) can be used to reconstruct the global dark matter distribution of the stellar halo (e.g., \JN{cite}), while local variations in the stream track and surface density encode information about the stream's encounters with baryonic structures such as (e.g.) the stellar bar (\JN{cite}), and non-luminous structures like dark matter subhalos (\JN{cite}). Importantly, these galactic components span orders of magnitude in mass, highlighting the importance of characterizing a population of stellar streams both in morphology and kinematics.

    While several methods have been developed to model stream properties as a function of global halo properties (e.g., flattening and orientation;\JN{cite}) and local substructures (e.g., the bar's pattern speed and the mass of subhalo perturbers; \JN{cite}), these methods and insights often rely on the existence of a homogenized stream-tracer population that characterizes the stream across each of the observable phase-space dimensions. Additionally, because even the fine-grain properties of streams are influenced by the detailed mass distribution of galaxies, it is important to propagate errors both in measurement uncertainty and membership uncertainty (i.e., which stars belong to the stream versus the background) when attempting to model the galaxy. Otherwise, a stream that has been poorly characterized could bias constraints on (e.g.) the total mass of the galaxy, the dark matter halo shape, and the mass of subhalo perturbers. In order to generate more precise constraints on the structure of the galaxy, it is therefore imperative that a statistically sound and homogenized catalog of stellar streams is produced so that theoretical models for the gravitational potential of the galaxy across a large range of spatial scales can be constrained.

    While streams are well described in (e.g.) action angle coordinates or $N-$body simulations, these methods relying on prior assumptions about the galaxy and its underlying potential. Therefore, it is important to devise data-driven methods to characterize stellar streams without appealing to strong prior assumptions about the galaxy's structure.

    There are a few previous works which have modeled stellar streams in this context. \citet{2022MNRAS.514.1757P} developed a data-driven spline-based model for modeling the photometric qualities of previously identified streams. Their method works by fitting the stream's stellar surface density with a series of splines, capturing both surface brightness variations along the stream and changes in the position of the stream track and width. Their method also utilize a simultaneous ischorone fit to the stream, from which a distance modulus can be derived by assuming that the entire stream originates from the same stellar population. While this method is very successful at characterizing the photometric properties of a stream, it does not consider the kinematic dimensions which have become increasing well-measured with successive data releases from {\it Gaia}, as well as targeted observations. From a population of RR Lyrae stars, \citet{2019AJ....158..223P} utilized a mixture model to fit the globular cluster stream Pal-5 in the space of proper motions and heliocentric distances. The mixture modeling approach enables a simultaneous fit to both the stream and background density, enabling better separation of a coherent stream from an otherwise noisy background distribution. However, the track of the stream in on-sky coordinates is not fit. The method identified 27 RR Lyrae stars consistent with being members of the stream. While useful for determining the average properties of the stream (e.g., its mean distance track), a limited population of RR Lyrae stars are not sufficient to capture the fine-grained properties of streams expected to inform constraints on (e.g.) a population of perturbers in the galaxy. \JN{are there other methods?}

    Other methods for extracting stellar streams from the data include matched-filter approaches, where over-densities in color-magnitude space are inspected as possible clusters undergoing tidal stripping \JN{cite the many key papers using this method}. While the matched-filter technique has been extremely successful at stream discovery, it is not equipped to characterize streams at the level of detailed necessary to inform precise constraints on the potential of the galaxy and its substructure.

    In order to address the need for a flexible model of stellar streams both in photometric and kinematic spaces, in this work we develop a new method for characterizing streams and quantifying the membership probability of possible stream stars. Our approach provides several key advantage over previous stream modeling efforts, specifically in its ability to jointly model the kinematic and photometric properties of streams simultaneously. Furthermore, our method is well-suited to fitting streams with only partial or incomplete phase-space or photometric observations, enabling a full characterization of streams insofar as the available data allows.

    \begin{quotation}

        A stellar stream is the association of stars tidally stripped from a common
        progenitor system that orbits in an underlying galactic gravitational field
        \citep{Johnston1998, HelmiWhite1999}. Stream progenitors are gravitationally
        bound systems, such as satellite dwarf galaxies or star clusters
        \citep{Odenkirchen2001, Majewski2003}. In the absence of external interactions
        these systems are stable, though still capable of mass loss driven by internal
        processes \citep{Hills1975, HeggieHut2003}, e.g. relaxation and binary
        interactions. Considering external interactions, these progenitor systems may
        be disrupted: completely, as in the merger events that formed the Galactic
        stellar halo \citep{Lynden-Bell1967, Searle1978}; or partially, such as tidal
        disruption forming a stellar stream.

        In the approximation of a smooth galactic gravitational field, the
        progenitor's extent is limited to the tidal radius. Through processes internal
        and external, progenitor stars can achieve sufficient energy to bring them
        beyond the tidal boundary, thus ``escaping'' into the galactic field. Escape
        primarily occurs at both an inner and outer saddle point of the combined
        potential -- e.g. the L2 and L3 Lagrange points for a spherical galactic field
        \citep{Ross1997, Fukushige2000, BinneyTremaine2008}. Generally, the initial
        phase space position of the escaped star is very similar to that of the
        progenitor: the configuration position is the tidal boundary and the star has
        a low peculiar velocity. Since both the escaped star and its progenitor orbit
        in the same underlying galactic potential and have similar phase space
        positions their orbits are similar \citep{BinneyTremaine2008}. Given
        sufficient time the phase space separation will increase and an escaped star
        will lead or trail the position of the progenitor, depending on the location
        from which it escaped. Therefore, a collection of tidally stripped stars will
        form extended structures, aka tails or streams, that are few to thousands of
        parsecs in length \citep{Johnston1998, HelmiWhite1999, Bovy2014}.

        To be a coherent, observable structure the peculiar velocity dispersion of
        escaped stars relative to the progenitor must be relatively low, but does vary
        from system to system. Stars in kinematically hot streams (e.g. from
        progenitors with larger kinematic dispersions) have more dissimilar orbits
        from each other and the progenitor than do stars in cold streams, with smaller
        dispersions \citep{Johnston1998, Hendel2015}. The orbit of a star depends on
        its phase space position and the gravitational potential; therefore measuring
        a star's position and velocity and observing its orbit reveals the local
        potential, which is normally dominated by the Galaxy (primarily the dark
        matter halo at large distances) \citep{Gibbons2014}. However, an orbit is
        impractical to observe on human timescales. Cold streams offer a workaround.
        Like measuring one star at many points in one orbit over an extended period,
        streams are many stars measured at different points along approximately one
        orbit at the same time. Therefore stellar streams are one of the most
        promising means to study the Galactic gravitational potential: the stellar
        components as well as the dark matter \citep{Johnston2016}.

        On the observational front there has been a large concerted effort in the
        field to find new stellar streams \citep[for an extensive list of known
        Galactic streams see][]{Mateu2022}, extend the detected extents of known
        streams \citep[e.g. for \stream{Palomar 5} alone:][]{Rockosi2002, Ibata2017,
        Grillmair2006, Carlberg2012, StarkmanEtAl2019}, and increase the general
        purity of stream catalogs.

    \end{quotation}

    \todo[inline]{--- Rephrase}

    \paragraph{one orbit and a single population}

        Stellar streams consist of stars that are tidally disrupted from a progenitor system.
        To a first approximation, the stars occupy a single orbit, that of their progenitor.
        Additionally, as stars formed in the progenitor they retain its photometric and chemical characteristics: which for globular clusters means following an isochrone.
        In this simplistic picture, astrometrically characterizing a stream requires only finding it's orbit. Member stars fall directly on the orbit and the isochrone, making membership likelihood determination trivial.

        % Observational errors complicate this picture. Instead of a certain membership calculation, we must build a probabilistic model, assigning a likelihood that an observed star falls along the stream's orbit.

    \paragraph{similar orbits and a somewhat noisy population}

        Occupying a single orbit implies that streams remain a point in action space, and having a common origin would thus not become extended structure. This is not true.
        Differences in stream stars' orbits from the progenitor's arise from various factors including internal dynamics of the progenitor and the host Galactic potential.

        An intrinsic width, in configuration and kinematic space.

        Centered on a track, not an orbit.

        This defines an astrometric phase-space hyper-tube \cite{2018MNRAS.477.4063M}, with a distribution radial to the track.

        Instead of characterizing the stream by its orbit, we characterize it by it's track and the distribution around that track.
        Computing membership likelihoods now requires fitting the track and distribution. This cannot be
        done without then also fitting the field of stars in which the stream lives, since the stream 
        and field cannot be clearly separated.

    \paragraph{gaps orbits and a somewhat noisy population}

        This is still a simplistic picture of streams.

        Perturbations to the Galactic potential, in the form of time evolution and non-axisymmetry, as well as substructure can disrupt the stellar stream. 
        Even most perturbations don't ruin the idea of the phase-space hypertube. They either drop the mixture weight (rephrase, not yet introduced) to 0 or change the width of the hypertube.

    \paragraph{Review of detection methods}

        Identifying streams of globular of clusters is a challenging task. Streams are very low signal-to-noise structures spread over comparatively large regions of the sky. They live in a vast background of field stars.

        Historically streams were detected by matched-filter techniques
        % Add citations
        You took an isochrone and shifted it by a distance modulus until a stream stood out in longitude and latitude as an extended overdensity structure.

        With the advent of Gaia, we gained access to detailed astrometric data. For most streams this

        poses a challenging task due to their intricate and diffuse nature
        from large-scale surveys is challenging due to the complex and overlapping distributions of stars in different regions of the sky; disentangling members from the vast background of field stars

        \begin{enumerate}
            \item Streamfinder \citep{2018MNRAS.477.4063M}. This method requires the potential
            \item VIA MACHINAE \citep{2022MNRAS.509.5992S}
            \item Kiyan and Adrian
            \item Matched Filter
            \item Gaia stuff: Adrian and Ana and Sara et al
            \item Uniform Modeling of 12 Stellar stream \citep{PatrickEtAl2022}. Only on phi1, phi2, and CMD. Their CMD method is more sophisticated. Worth discussing.
        \end{enumerate}

    Review of available data:
    \begin{enumerate}
        \item Gaia. Streams in Gaia have on-sky positions and proper motions, and parallaxes, though for most streams the parallax errors are prohibitively large. Likewise, most streams in Gaia do not have good radial velocity measurements. The photometry is generally poor, generally necessitating cross-matching with other datasets.
        \item S5. This is a much smaller dataset. It has deeper photometry and radial velocity measurements.
    \end{enumerate}

    \paragraph{Paper organization}

        The paper is organized as follows.
        In \autoref{sec:method} we introduce our method for characterizing stream tracks and computing stellar membership.
        We discuss the data in \autoref{sub:framing_the_data} and build a probabilistic framework to model the stream and background in \autoref{sub:likelihood_setup}-\autoref{sub:priors}.
        In \autoref{sec:results} we apply our method,
        % TODO: rephrase the mock data
        first to simulated data (\autoref{sub:mock_data}) to demonstrate the model with a known ground truth dataset,
        then to the stellar streams \stream{GD-1} (\autoref{sec:results_gd1}) and \stream{Pal\,5} (\autoref{sub:pal5}).
        We compare our approach to characterizing stellar streams to other methods in \autoref{sec:comparison}.
        We conclude in \autoref{sec:conclusions}, discussing the results of this work and future directions.

% section introduction (end)

\section{Method} \label{sec:method}

    We now introduce our method for characterizing stellar stream tracks and computing stream membership likelihoods from N-D phase-space measurements.
    % TODO: rephrase following sentence
    We develop a flexible Bayesian model framework, that allows arbitrary numbers of phase-space dimensions to be incorporated in the model and permits missing phase-space observations. 
    Below, we offer a concise overview of our approach, followed by a more in-depth technical explanation in subsequent sections.
    We start, in \autoref{sub:framing_the_data}, by transforming the astrometric data into
    a stream-oriented on-sky reference frame ($\phi_1, \phi_2, \parallax, \mu_{\phi_1}^*, \mu_{\phi_2}, v_r)$, for each stream. In this frame, 
    the stream track lies along $\phi_1$, and the stream density may be described by distributions in the other parameters as injective functions of $\phi_1$.
    % TODO: ⬆ I think a few more sentences explaining why this is important, feature variations, etc?
    % However, previously we discussed how stream will vary along the track, with changes in the width and possible gaps.
    % Also observational effects, e.g the distance gradient changes the on-sky width (should this be mentioned earlier?)
    Likewise, the background of the Galactic field may be described with $\phi_1$-conditioned distributions.
    % TODO: ⬆ I think this should be expanded.
    Non-astrometric coordinates, e.g. photometry and metallicity, are also modelled along the stream and background as functions of $\phi_1$.

    The distributions for the stream and background are collected in a single mixture model that describes the entire field.
    We take the parameters of the mixture model -- the mixing coefficients and component distribution parameters -- to be general (continuous) functions of the conditioning variable, $\phi_1$. This is done using feed-forward neural network that takes the conditioning variable as its input and output the parameter value. Mixture models of conditional probability distributions using neural networks are called Mixture Density Networks (MDNs). Since the model coefficients may evolve with $\phi_1$, MNDs are well suited to characterize streams (and the background) with all their
    variations in position, width, and linear density.

    % subsection framing_the_data (end)

    \subsection{Framing the Data} \label{sub:framing_the_data}

        As a stream progenitor orbits the Galactic center of mass (GCOM), its orbital path
        describes an ellipse in the orbital plane. The associated arms of the stream generally
        align closely with its orbital plane.
        From a Galactocentric reference frame, an orbital plane assumes the form of a great
        circle in the observer's celestial ``sky," thereby allowing the potential realignment of
        longitudinal ($\ell$) and latitudinal ($b$) coordinates to a coordinate system aligned
        with the stream ($\cmp{GCOM}{\phi_1}, \phi_2^{(GCOM)}$). Here, the progenitor is
        established at the origin, with the stream progressing along
        $\cmp{GCOM}{\phi_1} (\phi_2^{(GCOM)} \approx 0)$. $\cmp{GCOM}{\phi_1}$ is clearly a good
        affine parameter along which to construct models and analyze the stream's characteristics.

        However, data observations are not taken from the GCOM, but from  $\sim 8$ kpc distant
        \citep{GRAVITY2018, Leung2022} in a local frame, typically from vantage points such as
        Earth or its solar Lagrange points, and subsequently represented in the International
        Celestial Reference System (ICRS) \citep{ICRS1997}.
        Restricted as we are to a very local volume, there is no means to observe systems from
        the more ``intrinsic" reference frames. In our observational frames there is no inherent
        assurance of effecting a rotation that would similarly align the stream within observer
        coordinates such that $\phi_2(\phi_1) \approx 0$.
        Thankfully, most streams are adequately distant from both the Galactic center and the
        observer, or exhibit orientations such that a rotation a) can be established or
        b) holds valid. For most of the best-studied globular cluster streams -- \stream{GD-1},
        \stream{Palomar 5}, \stream{Jhelum}, etc. -- this holds true. However for some streams
        of dwarf galaxies, like the Magellanic stream which wraps multiple times around the
        Milky Way \citep{Wannier1972}, no such transformation is possible. More generally,
        a stream's ``sky"-projected path may be kinked, e.g. by subhalo interactions, such
        that it is a many-to-one function in $\phi_1$.

        In this work we treat with the vast majority of streams, which have a ``sky" transformation such that the stream is an injective function in $\phi_2(\phi_1)$.
        Since we are examining known streams this transformation is known \textit{a priori}, or
        may be constructed during pre-processing \citep[see][\S2.1]{StarkmanEtAl2023}.
        We note that, with sufficient prior knowledge, many-to-one functions may be broken into injective segments and their connection constrained by a prior. Thus the methods
        developed subsequently may extented and applied even to wrapped and kinked streams. However, 
        we leave extending the model framework to a future work.

    \subsection{Likelihood Setup}\label{sub:likelihood_setup}

        \begin{figure}
            \centering
            \includegraphics[scale=0.5]{static/model/DAG_StreamDensity.pdf}
            \caption{Probabilistic Graphical Model of the Mixture Density Network (MDN)}
            \todo[inline]{todo}
            \begin{enumerate}
                \item remake this example this with \url{https://github.com/daft-dev/daft}.
            \end{enumerate}
            \label{fig:DAG}
        \end{figure}

        Mixture models (MMs) are a statistical method to represent a population as a set of a sub-populations specified by probability distributions. These distributions are characterized by their parameters and the mixture by a corresponding set of mixture coefficients. Scalar mixture models may be extended to model completely general conditional distributions by making the distribution parameters and mixture coefficients into functions of a conditional input \citep{McLachlan1989MixtureM}.
        Mixture Density Networks (MDNs) are a machine learning method to implement these general MM, taking the parameters and coefficients to be the outputs of feed-forward neural networks and as such, general continuous functions \citep{Bishop1994}. Functional parameters allow the model to capture variation over those parameters that is not possible with scalar-valued models. Moreover, by using neural networks the inferred functional forms are ``model-free" and driven entirely by the data.

        For the MDN we take $x_{\cdot, 0}$, which is the first feature column of the data $\mbf{X}$, as the `independent' coordinate over which the MDN parameters are conditioned. In the context of streams $x_{\cdot, 0} \equiv \phi_1$ is the longitude coordinate in a reference frame rotated such that the stream is aligned along $\phi_1$. The parameters of the stream, the background field, and the mixing coefficients of the two are now general data-driven continuous functions along $\phi_1$.

        The PDF of a general MDN is
        \begin{equation} \label{eq:general_mixture_network}
            \!\!\! \pdf(\nth{\mbs{x}} | \mbs{\theta}(\phi_1))
            \!=\! \sum_{\mrm{q} \in \mbs{I_Q}} \! \qth{f}(\phi_1) \cmp{q}{\pdf}(\nth{\mbs{x}}|\qth{\mbs{\theta}}(\phi_1)),
            % \prior(\nth{\mbs{x}}, \phi_1)
        \end{equation}
        where $\cmp{q}{\pdf}$ is the PDF of the $q$-th model and $f_q$ its mixing coefficient for all $q$ models in the set of models indexed by $\mbs{I_Q}$. $\mbs{\theta}$ is the $\phi_1$-conditional output of the feed-forward neural network, and $\nth{\mbs{x}}$ the data on the $n$-th star in $\mbs{X}$, not including $\phi_1$. For convenience we drop the implied subscript $n$ in $\phi_1$, the explicit notation of $\mbs{\theta}$'s dependence on $\phi_1$, and the index set $\mbs{I}_X$ in summations over its elements $x$. A full detailing of notation and symbols is included in \autoref{app:notation}.

        The mixture coefficients $\qth{f}$ are normalized s.t.
        \begin{equation}
            \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
        \end{equation}
        In practice this is enforced by defining
        \begin{equation} \label{eq:mixture_weight_normalization}
            f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash \{b\}} \qth{f}(\phi_1),
        \end{equation}
        the summation over all models except for some ``background" index $b \in \mbs{I_Q}$.
        We discuss the mixture weights further in \autoref{sub:mixture_weight}.

        In this work we consider primarily astrometric and photometric data, (notated $\mbs{W}$ and $\mbs{M}$, respectively, as detailed in \autoref{app:notation}), but note that the MDN framework may be extended to arbitrary feature dimensions.
        Assuming the conditional independence of the astrometry and photometry,
        we split each model into linearly separable astrometric and photometric models $\cmp{q}{\pdf} = \cmp{q,w}{\pdf} \cmp{q,m}{\pdf}$ .

        The likelihood for a single star then becomes, simplifying from \autoref{eq:general_mixture_network},
        \begin{equation} \label{eq:general_model}
            \!\!\!\!\! \pdf(\nth{\mbs{w}}, \! \nth{\mbs{m}} | \mbs{\theta})
                \!=\! \sum_{q} \! f_q(\phi_1) \cmp{q,w}{\pdf}(\nth{\mbs{w}}|\mbs{\theta}) \cmp{q,m}{\pdf}(\nth{\mbs{m}}|\mbs{\theta})
        \end{equation}

        Assuming the measurement of each star is independent, the total
        log-likelihood is a sum over all stars:
        \begin{equation} \label{eq:general_likelihood}
            \ln\mcal{L}\left(\{\nth{\mbs{w}},\nth{\mbs{m}}\} | \mbs{\theta}\right) = \sum_n \ln \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}).
        \end{equation}

        It will prove convenient to distinguish the models for $q$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:
        \begin{align} \label{eq:stream_and_bkg_prob}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} \Scmp{P}(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) \Bpdf(\nth{\mbs{w}}|\mbs{\theta}) \Bpdf(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}
        where we have dropped the $(w), (m)$ superscript as the model types are evident from their inputs. We applied \autoref{eq:mixture_weight_normalization} to require only a single mixture coefficient.
        Of course, if more components are included in the mixture then \autoref{eq:stream_and_bkg_prob} becomes the more general \autoref{eq:general_model}. However, this an implementation detail as any more complex model, e.g a model including a cold stream, extra-tidal features, the background stellar halo, and field globular clusters may be rearranged such that all stream-related components (cold or extra-tidal) and all other components (stellar halo and globular clusters) are grouped into multi-component \textit{stream} and \textit{background} models which themselves adhere to a \autoref{eq:general_model}-type relation.

        The probability of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\left(S | \nth{\mbs{w}}, \nth{\mbs{m}}, \mbs{\theta} \right) = \frac{f \Spdf(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}

        Hereafter, when we refer to ``membership probability,"
        \autoref{eq:membership_prob} provides the formal definition. We nuance this definition by considering missing feature dimensions in the following subsection.

    \subsection{Missing Phase-Space Observations} \label{sub:missing_data}

        Streams have a diversity of measurement coverage: several have full 6-D astrometrics as well as photometrics and chemical abundances available \citep[e.g.,][]{Koposov2019, 2020A&A...635L...3A, 2022ApJ...928...30L}, while most have partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \gaia {\color{red} [cite]}, for instance, have proper motions but not radial velocities nor reliable parallax measurements. Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many feature dimensions, however only for a select set of streams. Given the diversity of coverage it is important that the likelihood model -- \autoref{eq:membership_prob} -- account for missing data.

        The method of treatment of missing data depends on the cause for that data to be missing.
        Broadly, there are two categories of causes: randomness and systematics.
        Consider a dataset with all features measured but with a random subset of features masked.
        One possible approach is to attempt to impute the missing features, then feed the full-featured data to the likelihood model. Alternatively, the likelihood model might use only the present features, ignoring those missing. If however, features are missing for systematic reasons, then this approach can introduce systematic bias. Consider a star too far away to measure it's parallax distance or radial velocity. This is a systematic reason and this information should be incorporated into any model as a prior. Not doing so, the data imputation approach might predict too-close distances. Likewise, using only present features in the likelihood model misses that the star is likely very distant.
        The effects of mask systematics are mitigated by not including such data,
        moving the contribution of the systematics from a prior to the evidence.
        Our models compute likelihoods under a specific dataset, and \autoref{eq:membership_prob} is a ratio of the probabilities, so the removed systematics are not expected to be impactful.

        We adopt the approach of modeling the data masks as randomly-distributed and not systematic. We do not impute missing data, instead the per-star likelihood is modelled with a distribution with the same dimensionality as the data vector.
        For example, a star with a 6D astrometric measurement will be modelled with a full 6D distribution, while a star missing its radial velocity measurement will be modelled with a dimensionally-reduced (5D) form of the 6D distribution. Using dimensionally reduced distributions ensure that missing phase-space dimensions do not amplify nor suppress the likelihood. 
        Notwithstanding, we include a quality flag on our calculations indicating the number of missing features.
        We leave incorporating priors on the data mask distribution to future work.

    % subsection likelihood_setup (end)

    \vspace{10pt}
    \subsection{Determination of the Mixture Weight} \label{sub:mixture_weight}
        The mixture weight $f_q(\phi_1)$ as used in \autoref{eq:general_model} is characterized as the fraction of stars belonging to the mixture component $q$ over an increment $[\phi_1, \phi_1 + \dif{\phi_1}]$. Thus, the mixture weight encodes information about the linear density of the stream, which has been shown to depend on substructure in the galaxy (CITE-x). The dependence of the mixture weight on the linear density is captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(w_{i,0}\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$ within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and 
        $N_{\rm tot}$ is the total number of stars in the same band (including those in component $q$). The linear  density for component $q$ is then entirely specified by the function $N_q(\phi_1)$.

        The mixture model specified in \S-X only models the weight parameter $f_q$, and not the components of the ratio in \autoref{eq:fraction_param}. However, the linear density $N_q(\phi_1)$ can be readily obtained from $f_q$, provided that $N_{\rm tot}\left(\phi_1\right)$ can be calculated. Importantly, this function does not depend on the parameters of the on-stream or off-stream model. Instead, $N_{\rm tot}(\phi_1)$ is simply the number of stars in a small $\phi_1$ increment. We therefore model $N_{\rm tot}\left(\phi_1\right)$ as a post-processing step using a single normalzing flow in $\phi_1$.

        In particular, for a given field the distribution of $\phi_1$ coordinates is specified by the set $\{\phi_1\}_f$. A normalizng flow is trained to this dataset, and will satisfy $\int d\phi_1 P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of stars in the field considered. We can then obtain the linear stream density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) .
        \end{equation}

    \subsection{Astrometric Model} \label{sub:astrometric_model}

        \subsubsection{On-Stream} \label{ssub:astrometric_model_on_stream}

            The astrometric track is modeled with a neural network, encompassing each phase space dimension, parameterized as a function of $\phi_1$. The neural network output for a single stream component is the mean location of the stream track in each dimension, $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is then
            \begin{equation}
                \cmp{S,w}{\pdf}(\nth{\mbs{w}} | \mbs{\theta}) \! = \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}}\!(\phi_1), {\cmp{w}{\nth{\mbs{\Sigma}}}}^2 \! + \!  \astroM{\mbs{\Sigma}}^2\!\left(\phi_1\right) \!\big),
            \end{equation}
            where $\astroM{\mbs{\theta}}$ is the set of neural network parameters that define $\astroM{\mbs{\mu}}$ and $\astroM{\mbs{\Sigma}}$; and $\cmp{w}{\nth{\mbs{\Sigma}}}$ the data uncertainties.
            In the simplest case, we take $\cmp{w}{\nth{\mbs{\Sigma}}}, \astroM{\mbs{\Sigma}}$ to be diagonal, consisting of both intrinsic dispersion for each dimension of $\mbs{w}$ and the data uncertainties added in quadrature. The data uncertainties $\cmp{w}{\nth{\mbs{\Sigma}}}$ are drawn from $\mbs{\delta}\mbf{W}$, the dataset of astrometric uncertainties (see \autoref{app:notation}).

            In practice, we work with truncated Gaussians, where the truncation is set by the size of the field in each astrometric dimension. Details of the truncated Gaussian are included in \autoref{app:sub:normal_distribution}. Truncating to the field ensures that the probability density is zero wherever there is no data. This has an added benefit of correctly normalizing the Gaussian distribution, which is not compact (has non-zero probability over all space) and would otherwise not integrate to one within the field.

            As a MDN, we implement the Gaussian distribution parameters with a multi-layered feed-forward neural network. The number of layers required depends on the maximum dimensionality of the Gaussian, and whether the neural network employs regularization techniques like dropout. We use alternating layers of linear and tanh units. Details of the network architecture are discussed in \autoref{sub:model_implementation}.

            In the presence of missing phase-space coordinate $f_k$ we reduce the dimensionality of the Gaussian to that of the data-vector. Namely, for $\mbs{I}_{\tilde{F}} \equiv \mbs{I}_{F} \backslash \{f_k\}$
            \begin{align}
                \nth{\tilde{\mbf{w}}} &\mapsto \nth{\tilde{\mbf{w}}} = \{ w_{n,f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\mu}} &\mapsto \astroM{\tilde{\mbs{\mu}}} = \{ \mu_{f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\Sigma}} &\mapsto \astroM{\tilde{\mbs{\Sigma}}} = \{ \Sigma_{i,j} | i,j \in \mbs{I}_{\tilde{F}} \}, & \nonumber
                \intertext{such that,}
                \Spdf(\nth{\mbs{w}} & | \astroM{\mbs{\theta}}(\phi_1))
                    \mapsto \Spdf(\nth{\tilde{\mbs{w}}} | \astroM{\tilde{\mbs{\theta}}} (\phi_1)) = \mcal{N}_{\tilde{F}}.
            \end{align}

        % ssub: on_stream (end)

        \subsubsection{Off-stream} \label{ssub:astrometric_model_off_stream}

            We now outline the off-stream astrometric model. We take a similar approach to the on-stream astrometric model discussed in \autoref{ssub:astrometric_model_on_stream}, though we typically do not characterize the backgrounds with a Gaussian distribution. Instead, we utilize a range of distributions (i.e., exponential, skew-normal) discussed in \autoref{app:distributions}. The parameters of the user-specified background distributions are themselves neural network outputs, such that any $\phi_1$ ``slice" of the background should be characterized by some analytic density, but the evolution of the backgrounds with $\phi_1$ can be complex. Details of the network architecture are discussed in \autoref{sub:model_implementation}.

            Let $\Bcmp{\fth{\mbs{\theta}}}(\phi_1)$ represent the neural network which outputs a vector of parameters characterizing the background distribution at a given $\phi_1$ for astrometric dimension $f$. The number of outputs for the background model is equal to $\mathrm{dim}\left(\fth{\Bcmp{\mbs{\theta}}}\right)$ (where $f=0$ corresponds to $\phi_1$, which we do not model). At a given $\phi_1$, the probability density for the astrometric dimension $f>0$ is $\Bcmp{\fth{\pdf}}\left(\mbs{w}_{n,f} | \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right)$. We treat each background component as independent, allowing us to write the likelihood as a product over the astrometric dimensions:
            \begin{equation} \label{eq:astrometric_model_off_stream_probability}
                \pdf^{(B)}\left(\nth{\mbs{w}} | \mbs{\theta}(\phi_1) \right) = \prod_{f\in \mbs{I}_{\tilde{F}}} \pdf_f^{(B)}\left(\mbs{w}_{n,f} | \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right).
            \end{equation}
            Where $\mbs{I}_{\tilde{F}}$ is the set of indices of the features, omitting missing phase space dimensions.

        % ssub: off_stream (end)

    % sub: astrometric_model (end)

    \subsection{Photometric Model} \label{sub:photometric_model}

        \subsubsection{On-Stream} \label{sub:photometric_model_on_stream}
            In photometric coordinates we model the stream as a single-population isochrone with the possibility of a non-zero distance gradient. We accomplish this by modeling the distance of the isochrone as a neural network output, parametrized as a function of $\phi_1$. The result is a distance track along the stream, estimated from the variation in its isochrone with $\phi_1$.

            At any given $\phi_1$, the isochrone is modeled in absolute magnitudes
            and parameterized by the normalized stellar mass, labeled $\gamma \in (0, 1]$, in the modeled mass range. For example, increasing $\gamma$ corresponds to movement up the main sequence towards the red giant branch. We denote the isochrone as $\mcal{I(\gamma)}$. The intrinsic dipsersion around the isochrone is $\Sigma_\mcal{I}(\gamma)$. Since the isochrone is in absolute magnitudes, but the data are in apparent magnitudes,
            it is necessary to shift the isochrone by a distance modulus to the predicted distance of the stream, labeled  $\mu(\phi_1)$, with intrinsic stream distance variation $\sigma_\mu(\gamma)$.

            For the $n$-th star, the photometric model is:
            \begin{multline} \label{eq:photometric_probability_point}
                \cmp{S,m}{\pdf}(\mbs{m_i}, \gamma | \mbs{\theta}(\phi_1)) 
                \\ \sim \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1} \sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \\ \times \prior(\gamma, \nth{\mbs{m}}, \phi_1),
            \end{multline}

            where $\cmp{m}{\nth{\mbs{\Sigma}}}$ is the data covariance and $\prior(\gamma, \nth{\mbs{m}}, \phi_1)$ encodes both the present-day mass function of the stream, as well as any observational constraints (e.g. $g < 20$ mag).

            We are interested in the total probability (at some $\phi_1$) and so integrate over $\gamma$ to find:
            \begin{small}
            \begin{align} \label{eq:photometric_probability}
                &\cmp{S,m}{\pdf}(\nth{\mbs{m}} | \mbs{\theta}(\phi_1)) = \\
                &\ \int\limits_{\gamma=0}^{1} \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1}\sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \nonumber
                \\
                &\qquad \times \prior(\gamma, \nth{\mbs{m}}, \phi_1) \rm{d}\gamma. \nonumber
            \end{align}
            \end{small}
            Then, for a given $\phi_1$ we can obtain a distribution over the distance modulus using 
            \begin{align}
                \mu_{\rm sample}\left(\phi_1\right) \sim \mathcal{N}\left(\mu(\phi_1), \sigma^2_\mu(\phi_1) \right) & \quad\unit{mag}.
            \end{align}
            For each sample, we can convert the distance-modulus distribution (and error) to the distance track using 
            \begin{align}
                \rm{dist}\left(\phi_1\right) &= 10^{\frac{1}{5}\mu_{\rm sample}(\phi_1)+1} & \quad\unit{pc}, \\
                \delta \rm{dist}(\phi_1) &\simeq \frac{\ln{10}}{5} \rm{dist}(\phi_1) \ \sigma_\mu(\phi_1) & \quad\unit{pc}.
            \end{align}

            Let $\cmp{S,m}{\mbs{\theta}}(\phi_1)$ represent the feed-forward neural network which outputs a vector of parameters characterizing the stream's photometric distribution at a given $\phi_1$. The network has 2 outputs -- $\mu, \Sigma_\mu$ -- as we only consider the distance track of the isochrone, holding fixed stellar population parameters internal to modeling the isochrone, such as its age and metallicity.
            In principle we might include parameters of the isochrone in the model, now $\mcal{I}(\mbs{\theta}, \gamma)$, where $\theta = {a, Z, ...}$ and any other generating parameters of the isochrone.
            % FIXME! DO this Likewise, the dust map may be treated as a prior and it's errors incorporated into the data errors.
            However, stellar streams are generally poorly fit by theoretic isochrones, at least ones with realistic ages and metallicities.
            Consequently including the age and metallicity as model parameters
            does not meaningfully contribute to constraining results on the stream's stellar population. Moreover, age and metallicity are partially degenerate with the distance modulus. An alternate approach is to use data-driven isochrones instead of theoretic ones. In some forms of this approach the age and metallicity are not meaningful quantities. In others, e.g. with a suite of models or a trained emulator, the data-driven isochrones might have realistic ages and metallicities and thus meaningfully contribute to the model.
            % Is this true? re-read Uniform Modeling of 13 streams
            We defer this to a future work.

            The photometric probability, \autoref{eq:photometric_probability} includes the term $\prior(\gamma, \nth{\mbs{m}}, \phi_1)$, which encodes all priors on both the isochrone as well as observational constraints. It is safe to assume that the properties of the isochrone are conditionally independent from observational constraints; namely
            \begin{equation} \label{eq:isochrone_prior_components}
                \prior(\gamma, \nth{\mbs{m}}, \phi_1) = \prior_{\cal{I}}(\gamma, \phi_1) \prior_{obs}(\nth{\mbs{m}}, \phi_1)
            \end{equation}
            where $\prior_{\cal{I}}$ and $\prior_{obs}$ are the isochrone and observational priors, respectively.
            \autoref{eq:photometric_probability} already encodes the photometric distribution -- the ``track" -- of the isochrone model at each point along the track, but not the amplitudes, the relative probability of points. Physically, this amplitude is the the present-day mass function (PDMF). We include the PDMF as $\prior_{\cal{I}}(\gamma, \phi_1)$. $\prior_{\cal{I}}$ is a joint distribution, depending on $\gamma$ since it is the $\gamma$-dependent distribution amplitude, and also on $\phi_1$. This latter dependence is confusing termed the ``mass function of the stream", which is the expected variation in the average stellar mass along the stream, due to e.g. mass segregation in the progenitor \TODO{[cite]}. To date this effect has not been observed \TODO{[cite]}, so in practice we drop the $\phi_1$ dependence -- $\prior_{\cal{I}}(\gamma, \phi_1) = \prior_{\cal{I}}(\gamma)$ -- assuming that the PDMF is constant along the stream.

            % link to the code!
            In the code we offer a few ansatz for the PDMF, though any user-defined function may be used instead. The default assumption, which is known to be non-physical is a uniform distribution over $\gamma$. Similarly are truncated uniform distributions, considering only a segment of the isochrone, e.g. the main sequence turnoff (MSTO). We include also a Kroupa initial mass function (IMF) \citep{2001MNRAS.322..231K}. The Kroupa IMF, and all physically-motivated IMFs,
            have an much larger fraction of low-mass stars then high-mass stars, the latter being less likely to form. This is observed in streams, particularly their progenitors. However the Kroupa IMF, and all IMF, are poor fits to the observed PDMF as stream progenitors like globular clusters are extremely old ($\sim10$ Gyr) while an IMF is the zero-age distribution.
            How then should the PDMF $\prior_{\cal{I}}$ be treated?
            % ⬆ rephrase previous sentence
            One approach to the PDMF is to use a more informed distribution, for example time-evolving an IMF, either analytically or by simulation.
            Using any fixed distribution is simple in implementation, however the choice of distribution is not \TODO{[cite]}.
            There is large uncertainty in the PDMF of most globular clusters.
            % ⬆ rephrase previous sentence

            Similarly, the PDMF might be incorporated into the model,
            parameterized by some distribution and fit simultaneously to the distance modulus and other $\cmp{m}{\mbs{\theta}}$.
            Using a fixed distribution is a simple matter, however 
            \TODO{[cite]}

            characterizing the stream PDMF, and the

            % Mention uniform modeling?
            An alternative approach to the PDMF issue is to restrict the mass range of the isochrone such that the PDMF is unimportant.
            This restriction means we are not modeling the full range of the 
            photometric data and not

            \todo[inline]{More on the PDMF}
            \begin{enumerate}
                \item could parameterize a more general distribution and fit for the PDMF
                \item Alternatively, we are interested in the photometric model for a number of reasons, but an important one is that we can fit the distance modulus of the stream, for this only a subsection of the isochrone is needed. The best section is the MS turnoff.
                Consequently, if we focus on the MS turnoff then the variation in the PDMF is much smaller and we can use a flat distribution. (Maybe need a plot demonstrating this?)
            \end{enumerate}

            The second term in \autoref{eq:isochrone_prior_components} -- $\prior_{obs}(\nth{\mbs{m}}, \phi_1)$ -- encodes all observational effects in the photometric dataset. In particular we are

            \todo[inline]{discussion of the $\prior_{obs}(\nth{\mbs{m}}, \phi_1)$}

            \begin{enumerate}
                \item survey completeness is challenging. We pre-select the dataset so that we don't need to concern ourselves with this. For Gaia this means cutting $G_{ \Gaia} > 20$. Our discussion of the PDMF emphasizes why this isn't a problem for the isochrone.
                \item we don't actually cut the data, but apply a mask. \autoref{sub:missing_data} explains how this is done. In this case ALL the photometric feature dimensions are masked. This means that the on-stream component is a track of 0-dimensional Gaussians, which means it doesn't enter into the probability. Any astrometric data enters the astrometric model, so the membership likelihood of these stars is driven entirely by the astrometric model.
            \end{enumerate}

            \todo[inline]{Dust}
            We use \package{dustmaps} Bayestar 19 for the dust modeling. We treat it as known. Include as a source of error in $\Sigma_n$
            % TODO: actually implement that in the model

        % ssub: on_stream (end)

        \subsubsection{Linking to Astrometrics: the Distance Modulus and the Parallax}

            % TODO: rephrase
            By modeling jointly the astrometrics and photometrics we may tie together components of the models.
            In particular, the distance modulus in the photometric model and the parallax in the astrometric model are both measures of the distance track of the stream.

            We introduce a prior to connect these components of the model:
            \begin{align}
                \prior(\{\mu_I, \mu_\parallax \} | \phi_1) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1), \label{eq:distance_track_prior}
                % \prior(\{\sigma_I, \sigma_\parallax \}(\phi_1)) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1) \label{eq:distance_width_prior}
            \end{align}
            % TODO: should we show that these are different distance functions? or is it obvious
            where $\delta$ is the Dirac delta function, and simply expresses that the distance modulus track and parallax track must be equal (when converted to the actual distance) as a function of $\phi_1$.
            The widths are similarly constrained.
            \begin{align}
                \mu_I(\parallax) &= 10 - 5\log_{10}(\parallax) \\
                \sigma_I(\sigma_\parallax)  &= \frac{5}{\ln{10}} \mu_I \lvert\sigma_\parallax\rvert
            \end{align}

            In practice, at the distance of many streams the errors in the parallaxes are of order 1 and we exclude the feature from the model.
            % TODO!

        \subsubsection{Off-stream} \label{sub:photometric_model_off_stream}
            In this section we introduce our model for the distribution of backgrounds in photometric coordinates. 
            In general, the distribution of stars in color (i.e., $g-r$) and magnitude (i.e., $g$) is a complicated function of location in the galaxy \TODO{(e.g. [cite])}. Even within a field centered on the stream of interest, the color-magnitude diagram (CMD) is a combination of stellar populations distributed over a range in distances. Because of this, we do not expect an analytic density to provide a realistic description of the data. In this work we take a data-driven approach, and represent the CMD using a normalizing flow model.

            Normalizing flows provide a flexible description of a probability density, by transforming samples generated from a simple base distribution (i.e., a Gaussian) to the more complicated target distribution. The transformations are typically non-linear outputs of a neural neural network, whose parameters are optimized until the target density accurately characterizes the data. The neural network transformation is differentiable, so that jacobian factors can be combined to ensure that the the target distribution remains a valid probability density (i.e., integrates to one) \TODO{[cite]}.

            Our implementation is as follows. For a given stream, we restrict the data to some field characterized by a polygon drawn around the stream. Because our density modeling is aimed at stream characterization rather than discovery, we assume that the stream is sufficiently well known so that it can be masked out as a simple cut in (e.g.) $\phi1-\phi_2$ coordinates. With the stream masked out, we then fit a normalizing flow to the photometric data. Because the background CMD can evolve with $\phi_1$, we fit a \emph{conditional} normalizing flow, $q_{\rm flow}(\mathbf{m}|\phi_1)$, so that the color-magnitude diagram is conditionally dependent on position along the stream. In practice, we model magntiudes and not colors (i.e., $\mathbf{m} = (g,r)$ rather than $g \ \mathrm{and} \ g-r$. Otherwise, the two arguments will always be covariate). Our central assumption with this model is that the distribution of colors and magnitudes for background stars in the masked region are quantitatively similar to those in the unmasked region. For a sufficiently narrow mask we expect this to be the case, and find that this assumption does not introduce bias for the streams considered in this work.

            Once the normalizing flow, $q_{\rm flow}\left(\mathbf{m} | \phi_1\right)$, is fit to the data, its parameters are fixed while the other components of the model begin training. Therefore, the normalizing flow can be rapidly trained and computed as a pre-processing step.

            % The specific architecture of the photometric normalizing flow are discussed in \autoref{sub:model_implementation}. But
            It is easy to evaluate the log-likelihood a normalizing flow at a point $\mbs{m}$ when all feature dimensions are present. When one (or more) feature dimension is missing, however, evaluating the marginal log-likelihood involves integrals over the flow. While the procedure is straightforward it is computationally expensive. As an alternative, we conservatively set the log-likelihood to $0$, the maximum value, overemphasizing the flow's contribution to the background model. Conceptually, up-weighting the background model decreases the false-positive rate in stream membership identification, at the
            cost of increasing the false-negative rate. However, the intrinsic width of stellar streams tends to be quite small, with correspondingly small $\sigma_\mu$ (see \autoref{eq:photometric_probability}); thus the effective region of the isochrone model is well determined and we do not find the marginal log-likelihood approach to be impactful. Another approach, which we leave to a future work, is to construct normalizing flows of each marginal distribution, e.g. $q_{\rm flow}\left(\{m_i\}_{i \in \mbs{I}_F, i \neq {j, k}} | m_j, m_k, \phi_1\right)$, and select for each $\nth{\mbs{m}}$ the correct flow given the features.

        % ssub: off_stream (end)

    % sub: photometric_model (end)

    % subsection photmetric background (end)

        \begin{figure*}
            \script{mock/plot/results.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/mock/results.pdf}
            \caption{Caption: .}
            TODO:
            \begin{enumerate}
                \item the linear density in the stream-fraction plot
            \end{enumerate}
            \label{fig:mock_data_result}
        \end{figure*}

    \subsection{Priors} \label{sub:priors}

        \todo[inline]{intro sentences}

        \subsubsection{Parameter Bounds}

            \todo[inline]{this section is very rough}

            We implement a hard bound on the range of all parameters.
            Hard cutoffs can be an unwise choice of prior, but we set the ranges to larger than the data prefers and check that the trained parameters are never ``up against" the bounds.
            This makes the priors safe.
            We require these priors for training the neural network. It helps to have a well-bounded problem. Simple hard priors have advantages and disadvantages in this regard.
            An advantage is that in the allowed region the prior does not not contribute to the posterior at all. We don't want it to. We want to keep the posterior as simple as possible, driven as much as possible by the likelihood and not any priors.
            A marked disadvantage is that outside the bounded region the prior does not contribute a gradient to the loss function the optimizer can use in the fitting.
            However, we sidestep this issue by encoding the prior bounds into the ``geometry" \TODO{(What is this called?)} of the neural network, so that the network can never
            produce outputs outside the hard bounds. This is a standard technique in ML to ensure that the neural network is efficient, only covering the required region of the output space \TODO{[cite]}.

        \subsubsection{Region Prior} \label{sub:track_region_prior}

            \todo[inline]{this section is very rough}

            Region priors encourage the model to pass through a user-defined control region, while also remaining compatible with the data.

            The overall track of the stream is known, so it makes sense to give this information to the system. However, we don't want the prior to dominate the results on small scales. A good balance between these two goals is to have a prior that only influences the track on large scales.
            We introduce the ``split" Gaussian prior, which is a Gaussian split at the peak and pieced back together with a flat region connecting the halves. 
            It is smooth up to the first derivative, so works well in an auto-grad setting.
            The unnormalized PDF is described by:
            \begin{small}
            \begin{equation}
                \pdf(x,\mu,\sigma,w) \propto  \frac{1}{\sigma \sqrt{2 \pi}} \begin{cases} 
                   \Exp{-\frac{1}{2}\left(\frac{x-(\mu-w)}{\sigma}\right)^2} & \phantom{\mu - w <}\ x < \mu - w \\
                    1 & \mu - w \leq x < \mu + w \\
                    \Exp{-\frac{1}{2}\left(\frac{x-(\mu+w)}{\sigma}\right)^2} & \mu + w < x
                \end{cases}.
            \end{equation}\end{small}

            Since this prior is equi-probable in the region $\mu - w < x < \mu + w$, the  impact is to encourage the stream to fall anywhere within this region. 
            It is convenient to reparametrize, with $\tau = \frac{1}{\sigma}$.
            \begin{small}
            \begin{equation}
                \ln \pdf(x,\mu,w) \propto \tau \begin{cases} 
                    -\left(x-(\mu-w)\right)^2 & \phantom{\mu - w <}\ x < \mu - w \\
                    0 & \mu - w \leq x < \mu + w \\
                    -\left(x-(\mu+w)\right)^2 & \mu + w \leq x
                \end{cases},
            \end{equation}\end{small}
            where the additive normalization is 
            $-\ln\left(\sqrt{\frac{\pi}{\tau}} \left(\frac{1}{2}\text{erf}\left(2 \sqrt{\tau } w\right)+1\right)\right)$. In this form we see that $\tau$ may be increased to strengthen the prior and encourage $\mbs{\theta}$ to pass through

    \subsection{Model Implementation, aka the code} \label{sub:model_implementation}

        \todo[inline]{TODO: how much should be included here?}

        \begin{enumerate}
            \item A network per model component, e.g. one for the weights, one for the stream track and width, one for the background distribution.
            \item MLP for the components - lin / tanh + dropout
            \item Normalizing flow - specific pieces
            \item connective priors
            \item 
            \item Architecture in the Appendix
        \end{enumerate}

        \begin{table}
            \centering
            \caption{MLP structure}
            \begin{tabular}{@{}rrcl@{}}
            \toprule
            & Layer & Shape & Param \# \\
            \midrule
            & Linear & (None, 1, F) & 1  \\
            & Tanh & (None, 1, F) & 1  \\
            & Dropout & (None, 1, F) & 0  \\
            & Linear & (None, 1, F) & 1  \\
            \bottomrule
            \end{tabular}
        \end{table}

        Dropout layers were used to prevent network co-adaptation and over-fitting. Dropout works by randomly zeroing out nodes in the network, decreasing the importance of any single node, encouraging instead emergent patterns across the network.

% section method (end)

%\newpage
\section{Results: Mock} \label{sec:results_mock}

    In this section we present an application of our model to a synthetic stream with known ground truths, and demonstrate the model's ability to characterize stream density variations and membership probabilities in an unsupervised manner. We then apply the method to real stellar streams in \autoref{sec:results_gd1} and \autoref{sub:pal5}.

    \subsection{Test on a Synthetic Stream} \label{sub:mock_data}

        In this section we apply our model to a synethic stream in order to demonstrate an accurate recovery of stream members and density variations.

        We generate synthetic observations by sampling points from a third degree polynomial \JN{check this} with a constant scatter. Density variations are then simulated by evaluating the stream density with two Gaussian distributions centered at different values of $\phi_1$ with random amplitudes (which sum to $1$) and widths. Stream stars are then removed using random sampling, weighted by their likelihood of belonging to the density. The background astrometric  points in are generated from exponential distributions at each $\phi_1$. The number of stream stars is $\variable{output/mock/nstream_variable.txt}$, while the number of background stars is $\variable{output/mock/nbackground_variable.txt}$.

        \begin{figure}
            \script{mock/plot/photometric_background_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/mock/photometric_background_selection.pdf}
            \caption{Caption: TODO need to show this with a ground truth (e.g. show true weight in top panel).}
            \label{fig:mock_data_photometric_background_selection}
        \end{figure}

        The synthetic stream is also simulated in magnitudes using a 12~Gyr isochrone with [Fe/H] = -1.35~dex. The isochrone truncates just short of the giant branch, and sampled from uniformly (i.e., the initial mass function, $P(\gamma)$, is uniform). The background photometric coordinates are generated as a 2-dimensional Gaussian with non-zero covariance. The distance modulus is derived from a polynomial parallax track with uniform thickness. Thus, the synthetic data scenario described here generates a mock stream in astrometric coordinates and photometric coordinates, with density variations, a distance gradient which is consistent in parallax and distance modulus (i.e., magnitudes), all superimposed over a noisy field of background points in each astrometric and photometric dimension.

        The result of applying our density model to the synthetic data is illustrated in \autoref{fig:mock_data_result}. The top panel shows the weight parameter that controls membership probability as a function of $\phi_1$. The dashed curve represents the output of our data-driven model, while the solid curve is the ground truth. The second panel shows the stream and background in the $(\phi_1,\phi_2)$ plane, and the third panel shows the stream and background in the $(\phi_1,\parallax)$ plane. Points are color-coded according to their membership probability (Eq.~X) \JN{would be good to remove the $\sigma$ bands in this plot.} The weight parameter clearly shows that the model has captured the two prominent density variations along the stream. The are additional variations  mostly around the edges of the stream, though re-sampling the neural network weights with dropout reveals that these variations represent regions of higher model uncertainty.

        We find that the stream is successfully recovered in both astrometric and photometric coordinates, with $x\%$ stream stars being recovered with membership probability greater than $y$. The false positive rate (i.e., stars misidentified as stream members) for this test is found to be $z\%$ when cutting on this threshold.

        In the bottom two rows of \autoref{fig:mock_data_result} we illustrate the synthetic data color-coded by stream membership probability (top) and background membership probability (background) for different $\phi_1$ slices. Importantly, the photometric portion of the mixture model successfully identifies the stream in magnitude space, showing a high stream membership probability where the stream is present, and a low probability ($\sim 0$) where the stream is absent.

    % subsection test_on_a_synthetic_stream (end)

\section{Results: GD-1} \label{sec:results_gd1}

        We now apply our density modeling approach to the stellar stream GD-1. We discuss the data selection in \autoref{ssec:results_gd1_data} and results in \autoref{ssub:trained_gd1}.

        \subsubsection{Data Selection} \label{ssec:results_gd1_data}

            \begin{enumerate}
                \item Gaia data.
                    \begin{enumerate}
                        \item box size
                        \item selection criteria
                    \end{enumerate}
                \item crossmatched with panstarrs1
                \item extinction corrected with bayestar2019
            \end{enumerate}

            We utilize data from {\it Gaia} DR3 for the astrometric coordinates of each star, and Pan-STARRS1 to obtain more accurate color information. We adopt broad cuts on the cross-matched data to limit the field to a region known to contain GD-1. We perform these rough cuts by requiring \JN{fill in what the cuts are and reference proper motion figure and CMD figure.}

            The result is a field populated by \JN{x} stars, from which we would like to identify stream members.

        \subsubsection{Model Specification}\label{sec:GD_1_ModelSpecification}

            We model GD-1 as a three-component mixture consisting of a main stream component, a spur, and background. In order to guide our model towards GD-1, we set down 4 ``control points" along GD-1's main track, and 3 control points along the spur (illustrated in \JN{x}). These control points act as Gaussian priors on the model. We set the Gaussian error on these control points to large values, so that their influence on the model is not dominant. Indeed, we find that the value of the control points is to help reduce training time, since otherwise it will take the flexible density model longer time to converge to the actual stream.

            The weight parameter is also set to truncate to $0$ at $\phi_1 < X$ and $\phi_1 > Y$. This ensures that we do not model stream stars that fall

        \subsubsection{Trained Density Model}\label{ssub:trained_gd1}

            In this section we present results from our model applied to real {\it Gaia} and Pan-STARRS data of GD-1.

            We visualize the performance of our model in identifying stream stars and characterizing the astrometric background with \autoref{fig:gd1-results-full} and \autoref{fig:gd1-results-panels}. In \autoref{fig:gd1-results-full} The top panel shows the stream in $\phi_1$, $\phi_2$ coordinates, color-coded by membership probability (derived from \autoref{eq:membership_prob}). The following three rows represent background distributions in $\phi_2, \mu_{\phi_1}, \rm{\ and} \ \mu_{\rm phi_2}$, respectively taken at different $\phi_1$ increments centered on the red lines. The $\phi_1$ increments are roughly $\sim X$ in size.

            The blue lines over-plotted represent the best fit model at the given $\phi_1$ slice. We find that the backgrounds are typically well characterized by the distributions adopted (described in \S-x). We can also see from \autoref{fig:gd1-results-panels}
            that the stream and spur components appear to be well characterized.

            Our ability to recover the stream is demonstrated further in \autoref{fig:gd1-results-full}. In this figure, we only plot stars if their membership probability is greater than \JN{X}. The result is a ``gold" catalog of probable stream and spur members, extending over $X$-degrees in $\phi_1$. We find the expected gap around $\phi_1 \sim -20~\rm{deg}$, which has been proposed to be the location of the dissolved progenitor (CITE).

            Our model also fits the distribution of stream stars in the CMD. These fits are illustrated in \autoref{fig:gd1-results-panels}, where the CMD for the population in a given $\phi_1$ band is plotted, color-coded by stream membership probability. For the on-stream regions, the isochrone appears to provide a decent description of the stream, though there is room for further improvements (discussed in \S-x). However, we note that the stream's presence and absence in the isochrone is well captured by the mixture weight, which is a neural network output.

            Additionally, the distance modulus is also a neural network output, conditioned on $\phi_1$. This allows us to obtain a distance modulus at each $\phi_1$ along the stream. The distance modulus track and (converted) distance track are illustrated in \autoref{fig:gd1-results-full}. The left panel shows the distance modulus (in magnitudes), while the right shows the corresponding distance track. Importantly, the prior on the distance track is that the stream must fall between $[4,40]~\rm{kpc}$. Thus, our ability to recover the distance to GD-1 without substantial bias provides a useful indicator that our color-magnitude model combined with the astrometric model provide an adequate description of the data. \JN{more on this...}

            \begin{figure*}[h]
                \script{gd1/plot/data_selection.py}
                \centering
                \includegraphics[width=1\linewidth]{figures/gd1/data_selection.pdf}
                \caption{Caption}
                \label{fig:gd1-data_selection}
            \end{figure*}

            \begin{figure}
                \script{gd1/plot/photometric_background_selection.py}
                \centering
                \includegraphics[width=1\linewidth]{figures/gd1/photometric_background_selection.pdf}
                \caption{Caption}
                \label{fig:gd1-photometric_background_selection}
            \end{figure}

            \begin{figure*}[h]
                \script{gd1/plot/results_full.py}
                \centering
                \includegraphics[width=1\linewidth]{figures/gd1/results_full.pdf}
                \caption{CAPTION}
                TODO:
                \begin{itemize}
                    \item Add plot of the stream linear density
                    \item Color palette for the spur
                    \item more subtle stream MLE path
                    \item rerun model with larger pmphi2 sigma
                    \item add the isochrones to the cmd plots
                \end{itemize}
                \label{fig:gd1-results-full}
            \end{figure*}

            \begin{figure*}[h]
                \script{gd1/plot/results_panels.py}
                \centering
                \includegraphics[width=1\linewidth]{figures/gd1/results_panels.pdf}
                \caption{CAPTION}
                TODO:
                \begin{itemize}
                    \item Add plot of the stream linear density
                    \item Color palette for the spur
                    \item add the isochrones to the cmd plots
                \end{itemize}
                \label{fig:gd1-results-panels}
            \end{figure*}

            \begin{figure*}
                \script{gd1/plot/smooth_likelihood.py}
                \centering
                \includegraphics[scale=0.8]{figures/gd1/smooth_likelihood.pdf}
                \caption{Caption}
                \label{fig:gd1-heatmap}
            \end{figure*}

            \variable{output/gd1/select_members.tex}

    % subsection gd1 (end)

    \subsection{Palomar 5} \label{sub:pal5}

        Content

    % subsection pal5 (end)

% section results (end)

\section{Discussion}

    \subsection{Comparison to Other Methods} \label{sec:comparison}

        \subsubsection[STREAMFINDER]{\cite{2018MNRAS.477.4063M}}

            Assumes a potential

        % subsection streamfinder (end)

        \subsubsection{VIA MACHINAE}{\cite{2022MNRAS.509.5992S}}

            Only astrometric.

        % subsection via_machinae (end)

        \subsubsection[Uniform Modeling of 13 Stellar Streams]{\cite{PatrickEtAl2022}}

            Doesn't do simultaneous modeling.

        % subsection uniform_modeling_of_13_stellar_streams (end)

\section{Conclusions} \label{sec:conclusions}

% section comparison_to_other_methods (end)

\section{acknowledgements}

    We would like to thank the Academy.

    The data availability statement is modified from

    \paragraph{Software (alphabetical)}

        \package{asdf} \citep{Greenfield2015},
        \package{astropy} \citep{Astropy2013, Astropy2018, Astropy2022},
        \package{astroquery} \citep{Astroquery2019},
        \package{Matplotlib} \citep{Hunter2007},
        \package{NumPy} \citep{Harris2020},
        \package{Pytorch} \citep{Pytorch2019},
        \package{SciPy} \citep{Scipy2020},

\section*{Data Availability} \label{sec:data_availability}

    This study was carried out using the reproducibility software
    \href{https://github.com/showyourwork/showyourwork}{\showyourwork}
    \citep{Luger2021}, which uses continuous integration to
    programmatically download the data, perform the analyses, create the
    figures, and compile the manuscript. Each figure caption contains two links:
    one to the dataset used in the corresponding figure, and the other to the
    script used to make the figure. The datasets are stored at
    \todo[inline]{zenodo url}. The git repository associated
    with this study is publicly available at $\GitHubURL{}$.

% section data_availability (end)

\bibliography{bib}

\appendix

\section{Notation} \label{app:notation}

    In this work we use multiple data sets and a variety of mathematical tools to analyze the data, including machine learning algorithms and statistical models.
    For clarity and to connect related components of the models used in this work, a consistent mathematical notation is crucial. Here we establish the notation and symbols used throughout this work.
    \autoref{table:symbol_table} enumerates the symbols used for the statistical models and the different datasets, e.g. distinguishing astrometric from photometric data.

    \TODO{position the table here}

    \begin{table}
    \centering
    \caption{Symbol Table}
    \label{table:symbol_table}
    \begin{tabular}{p{0.35\linewidth} p{0.6\linewidth}}
    Symbol                         & Explanation \\
    \toprule
    % -----------------------------------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Datasets}:} \\
    \midrule
    % -------------------------------------------------
    \multicolumn{2}{l}{\textbf{Full Data}:}  \vspace{5pt} \\
    $\mbf{X} \in \mbb{R}^{(N, F+1)}$ & The full dataset, including all measurements (rows) and features (columns).  \\
    $\mbs{\delta}\mbf{X} \in \mbb{R}^{(N, F+1)}$ & The full set of uncertainties in $\mbf{X}$, e.g. Gaussian $\sigma_{n,f}$ and correlation terms $\sigma_{n,(f_i, f_j)}$.\\
    $N$ & The number of independent measurements (aka stars) in the dataset.  \\
    $F$ & The number of measured features, e.g. $(\alpha, \delta, \parallax, \mu_{\alpha^*}, \mu_\delta, v_r, g, r)$,  not including the the `independent' feature over which the others are conditioned. \\
    $\mrm{n} \in \mbs{I_N} \equiv \{1, ..., N\} \in \mbb{Z^+}$ & Index into the set of measurements $\mbf{X}$ (and $\mbs{\delta}\mbf{X}$), i.e. the $n$-th star. \\
    $\mrm{f} \in \mbs{I_F}  \equiv \{1, ..., F\} \in \mbb{Z^+}$ & Index into the set of features of $\mbf{X}$  (and $\mbs{\delta}\mbf{X}$). $ f=0$ is the `independent' feature.\\
    $(\delta)x_{\mrm{n,f}}$ & $n$-th star with $f$-th feature in $(\mbs{\delta})\mbf{X}$. \\
    $(\mbs{\delta})\nth{\mbs{x}}$ & $n$-th star as a row vector of all features of $(\mbs{\delta})\mbf{X}$. \\

    % -------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Astrometric Data}:}  \vspace{5pt} \\
    $\mbf{W} \in \mathbb{R}^{(N, F_w + 1)}$ & Full astrometric dataset. \\
    $\mrm{f_w} \in \mbs{I_{F_w}} \equiv \{1, ..., F_w\} \in \mathbb{Z^+}$ & Index into the set of astrometric features. $f_w = f = 0$ is the `independent' feature.\\
    $w_{\mrm{n,f_w}}$ & $f_w$-th astrometric feature of the $n$-th star in $\mbf{W}.$\\
    $\nth{\mbs{w}}$ & $n$-th star as a row-vector of all astrometric features. \\
    $\phi_1 \equiv w_{\cdot,0}$ & The `independent' (index 0) feature column on which the models are conditioned. For convenience $\phi_1$ is used instead of $\nth{\phi_1}$ (or $w_{n,0}$) except where the $n$ subscript cannot be inferred from context.\\
    % -------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Photometric Data}:}  \vspace{5pt} \\
    $\mbf{M} \in \mathbb{R}^{(N, F_m)}$ & Full photometric dataset. \\
    $\mrm{f_m} \in \mbs{I_{F_m}} \equiv \{1, ..., F_m\} \in \mathbb{Z^+}$ & Index into the set of photometric features. \\
    $m_{\mrm{n,f_m}}$ & $n$-th measurement of the $f_m$-th feature of $\mbf{M}$. \\
    $\nth{\mbs{m}}$ & $n$-th row-vector of all features. \\
    % -----------------------------------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Statistics}:} \\
    \midrule
    $\pdf(...)$ & The PDF of the distribution. \\
    $\cdf(...)$ & The CDF corresponding to $\pdf(...)$. \\
    $\mbs{\theta} = \{\theta_f\}_f$ & The vector of a distribution's parameters. \\
    $\mcal{L}(\nth{\mbf{x}}, \mbs{\theta}(\phi_1))$ & The likelihood of the data given the parameters. \\
    $Q$ & The number of models in a mixture model. \\
    $\mrm{q} \in \mbs{I_Q} \equiv \{1, ..., Q \} \in \mbb{Z^+}$ & Index into the set of $Q$ models. \\
    $\mrm{b} \in \mbs{I_Q}$ & The index of the background model in the set of $Q$ models.
    % -------------------------------------------------
    \\
    \end{tabular}
    \end{table}

    We use $\tilde{F}$ to indicate that there a missing features, e.g. radial velocities.
    The index set $\mbs{I}_F \mapsto \mbs{I}_{\tilde{F}} = \{1, ..., k-1, k+1, ... \}$ where $k$ is the index of the missing feature.

% section notation (end)

% \section{Theory of Mixture Density Networks (MDNs)} \label{app:theory_of_mixture_density_networks}

%     Mixture models (MMs) are a statistical method to represent a population as a set of a sub-populations specified by probability distributions. These distributions are characterized by their parameters and mixture by a corresponding set of mixture weights. Mixture Density Networks (MDNs) 
%     are a machine learning method to combine a MM with neural networks, increasing the flexibility of the model. An MDN replaces the scalar distribution parameters of the MM with neural networks, allowing the distribution parameters and mixture weights to vary as smooth functions \citep{Bishop1994}. Functional parameters allows the model to capture variation over those parameters that is not possible with scalar-valued models. Moreover, by using neural networks the inferred functional forms are ``model-free" and driven entirely by the data.

%     % Quote from https://mikedusenberry.com/mixture-density-networks
%     % A mixture density network (MDN) is an interesting model formalism built within the general framework of neural networks and probability theory for working on supervised learning problems in which the target variable cannot be easily approximated by a single standard probability distribution. Essentially an MDN allows one to model a conditional probability distribution as a mixture of distributions, in which the individual distributions and the corresponding mixture coefficients are parameterized by functions of the inputs

%     For the MDN we take $x_{\cdot, 0} \equiv x_{\cdot}$ (indexed from 0), which is the first feature column of $\mbf{X}$, as the `independent' coordinate over which the MDN parameters are conditioned. In the context of streams $x_{\cdot, 0} \equiv \phi_1$ is the longitude coordinate in a reference frame rotated such that the stream is aligned along $\phi_1$.

%     The PDF of the MDN is

%     \begin{equation} \label{eq:general_mixture_network}
%         \pdf(\nth{\mbs{x}} | \mbs{\theta}(\phi_1))
%         = \sum_{\mrm{q} \in \mbs{I_Q}} \ \qth{f}(\phi_1) \qth{P}(\nth{x}|\qth{\mbs{\theta}}(\phi_1),q\!=\!\mrm{q})
%     \end{equation}
%     The mixture coefficients $\qth{f}$ are normalized s.t.
%     \begin{equation}
%         \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
%     \end{equation}
%     In practice this is enforced by defining
%     \begin{equation}
%         f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash b} \qth{f}(\phi_1)
%     \end{equation}
%     for some ``background" index $b \in \mbs{I_Q}$.

%     Assuming the measurement of each datum is independent, the likelihood is a sum over all data:
%     \begin{equation}
%         \mcal{L}\left(\mbf{x}_n | \mbs{\theta}(\phi_1)\right) = \sum_{q\in\mbs{I_Q}} \qth{P}\left(\nth{\mbs{x}} | \qth{\mbs{\theta}}(\phi_1)\right).
%     \end{equation}

%     The full posterior distribution includes the priors.

%     \begin{equation}
%         \ln\posterior\left(\mbf{X} | \mbs{\theta}(\mbs{x})\right) = \ln\mcal{L}\left(\mbf{X} | \mbs{\theta}(\mbs{x})\right) +  \sum_{n \in \mbs{I_N}} \ln \left[ \sum_{q\in\mbs{I_Q}} \qth{\prior}\left(\qth{\mbs{\theta}}(\mbf{x}_n)\right) \right].
%     \end{equation}

\section{Probability Distributions} \label{app:distributions}

    A variety of probability distributions are used when modeling the stream and Galactic background. This appendix presents relevant mathematical details of these distributions.
    In particular, observational errors are modeled as Gaussian distributions, which requires modifying each distribution to account for the Gaussian noise.

    \vspace{10pt}

    Let $x \sim X$ be distributed as some specified distribution
    and $\delta \sim \Delta \equiv \mcal{N}(0, \sigma_*)$ as an independent, centered Gaussian.
    An observation $Y$ is distributed as
    \begin{equation}
        Z = X + \Delta,
    \end{equation}
    Which is a convolution of the PDFs.
    \begin{align}  \label{eq:general_convolution}
        \pdf_Z(x; \mbs{\theta}, \sigma)
            & \equiv (\pdf_X(x; \mbs{\theta}) * \mcal{N}(x; 0, \sigma)) \\
            &= \int_{-\infty}^{\infty} \rm{d}\tau \, \pdf(x; \mbs{\theta})\mcal{N}(x-\tau; 0, \sigma)
    \end{align}

    In the following subsections we work through the distributions $\pdf(x,\mbs{\theta})$ and the convolutions with $\mcal{N}(x,0,\sigobs)$.

    \vspace{5pt}
    \subsection{Flat Distribution} \label{sub:flat_distribution}

        The simplest distribution is that of the univariate uniform distribution.
        For a domain $x \in [a, b]$ the PDF is given by
        \begin{equation}\label{eq:pdf_flat_univariate}
            P_{\mcal{U}}(x|a,b) = \begin{cases}
                (b-a)^{-1} & a < x \leq b \\
                0 & \text{else}
            \end{cases}
        \end{equation}
        where $a,b$ are the bounds.

        In practice, the bounds $a,b$ are the bounds of the observation window, creating a subset of a larger field.
        Suppose the larger field is uniformly distributed in a region larger than the bounds $a, b$, ie. $P_{\mcal{U}}(x|\alpha,\beta)$.
        Let each observed datum $x_n$ in the field have Gaussian error $\sigobs_n$,
        then the PDF of the uniform-Gaussian-convolved distribution is given by:

        \begin{equation}
            \pdf_{(\mcal{U}*\mcal{N})}(x; \alpha, \beta, \sigobs) = \begin{cases}
                \pdf_\mcal{U}(x; \alpha, \beta) \left( \cdf_{\mcal{N}}(x; \alpha, \sigobs) - \cdf_{\mcal{N}}(x; \beta, \sigobs) \right) & \alpha \leq x \leq \beta \\
                0 & \text{else}
            \end{cases}
        \end{equation}

        We then truncate this distribution to the observed field $(a, b]$
        \begin{equation} \label{eq:pdf_flat_univariate_convolved_error_full}
            \pdf_{(\mcal{U}*\mcal{N})}(x | a < X \leq b; \alpha, \beta, \sigobs)
        \end{equation}
        in the normal fashion: setting $\pdf_{(\mcal{U}*\mcal{N})} = 0$ when $x < a, x > b$ and normalizing $\pdf_{(\mcal{U}*\mcal{N})}$ within $(a,b]$. The truncated distribution is tractable but unwieldy. The distribution can be simplified enormously if two conditions hold: first, $\alpha \ll a, b \ll \beta$, the observed field is a small subset of the full field; and second, 
        $\sigobs \ll |\alpha - a|$ and $\sigobs \ll |\beta - b|$, the errors are smaller than the observation window size compared to the full field.
        In this case, the distribution reduces back to the uniform distribution!

        Therefore, for all cases considered in this work \eqref{eq:pdf_flat_univariate_convolved_error_full} becomes:

        \begin{equation}\label{eq:pdf_flat_univariate_convolved_error}
            \pdf_{(\mcal{U}*\mcal{N})}(x | a < X \leq b; \alpha, \beta, \sigobs) \cong P_{\mcal{U}}(x|a,b)
        \end{equation}

    \vspace{10pt}
    \subsection{Truncated Exponential Distribution} \label{sub:exponential_distribution}

        The univariate exponential distribution is:
        \begin{equation} \label{eq:pdf_exp_univariate}
            \pdf_{\mcal{E}}(x; \lambda) = \begin{cases}
                m \Exp{-\lambda x} & x \in [0, \infty) \\
                0 & x < 0
            \end{cases}
        \end{equation}

        The support of this distribution, $[0, \infty)$, rarely matches the support of the data. We generalize to the unit-normalized, truncated, univariate exponential distribution in the domain $(a, b]$:
        \begin{equation} \label{eq:pdf_truncexp_univariate}
            \pdf_{\mcal{E}_T}(x; \lambda, a, b) = \begin{cases}
                \frac{\lambda \Exp{-\lambda \, (x - a))}}{1 - \Exp{-\lambda(b - a)}} & a < x \leq b \\
                0 & \text{else},
            \end{cases}
        \end{equation}
        for bounds $x \in (a,b]$. However, as $\lambda \rightarrow 0$, which describes the flat distribution (\autoref{sub:flat_distribution}), this functional form can be numerically unstable. It is therefore practical to define $\pdf_{\mcal{E}}(x; |\lambda| < \epsilon, a, b) \cong P_{\mcal{U}}(x; a,b)$.

        Let each observed datum $x_n$ in the field have Gaussian error $\sigobs_n$,
        then the PDF of the Exponential-convolved-with-a-Gaussian distribution is given by:

        \begin{equation}
            \pdf_{\mcal{E}\cdot\mcal{N}}(x;m,a,\sigma_*) = \frac{m  e^{\frac{1}{2} \lambda  \left(2 b+\lambda  \sigma _*^2-2 x\right)}}{e^{\lambda  (b-a)}-1} = \pdf_{\mcal{E}}(x;\lambda,a) \ \Exp{\frac{1}{2}\lambda^2 \sigma_*^2},
        \end{equation}
        where the function is no longer truncated, since a Gaussian is not compact.

        We impose the same observation field $(a, b]$ and truncate $\pdf_{\mcal{E}\cdot\mcal{N}}(x | a < X \leq b;m,a,\sigma_*)$.
        We note the normalization, which is the CDF evaluated at $a, b$, cancels the factor
        $\Exp{\frac{1}{2}\lambda^2 \sigma_*^2}$. Following thtough, the truncated convolved distribution is identical to the original distribution.

        \begin{equation}
            \pdf_{(\mcal{E}\cdot\mcal{N})_T}x | a < X \leq b;m,a,\sigma_*) \equiv \pdf_{\mcal{E}}(x; \lambda, a, b)
        \end{equation}

    \vspace{10pt}
    \subsection{Normal Distribution} \label{app:sub:normal_distribution}

        The properties of the Normal distribution are well known. For completeness we state

        \begin{equation}
            P_X(x; \mu, \sigma) = \mcal{N}(\mu, \sigma).
        \end{equation}

        With the inclusion of observational errors the distribution becomes

        \begin{equation}
            P_{\mcal{N}\cdot\mcal{N}}(x; \mu, \sigma, \sigobs) = \mcal{N}(\mu, \sqrt{\sigma^2 + \sigobs^2}).
        \end{equation}

        The Normal distribution is not compact, with non-zero value over all $x$. In an observational window $(a,b]$ the PDF is truncated, giving

        \begin{equation}
            P_{\mcal{N}_T}(x; \mu, \sigma, a, b) = \begin{cases}
                \frac{\mcal{N}(x, \mu, \sigma)}{\Phi_\mcal{N}(b; \mu, \sigma) - \Phi_\mcal{N}(a; \mu, \sigma)} & a < x \leq b \\
                0 & \text{else}
            \end{cases}
        \end{equation}

        Note that the observation-window-truncation is performed on the full distribution, which includes the observational errors, if present.
        Therefore, let $\tilde{\sigma}^2 \equiv \sigma^2+\sigma_*^2$

        \begin{equation}
            P_{(\mcal{N}\cdot\mcal{N})_T}(x; \mu, \tilde{\sigma}, a, b) = P_{\mcal{N}_T}(x; \mu, \tilde{\sigma}, a, b)
        \end{equation}

    \vspace{10pt}
    \subsection{Skew-Normal Distribution}

        The PDF of the skew-normal distribution is given by

        \begin{equation}
            \pdf_{SN}(x;\mu,\sigma,\alpha) = 2\pdf_{\mcal{N}}(x;\mu,\sigma)\cdf_{\mcal{N}}(\alpha x;\mu,\sigma).
        \end{equation}

        Despite the popularity of this distribution we are not aware of any work showing the convolution of the skew-normal distribution with a centered Gaussian. Moreover, our Mathematica (v13.2) does not successfully perform this convolution. Consequently we present a step-by-step derivation of $P_{X\cdot\mcal{N}}$.

        \autoref{eq:general_convolution} establishes that convolutions of PDFs are an integral operation. Integrals are often intractable, but the problem is solvable if recast in different terms. Rather that working with PDFs of distributions we can instead consider the moment-generating function (MGF) corresponding to the PDF's distribution. Like the PDF, the MGF uniquely determines a distribution, so a MFG uniquely maps to a PDF. For the MGF of two distributions the convolutions is a multiplication, not an integral.
        Therefore, for the convolution of two PDFs their MGFs are multiplied and if the product is the MGF of a known distribution, the corresponding PDF is the convolution of the two original PDFs.

        The MGF of the Normal distribution is

        \begin{equation} \label{eq:mgf_normal}
            M_{\mcal{N}}(t) = \Exp{\frac{\sigma_*^2 t^2}{2}}
        \end{equation}

        And the MGF of the skew-normal is

        \begin{equation} \label{eq:mgf_skewnormal}
            M_{S\mcal{N}}(t) = 2 * \Exp{\mu t + \frac{\sigma^2 t^2}{2}} \cdf(\sigma \delta t), \quad \delta = \frac{\alpha}{\sqrt{1 + \alpha^2}}.
        \end{equation}

        The product of the two MGF is

        \begin{equation}
            M_{S\mcal{N}*\mcal{N}}(t) = M_{S\mcal{N}}(t) * M_{\mcal{N}}(t) = 2 * \Exp{\mu t + \frac{(\sigma^2 + \sigma_*^2) t^2}{2}} \cdf(\sigma \delta t).
        \end{equation}
        We rewrite $\tilde\sigma^2 = \sigma^2 + \sigma_*^2$, and $\tilde{\sigma} \tilde{\delta} = \sigma \delta$.

        \begin{equation}
            M_{S\mcal{N}*\mcal{N}}(t) = 2 * \Exp{\mu t + \frac{\tilde{\sigma}^2 t^2}{2}} \cdf_{\mcal{N}}(\tilde{\sigma} \tilde{\delta} t)
        \end{equation}
        This has the same form as the skew-normal, only with $\sigma \rightarrow \tilde\sigma$ and $\delta \rightarrow \tilde\delta$.

        Therefore the PDF of the convolved distribution is

        \begin{equation}
            \pdf_{S\mcal{N}\cdot\mcal{N}}(x;\mu,\sigma,\alpha) = 2\pdf_{\mcal{N}}(x;\mu,\tilde{\sigma})\cdf_{\mcal{N}}(\tilde{\alpha} x;\mu,\tilde{\sigma})
        \end{equation}
        where $\tilde\sigma^2 = \sigma^2 + \sigma_*^2$ and $\tilde{\alpha} = \frac{\tilde{\delta}}{\sqrt{1-\tilde{\delta}^2}}$, $\tilde{\delta} = \frac{\sigma}{\tilde{\sigma}}\delta$.

        The truncation of this distribution to the range $(a, b]$ is

       \begin{equation}
            \pdf_{S\mcal{N}_T}(x; \mu, \sigma, \alpha, a, b)
                = \begin{cases}
                    \frac{\pdf_{S\mcal{N}}(x; \mu, \tilde\sigma, \tilde\alpha)}{\cdf_{S\mcal{N}}(b; \mu, \tilde\sigma, \tilde\alpha) - \cdf_{S\mcal{N}}(a; \mu, \tilde\sigma, \tilde\alpha)} & a < x \leq b \\
                    0 & \text{else}
                \end{cases}
        \end{equation}
        where
        \begin{equation}
            \cdf_{S\mcal{N}}(x; \mu, \sigma) = \frac{1}{2} \text{erfc}\left(\frac{\mu -x}{\sqrt{2} \sigma }\right)-2 T\left(\frac{x-\mu }{\sigma },\alpha \right)
        \end{equation}
        where $T$ is Owen's T distribution \citep{Owen_1956}.

% \section{Plots} \label{app:developing_plots}

%     \begin{figure*}
%         \centering
%         \includegraphics[scale=0.7]{figures-not-syw/gd1/tot_stream_and_spur_prob.png}
%         \caption{Caption}
%         \label{fig:gd1-astrometrics}
%     \end{figure*}

%     \begin{figure*}
%         \centering
%         \includegraphics[scale=0.8]{figures-not-syw/gd1/Unknown.png}
%         \caption{Caption}
%         \label{fig:gd1-distance_modulus}
%     \end{figure*}

%     \begin{figure}
%         \centering
%         \includegraphics[scale=0.6]{figures-not-syw/gd1/Unknown-2.png}
%         \caption{Caption}
%         \label{fig:gd1-cmd}
%     \end{figure}

%     \begin{figure*}
%         \centering
%         \includegraphics[scale=0.6]{figures-not-syw/gd1/Unknown-4.png}
%         \caption{Caption}
%         \label{fig:gd1-cmds}
%     \end{figure*}

\end{document}