### 2023-08-26 src/tex/ms.tex
[showyourwork] overleaf sync

% Define document class

\PassOptionsToPackage{table}{xcolor}

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath, amssymb}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage[inline]{enumitem}
\usepackage{cleveref}
\usepackage{todonotes}

% Change the section names
\renewcommand{\sectionautorefname}{\S\!}%
\renewcommand{\subsectionautorefname}{\S\!\!}%
\renewcommand{\subsubsectionautorefname}{\S\!}%
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\tableautorefname}{Tab.}

\newcommand{\package}[1]{\textsc{#1}}
\newcommand{\stream}[1]{#1}
\newcommand{\dataarchive}[1]{\texttt{#1}}
\newcommand{\Gaia}{\dataarchive{Gaia}}
\newcommand{\gaia}{\dataarchive{gaia}}

% Math fonts
\newcommand{\mrm}[1]{\mathrm{#1}}
\newcommand{\mbs}[1]{\boldsymbol{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mfk}[1]{\mathfrak{#1}}
\newcommand{\mcal}[1]{\mathcal{#1}}

% Math symbols
\newcommand{\Exp}[1]{e^{#1}}
\newcommand{\Erf}[1]{\mathrm{erf}\left(#1\right)}
\NewDocumentCommand \dif {o m} {
	\IfNoValueTF{#1}
		{ \mathrm{d}{#2} }
		{ \mathrm{d}^{#1}{\!#2} }
}

% Statistics
\newcommand{\pdf}{\mcal{P}}
\newcommand{\cdf}{\mfk{P}}
\newcommand{\likelihood}{\mcal{L}}
\newcommand{\prior}{\mcal{\pi}}
\newcommand{\posterior}{p}

% Labels
\newcommand{\nth}[1]{{#1}_{\mrm{n}}}  % measurement
\newcommand{\fth}[1]{{#1}_{\mrm{f}}}  % feature
\newcommand{\qth}[1]{{#1}_{\mrm{q}}}  % model index
\newcommand{\unit}[1]{[\text{#1}]}

\newcommand{\smallcomponent}[2]{#2^{\scriptscriptstyle (#1)}}
\newcommand{\cmp}[2]{\smallcomponent{#1}{#2}}
\newcommand{\Scmp}[1]{\cmp{S}{#1}}
\newcommand{\Bcmp}[1]{\cmp{B}{#1}}
\newcommand{\astroM}[1]{{#1}_w}
\newcommand{\photoM}[1]{\smallcomponent{m}{#1}}
\newcommand{\chemoM}[1]{\smallcomponent{z}{#1}}

\newcommand{\Spdf}{\Scmp{\pdf}}
\newcommand{\Bpdf}{\Bcmp{\pdf}}

\newcommand{\parallax}{\varpi}
\newcommand{\sigobs}{{\sigma_*}}

\newcommand{\TODO}[1]{{\textcolor{red}{#1}}}
\newcommand{\JN}[1]{\TODO{#1}}

% Begin!
\begin{document}

% Title
\title{Characterizing Stellar Streams by ML}

% Author list
\author{Nathaniel Starkman \& Jacob Nibauer}
\author{Jo Bovy}
\author{Jeremy Webb}
\author{Ana Bonaca}

% Abstract with filler text
\begin{abstract}
    Stellar stream membership likelihoods with ML. It works!
\end{abstract}

\section{Introduction} \label{sec:intro}
    %
    Stellar streams are the disrupted
    remnants of globular clusters and satellite galaxies, and their census has
    grown extensively in the Milky Way thanks to all-sky astrometric missions
    like {\it Gaia} and low-surface brightness photometric surveys (\JN{cite}).
    The formation of streams can be traced back to a progenitor cluster (i.e., a
    satellite galaxy or globular cluster), whose stars are gradually lost to the
    tidal field of the host galaxy and extend along a series of similar orbits
    producing tidal tails, which can sometimes span many tens of degrees across
    the sky. This orbital proximity of stars belonging to a stream makes them
    sensitive tracers of the mass distribution of galaxies, sourced by both the
    baryonic and dark matter components.
    % 
    The average properties of a stream (e.g., the mean position and velocity
    track of the tidal tails) can be used to reconstruct the global dark matter
    distribution of the stellar halo (e.g., \JN{cite}), while local variations
    in the stream track and surface density encode information about the
    stream's encounters with baryonic structures such as (e.g.) the stellar bar
    (\JN{cite}), and non-luminous structures like dark matter subhalos
    (\JN{cite}). Importantly, these galactic components span orders of magnitude
    in mass, highlighting the importance of characterizing a population of
    stellar streams both in morphology and kinematics. 
    % 
    While several methods have been developed to model stream properties as a
    function of global halo properties (e.g., flattening and
    orientation;\JN{cite}) and local substructures (e.g., the bar's pattern
    speed and the mass of subhalo perturbers; \JN{cite}), these methods and
    insights often rely on the existence of a homogenized stream-tracer
    population that characterizes the stream across each of the observable
    phase-space dimensions. Additionally, because even the fine-grain properties
    of streams are influenced by the detailed mass distribution of galaxies, it
    is important to propagate errors both in measurement uncertainty and
    membership uncertainty (i.e., which stars belong to the stream versus the
    background) when attempting to model the galaxy. Otherwise, a stream that
    has been poorly characterized could bias constraints on (e.g.) the total
    mass of the galaxy, the dark matter halo shape, and the mass of subhalo
    perturbers. In order to generate more precise constraints on the structure
    of the galaxy, it is therefore imperative that a statistically sound and
    homogenized catalog of stellar streams is produced so that theoretical
    models for the gravitational potential of the galaxy across a large range of
    spatial scales can be constrained. 
    %
    While streams are well described in (e.g.) action angle coordinates or
    $N-$body simulations, these methods relying on prior assumptions about the
    galaxy and its underlying potential. Therefore, it is important to devise
    data-driven methods to characterize stellar streams without appealing to
    strong prior assumptions about the galaxy's structure. 
    % 
    There are a few previous works which have modeled stellar streams in this
    context. \citet{2022MNRAS.514.1757P} developed a data-driven spline-based
    model for modeling the photometric qualities of previously identified
    streams. Their method works by fitting the stream's stellar surface density
    with a series of splines, capturing both surface brightness variations along
    the stream and changes in the position of the stream track and width. Their
    method also utilize a simultaneous ischorone fit to the stream, from which a
    distance modulus can be derived by assuming that the entire stream
    originates from the same stellar population. While this method is very
    successful at characterizing the photometric properties of a stream, it does
    not consider the kinematic dimensions which have become increasing
    well-measured with successive data releases from {\it Gaia}, as well as
    targeted observations. From a population of RR Lyrae stars,
    \citet{2019AJ....158..223P} utilized a mixture model to fit the globular
    cluster stream Pal-5 in the space of proper motions and heliocentric
    distances. The mixture modeling approach enables a simultaneous fit to both
    the stream and background density, enabling better separation of a coherent
    stream from an otherwise noisy background distribution. However, the track
    of the stream in on-sky coordinates is not fit. The method identified 27 RR
    Lyrae stars consistent with being members of the stream. While useful for
    determining the average properties of the stream (e.g., its mean distance
    track), a limited population of RR Lyrae stars are not sufficient to capture
    the fine-grained properties of streams expected to inform constraints on
    (e.g.) a population of perturbers in the galaxy. \JN{are there other
    methods?}
    % 
    Other methods for extracting stellar streams from the data include
    matched-filter approaches, where over-densities in color-magnitude space are
    inspected as possible clusters undergoing tidal stripping \JN{cite the many
    key papers using this method}. While the matched-filter technique has been
    extremely successful at stream discovery, it is not equipped to characterize
    streams at the level of detailed necessary to inform precise constraints on
    the potential of the galaxy and its substructure.
    % 
    In order to address the need for a flexible model of stellar streams both in
    photometric and kinematic spaces, in this work we develop a new method for
    characterizing streams and quantifying the membership probability of
    possible stream stars. Our approach provides several key advantage over
    previous stream modeling efforts, specifically in its ability to jointly
    model the kinematic and photometric properties of streams simultaneously.
    Furthermore, our method is well-suited to fitting streams with only partial
    or incomplete phase-space or photometric observations, enabling a full
    characterization of streams insofar as the available data allows.

    \begin{quotation}

        A stellar stream is the association of stars tidally stripped from a
        common progenitor system that orbits in an underlying galactic
        gravitational field \citep{Johnston1998, HelmiWhite1999}. Stream
        progenitors are gravitationally bound systems, such as satellite dwarf
        galaxies or star clusters \citep{Odenkirchen2001, Majewski2003}. In the
        absence of external interactions these systems are stable, though still
        capable of mass loss driven by internal processes \citep{Hills1975,
        HeggieHut2003}, e.g. relaxation and binary interactions. Considering
        external interactions, these progenitor systems may be disrupted:
        completely, as in the merger events that formed the Galactic stellar
        halo \citep{Lynden-Bell1967, Searle1978}; or partially, such as tidal
        disruption forming a stellar stream.

        In the approximation of a smooth galactic gravitational field, the
        progenitor's extent is limited to the tidal radius. Through processes
        internal and external, progenitor stars can achieve sufficient energy to
        bring them beyond the tidal boundary, thus ``escaping'' into the
        galactic field. Escape primarily occurs at both an inner and outer
        saddle point of the combined potential -- e.g. the L2 and L3 Lagrange
        points for a spherical galactic field \citep{Ross1997, Fukushige2000,
        BinneyTremaine2008}. Generally, the initial phase space position of the
        escaped star is very similar to that of the progenitor: the
        configuration position is the tidal boundary and the star has a low
        peculiar velocity. Since both the escaped star and its progenitor orbit
        in the same underlying galactic potential and have similar phase space
        positions their orbits are similar \citep{BinneyTremaine2008}. Given
        sufficient time the phase space separation will increase and an escaped
        star will lead or trail the position of the progenitor, depending on the
        location from which it escaped. Therefore, a collection of tidally
        stripped stars will form extended structures, aka tails or streams, that
        are few to thousands of parsecs in length \citep{Johnston1998,
        HelmiWhite1999, Bovy2014}.

        To be a coherent, observable structure the peculiar velocity dispersion
        of escaped stars relative to the progenitor must be relatively low, but
        does vary from system to system. Stars in kinematically hot streams
        (e.g. from progenitors with larger kinematic dispersions) have more
        dissimilar orbits from each other and the progenitor than do stars in
        cold streams, with smaller dispersions \citep{Johnston1998, Hendel2015}.
        The orbit of a star depends on its phase space position and the
        gravitational potential; therefore measuring a star's position and
        velocity and observing its orbit reveals the local potential, which is
        normally dominated by the Galaxy (primarily the dark matter halo at
        large distances) \citep{Gibbons2014}. However, an orbit is impractical
        to observe on human timescales. Cold streams offer a workaround.  Like
        measuring one star at many points in one orbit over an extended period,
        streams are many stars measured at different points along approximately
        one orbit at the same time. Therefore stellar streams are one of the
        most promising means to study the Galactic gravitational potential: the
        stellar components as well as the dark matter \citep{Johnston2016}.

        On the observational front there has been a large concerted effort in
        the field to find new stellar streams \citep[for an extensive list of
        known Galactic streams see][]{Mateu2022}, extend the detected extents of
        known streams \citep[e.g. for \stream{Palomar 5} alone:][]{Rockosi2002,
        Ibata2017, Grillmair2006, Carlberg2012, StarkmanEtAl2019}, and increase
        the general purity of stream catalogs.

    \end{quotation}

    \todo[inline]{--- Rephrase}

    \paragraph{one orbit and a single population}

        Stellar streams consist of stars that are tidally disrupted from a
        progenitor system.  To a first approximation, the stars occupy a single
        orbit, that of their progenitor.  Additionally, as stars formed in the
        progenitor they retain its photometric and chemical characteristics:
        which for globular clusters means following an isochrone.  In this
        simplistic picture, astrometrically characterizing a stream requires
        only finding it's orbit. Member stars fall directly on the orbit and the
        isochrone, making membership likelihood determination trivial.

        % Observational errors complicate this picture. Instead of a certain
        % membership calculation, we must build a probabilistic model, assigning
        % a likelihood that an observed star falls along the stream's orbit.

    \paragraph{similar orbits and a somewhat noisy population}

        Occupying a single orbit implies that streams remain a point in action
        space, and having a common origin would thus not become extended
        structure. This is not true.  Differences in stream stars' orbits from
        the progenitor's arise from various factors including internal dynamics
        of the progenitor and the host Galactic potential.

        An intrinsic width, in configuration and kinematic space.

        Centered on a track, not an orbit.

        This defines an astrometric phase-space hyper-tube
        \cite{2018MNRAS.477.4063M}, with a distribution radial to the track.

        Instead of characterizing the stream by its orbit, we characterize it by
        it's track and the distribution around that track.  Computing membership
        likelihoods now requires fitting the track and distribution. This cannot
        be done without then also fitting the field of stars in which the stream
        lives, since the stream and field cannot be clearly separated.

    \paragraph{gaps orbits and a somewhat noisy population}

        This is still a simplistic picture of streams.

        Perturbations to the Galactic potential, in the form of time evolution
        and non-axisymmetry, as well as substructure can disrupt the stellar
        stream.  Even most perturbations don't ruin the idea of the phase-space
        hypertube. They either drop the mixture weight (rephrase, not yet
        introduced) to 0 or change the width of the hypertube.

    \paragraph{Review of detection methods}

        Identifying streams of globular of clusters is a challenging task.
        Streams are very low signal-to-noise structures spread over
        comparatively large regions of the sky. They live in a vast background
        of field stars.

        Historically streams were detected by matched-filter techniques
        % Add citations
        You took an isochrone and shifted it by a distance modulus until a
        stream stood out in longitude and latitude as an extended overdensity
        structure.

        With the advent of Gaia, we gained access to detailed astrometric data.
        For most streams this

        poses a challenging task due to their intricate and diffuse nature from
        large-scale surveys is challenging due to the complex and overlapping
        distributions of stars in different regions of the sky; disentangling
        members from the vast background of field stars

        \begin{enumerate}
            \item Streamfinder \citep{2018MNRAS.477.4063M}. This method requires the potential
            \item VIA MACHINAE \citep{2022MNRAS.509.5992S}
            \item Kiyan and Adrian
            \item Matched Filter
            \item Gaia stuff: Adrian and Ana and Sara et al
            \item Uniform Modeling of 12 Stellar stream \citep{PatrickEtAl2022}. Only on phi1, phi2, and CMD. Their CMD method is more sophisticated. Worth discussing.
        \end{enumerate}

    Review of available data:
    \begin{enumerate}
        \item Gaia. Streams in Gaia have on-sky positions and proper motions, and parallaxes, though for most streams the parallax errors are prohibitively large. Likewise, most streams in Gaia do not have good radial velocity measurements. The photometry is generally poor, generally necessitating cross-matching with other datasets.
        \item S5. This is a much smaller dataset. It has deeper photometry and radial velocity measurements.
    \end{enumerate}

    \paragraph{Paper organization}

        The paper is organized as follows.  In \autoref{sec:method} we introduce
        our method for characterizing stream tracks and computing stellar
        membership.  We discuss the data in \autoref{sub:method:framing_the_data} and
        build a probabilistic framework to model the stream and background in
        \autoref{sub:method:likelihood_setup}-\autoref{sub:methods:priors}.  In
        \autoref{sec:results_mock} we apply our method to simulated data to
        demonstrate the model with a known ground truth dataset, then to the
        stellar streams \stream{GD-1} (\autoref{sec:results_gd1}) and
        \stream{Pal\,5} (\autoref{sec:results_pal5}).  We compare our approach
        to characterizing stellar streams to other methods in
        \autoref{sec:comparison}.  We conclude in \autoref{sec:conclusions},
        discussing the results of this work and future directions.

% section introduction (end)

\section{Method} \label{sec:method}

    We now introduce our method for characterizing stellar stream tracks and
    computing stream membership likelihoods from N-D phase-space measurements.
    % TODO: rephrase following sentence
    We develop a flexible Bayesian model framework, that allows arbitrary
    numbers of phase-space dimensions to be incorporated in the model and
    permits missing phase-space observations.  Below, we offer a concise
    overview of our approach, followed by a more in-depth technical explanation
    in subsequent sections.  We start, in \autoref{sub:method:framing_the_data}, by
    transforming the astrometric data into a stream-oriented on-sky reference
    frame ($\phi_1, \phi_2, \parallax, \mu_{\phi_1}^*, \mu_{\phi_2}, v_r)$, for
    each stream. In this frame, the stream track lies along $\phi_1$, and the
    stream density may be described by distributions in the other parameters as
    injective functions of $\phi_1$.
    % TODO: ⬆ I think a few more sentences explaining why this is important,
    % feature variations, etc?  However, previously we discussed how stream will
    % vary along the track, with changes in the width and possible gaps.  Also
    % observational effects, e.g the distance gradient changes the on-sky width
    % (should this be mentioned earlier?)
    Likewise, the background of the Galactic field may be described with
    $\phi_1$-conditioned distributions.
    % TODO: ⬆ I think this should be expanded.
    Non-astrometric coordinates, e.g. photometry and metallicity, are also
    modelled along the stream and background as functions of $\phi_1$.

    The distributions for the stream and background are collected in a single
    mixture model that describes the entire field.  We take the parameters of
    the mixture model -- the mixing coefficients and component distribution
    parameters -- to be general (continuous) functions of the conditioning
    variable, $\phi_1$. This is done using feed-forward neural network that
    takes the conditioning variable as its input and output the parameter value.
    Mixture models of conditional probability distributions using neural
    networks are called Mixture Density Networks (MDNs). Since the model
    coefficients may evolve with $\phi_1$, MNDs are well suited to characterize
    streams (and the background) with all their variations in position, width,
    and linear density.

    % sub: framing_the_data (end)

    \subsection{Framing the Data} \label{sub:method:framing_the_data}

        As a stream progenitor orbits the Galactic center of mass (GCOM), its
        orbital path describes an ellipse in the orbital plane. The associated
        arms of the stream generally align closely with its orbital plane.  From
        a Galactocentric reference frame, an orbital plane assumes the form of a
        great circle in the observer's celestial ``sky," thereby allowing the
        potential realignment of longitudinal ($\ell$) and latitudinal ($b$)
        coordinates to a coordinate system aligned with the stream
        ($\cmp{GCOM}{\phi_1}, \phi_2^{(GCOM)}$). Here, the progenitor is
        established at the origin, with the stream progressing along
        $\cmp{GCOM}{\phi_1} (\phi_2^{(GCOM)} \approx 0)$. $\cmp{GCOM}{\phi_1}$
        is clearly a good affine parameter along which to construct models and
        analyze the stream's characteristics.

        However, data observations are not taken from the GCOM, but from  $\sim
        8$ kpc distant \citep{GRAVITY2018, Leung2022} in a local frame,
        typically from vantage points such as Earth or its solar Lagrange
        points, and subsequently represented in the International Celestial
        Reference System (ICRS) \citep{ICRS1997}.  Restricted as we are to a
        very local volume, there is no means to observe systems from the more
        ``intrinsic" reference frames. In our observational frames there is no
        inherent assurance of effecting a rotation that would similarly align
        the stream within observer coordinates such that $\phi_2(\phi_1) \approx
        0$.  Thankfully, most streams are adequately distant from both the
        Galactic center and the observer, or exhibit orientations such that a
        rotation a) can be established or b) holds valid. For most of the
        best-studied globular cluster streams -- \stream{GD-1}, \stream{Palomar
        5}, \stream{Jhelum}, etc. -- this holds true. However for some streams
        of dwarf galaxies, like the Magellanic stream which wraps multiple times
        around the Milky Way \citep{Wannier1972}, no such transformation is
        possible. More generally, a stream's ``sky"-projected path may be
        kinked, e.g. by subhalo interactions, such that it is a many-to-one
        function in $\phi_1$.

        In this work we treat with the vast majority of streams, which have a
        ``sky" transformation such that the stream is an injective function in
        $\phi_2(\phi_1)$.  Since we are examining known streams this
        transformation is known \textit{a priori}, or may be constructed during
        pre-processing \citep[see][\S2.1]{StarkmanEtAl2023}.  We note that, with
        sufficient prior knowledge, many-to-one functions may be broken into
        injective segments and their connection constrained by a prior. Thus the
        methods developed subsequently may extented and applied even to wrapped
        and kinked streams. However, we leave extending the model framework to a
        future work.

    \subsection{Likelihood Setup}\label{sub:method:likelihood_setup}

        \begin{figure}
            \centering
            \includegraphics[scale=0.5]{static/model/DAG_StreamDensity.pdf}
            \caption{Probabilistic Graphical Model of the Mixture Density Network (MDN)}
            \todo[inline]{TODO:}
            \begin{enumerate}
                \item remake this example this with \url{https://github.com/daft-dev/daft}. See \url{https://en.wikipedia.org/wiki/Mixture_model} for examples.
            \end{enumerate}
            \label{fig:DAG}
        \end{figure}

        Mixture models (MMs) are a statistical method to represent a population
        as a set of a sub-populations specified by probability distributions.
        These distributions are characterized by their parameters and the
        mixture by a corresponding set of mixture coefficients. Scalar mixture
        models may be extended to model completely general conditional
        distributions by making the distribution parameters and mixture
        coefficients into functions of a conditional input
        \citep{McLachlan1989MixtureM}.  Mixture Density Networks (MDNs) are a
        machine learning method to implement these general MM, taking the
        parameters and coefficients to be the outputs of feed-forward neural
        networks and as such, general continuous functions \citep{Bishop1994}.
        Functional parameters allow the model to capture variation over those
        parameters that is not possible with scalar-valued models. Moreover, by
        using neural networks the inferred functional forms are ``model-free"
        and driven entirely by the data.

        For the MDN we take $x_{\cdot, 0}$, which is the first feature column of
        the data $\mbf{X}$, as the `independent' coordinate over which the MDN
        parameters are conditioned. In the context of streams $x_{\cdot, 0}
        \equiv \phi_1$ is the longitude coordinate in a reference frame rotated
        such that the stream is aligned along $\phi_1$. The parameters of the
        stream, the background field, and the mixing coefficients of the two are
        now general data-driven continuous functions along $\phi_1$.

        The PDF of a general MDN is
        \begin{equation} \label{eq:general_mixture_network}
            \!\!\! \pdf(\nth{\mbs{x}} | \mbs{\theta}(\phi_1))
            \!=\! \sum_{\mrm{q} \in \mbs{I_Q}} \! \qth{f}(\phi_1) \cmp{q}{\pdf}(\nth{\mbs{x}}|\qth{\mbs{\theta}}(\phi_1)),
            % \prior(\nth{\mbs{x}}, \phi_1)
        \end{equation}
        where $\cmp{q}{\pdf}$ is the PDF of the $q$-th model and $f_q$ its
        mixing coefficient for all $q$ models in the set of models indexed by
        $\mbs{I_Q}$. $\mbs{\theta}$ is the $\phi_1$-conditional output of the
        feed-forward neural network, and $\nth{\mbs{x}}$ the data on the $n$-th
        star in $\mbs{X}$, not including $\phi_1$. For convenience we drop the
        implied subscript $n$ in $\phi_1$, the explicit notation of
        $\mbs{\theta}$'s dependence on $\phi_1$, and the index set $\mbs{I}_X$
        in summations over its elements $x$. A full detailing of notation and
        symbols is included in \autoref{app:notation}.

        The mixture coefficients $\qth{f}$ are normalized s.t.
        \begin{equation}
            \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
        \end{equation}
        In practice this is enforced by defining
        \begin{equation} \label{eq:mixture_weight_normalization}
            f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash \{b\}} \qth{f}(\phi_1),
        \end{equation}
        the summation over all models except for some ``background" index $b \in
        \mbs{I_Q}$.  We discuss the mixture weights further in
        \autoref{sub:method:mixture_weight}.

        In this work we consider primarily astrometric and photometric data,
        (notated $\mbs{W}$ and $\mbs{M}$, respectively, as detailed in
        \autoref{app:notation}), but note that the MDN framework may be extended
        to arbitrary feature dimensions.  Assuming the conditional independence
        of the astrometry and photometry, we split each model into linearly
        separable astrometric and photometric models $\cmp{q}{\pdf} =
        \cmp{q,w}{\pdf} \cmp{q,m}{\pdf}$ .

        The likelihood for a single star then becomes, simplifying from
        \autoref{eq:general_mixture_network},
        \begin{equation} \label{eq:general_model}
            \!\!\!\!\! \pdf(\nth{\mbs{w}}, \! \nth{\mbs{m}} | \mbs{\theta})
                \!=\! \sum_{q} \! f_q(\phi_1) \cmp{q,w}{\pdf}(\nth{\mbs{w}}|\mbs{\theta}) \cmp{q,m}{\pdf}(\nth{\mbs{m}}|\mbs{\theta})
        \end{equation}

        Assuming the measurement of each star is independent, the total
        log-likelihood is a sum over all stars:
        \begin{equation} \label{eq:general_likelihood}
            \ln\mcal{L}\left(\{\nth{\mbs{w}},\nth{\mbs{m}}\} | \mbs{\theta}\right) = \sum_n \ln \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}).
        \end{equation}

        It will prove convenient to distinguish the models for $q$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:
        \begin{align} \label{eq:stream_and_bkg_prob}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} \Scmp{P}(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) \Bpdf(\nth{\mbs{w}}|\mbs{\theta}) \Bpdf(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}
        where we have dropped the $(w), (m)$ superscript as the model types are
        evident from their inputs. We applied
        \autoref{eq:mixture_weight_normalization} to require only a single
        mixture coefficient.  Of course, if more components are included in the
        mixture then \autoref{eq:stream_and_bkg_prob} becomes the more general
        \autoref{eq:general_model}. However, this an implementation detail as
        any more complex model, e.g a model including a cold stream, extra-tidal
        features, the background stellar halo, and field globular clusters may
        be rearranged such that all stream-related components (cold or
        extra-tidal) and all other components (stellar halo and globular
        clusters) are grouped into multi-component \textit{stream} and
        \textit{background} models which themselves adhere to a
        \autoref{eq:general_model}-type relation.

        The probability of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\left(S | \nth{\mbs{w}}, \nth{\mbs{m}}, \mbs{\theta} \right) = \frac{f \Spdf(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}

        Hereafter, when we refer to ``membership probability,"
        \autoref{eq:membership_prob} provides the formal definition. We nuance
        this definition by considering missing feature dimensions in the
        following subsection.

    \subsection{Missing Phase-Space Observations} \label{sub:method:missing_data}

        Streams have a diversity of measurement coverage: several have full 6-D
        astrometrics as well as photometrics and chemical abundances available
        \citep[e.g.,][]{Koposov2019, 2020A&A...635L...3A, 2022ApJ...928...30L},
        while most have partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \gaia {\color{red} [cite]}, for instance, have proper
        motions but not radial velocities nor reliable parallax measurements.
        Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many
        feature dimensions, however only for a select set of streams. Given the
        diversity of coverage it is important that the likelihood model --
        \autoref{eq:membership_prob} -- account for missing data.

        The method of treatment of missing data depends on the cause for that
        data to be missing.  Broadly, there are two categories of causes:
        randomness and systematics.  Consider a dataset with all features
        measured but with a random subset of features masked.  One possible
        approach is to attempt to impute the missing features, then feed the
        full-featured data to the likelihood model. Alternatively, the
        likelihood model might use only the present features, ignoring those
        missing. If however, features are missing for systematic reasons, then
        this approach can introduce systematic bias. Consider a star too far
        away to measure it's parallax distance or radial velocity. This is a
        systematic reason and this information should be incorporated into any
        model as a prior. Not doing so, the data imputation approach might
        predict too-close distances. Likewise, using only present features in
        the likelihood model misses that the star is likely very distant.  The
        effects of mask systematics are mitigated by not including such data,
        moving the contribution of the systematics from a prior to the evidence.
        Our models compute likelihoods under a specific dataset, and
        \autoref{eq:membership_prob} is a ratio of the probabilities, so the
        removed systematics are not expected to be impactful.

        We adopt the approach of modeling the data masks as randomly-distributed
        and not systematic. We do not impute missing data, instead the per-star
        likelihood is modelled with a distribution with the same dimensionality
        as the data vector.  For example, a star with a 6D astrometric
        measurement will be modelled with a full 6D distribution, while a star
        missing its radial velocity measurement will be modelled with a
        dimensionally-reduced (5D) form of the 6D distribution. Using
        dimensionally reduced distributions ensure that missing phase-space
        dimensions do not amplify nor suppress the likelihood.  Notwithstanding,
        we include a quality flag on our calculations indicating the number of
        missing features.  We leave incorporating priors on the data mask
        distribution to future work.

    % sub: likelihood_setup (end)

    \vspace{10pt}
    \subsection{Determination of the Mixture Weight} \label{sub:method:mixture_weight}

        The mixture weight $f_q(\phi_1)$ as used in \autoref{eq:general_model}
        is characterized as the fraction of stars belonging to the mixture
        component $q$ over an increment $[\phi_1, \phi_1 + \dif{\phi_1}]$. Thus,
        the mixture weight encodes information about the linear density of the
        stream, which has been shown to depend on substructure in the galaxy
        (CITE-x). The dependence of the mixture weight on the linear density is
        captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(w_{i,0}\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$
        within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and $N_{\rm tot}$
        is the total number of stars in the same band (including those in
        component $q$). The linear  density for component $q$ is then entirely
        specified by the function $N_q(\phi_1)$.

        The mixture model specified in \S-X only models the weight parameter
        $f_q$, and not the components of the ratio in
        \autoref{eq:fraction_param}. However, the linear density $N_q(\phi_1)$
        can be readily obtained from $f_q$, provided that $N_{\rm
        tot}\left(\phi_1\right)$ can be calculated. Importantly, this function
        does not depend on the parameters of the on-stream or off-stream model.
        Instead, $N_{\rm tot}(\phi_1)$ is simply the number of stars in a small
        $\phi_1$ increment. We therefore model $N_{\rm tot}\left(\phi_1\right)$
        as a post-processing step using a single normalzing flow in $\phi_1$.

        In particular, for a given field the distribution of $\phi_1$
        coordinates is specified by the set $\{\phi_1\}_f$. A normalizng flow is
        trained to this dataset, and will satisfy $\int d\phi_1
        P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the
        normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale
        $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of
        stars in the field considered. We can then obtain the linear stream
        density with
        \begin{equation}
            N_q\left(\phi_1\right) = f_q N_{\rm tot} \left(\phi_1\right)  = f_q N_{\rm tot,0} P_{\phi_1}\left(\phi_1\right) .
        \end{equation}

    \subsection{Astrometric Model} \label{sub:method:astrometric_model}

        \subsubsection{On-Stream} \label{ssub:method:astrometric_model:on_stream}

            The astrometric track is modeled with a neural network, encompassing
            each phase space dimension, parameterized as a function of $\phi_1$.
            The neural network output for a single stream component is the mean
            location of the stream track in each dimension,
            $\astroM{\mbs{\mu}}(\phi_1)$, and its associated covariance matrix
            $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is
            then
            \begin{equation}
                \cmp{S,w}{\pdf}(\nth{\mbs{w}} | \mbs{\theta}) \! = \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}}\!(\phi_1), {\cmp{w}{\nth{\mbs{\Sigma}}}}^2 \! + \!  \astroM{\mbs{\Sigma}}^2\!\left(\phi_1\right) \!\big),
            \end{equation}
            where $\astroM{\mbs{\theta}}$ is the set of neural network
            parameters that define $\astroM{\mbs{\mu}}$ and
            $\astroM{\mbs{\Sigma}}$; and $\cmp{w}{\nth{\mbs{\Sigma}}}$ the data
            uncertainties.  In the simplest case, we take
            $\cmp{w}{\nth{\mbs{\Sigma}}}, \astroM{\mbs{\Sigma}}$ to be diagonal,
            consisting of both intrinsic dispersion for each dimension of
            $\mbs{w}$ and the data uncertainties added in quadrature. The data
            uncertainties $\cmp{w}{\nth{\mbs{\Sigma}}}$ are drawn from
            $\mbs{\delta}\mbf{W}$, the dataset of astrometric uncertainties (see
            \autoref{app:notation}).

            In practice, we work with truncated Gaussians, where the truncation
            is set by the size of the field in each astrometric dimension.
            Details of the truncated Gaussian are included in
            \autoref{app:sub:normal_distribution}. Truncating to the field
            ensures that the probability density is zero wherever there is no
            data. This has an added benefit of correctly normalizing the
            Gaussian distribution, which is not compact (has non-zero
            probability over all space) and would otherwise not integrate to one
            within the field.

            As a MDN, we implement the Gaussian distribution parameters with a
            multi-layered feed-forward neural network. The number of layers
            required depends on the maximum dimensionality of the Gaussian, and
            whether the neural network employs regularization techniques like
            dropout. We use alternating layers of linear and tanh units. Details
            of the network architecture are discussed in
            \autoref{sub:methods:model_implementation}.

            In the presence of missing phase-space coordinate $f_k$ we reduce
            the dimensionality of the Gaussian to that of the data-vector.
            Namely, for $\mbs{I}_{\tilde{F}} \equiv \mbs{I}_{F} \backslash
            \{f_k\}$
            \begin{align}
                \nth{\tilde{\mbf{w}}} &\mapsto \nth{\tilde{\mbf{w}}} = \{ w_{n,f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\mu}} &\mapsto \astroM{\tilde{\mbs{\mu}}} = \{ \mu_{f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\Sigma}} &\mapsto \astroM{\tilde{\mbs{\Sigma}}} = \{ \Sigma_{i,j} | i,j \in \mbs{I}_{\tilde{F}} \}, & \nonumber
                \intertext{such that,}
                \Spdf(\nth{\mbs{w}} & | \astroM{\mbs{\theta}}(\phi_1))
                    \mapsto \Spdf(\nth{\tilde{\mbs{w}}} | \astroM{\tilde{\mbs{\theta}}} (\phi_1)) = \mcal{N}_{\tilde{F}}.
            \end{align}

        % ssub: on_stream (end)

        \subsubsection{Off-stream} \label{ssub:method:astrometric_model:off_stream}

            We now outline the off-stream astrometric model. We take a similar
            approach to the on-stream astrometric model discussed in
            \autoref{ssub:method:astrometric_model:on_stream}, though we
            typically do not characterize the backgrounds with a Gaussian
            distribution. Instead, we utilize a range of distributions (i.e.,
            exponential, skew-normal) discussed in \autoref{app:distributions}.
            The parameters of the user-specified background distributions are
            themselves neural network outputs, such that any $\phi_1$ ``slice"
            of the background should be characterized by some analytic density,
            but the evolution of the backgrounds with $\phi_1$ can be complex.
            Details of the network architecture are discussed in
            \autoref{sub:methods:model_implementation}.

            Let $\Bcmp{\fth{\mbs{\theta}}}(\phi_1)$ represent the neural network
            which outputs a vector of parameters characterizing the background
            distribution at a given $\phi_1$ for astrometric dimension $f$. The
            number of outputs for the background model is equal to
            $\mathrm{dim}\left(\fth{\Bcmp{\mbs{\theta}}}\right)$ (where $f=0$
            corresponds to $\phi_1$, which we do not model). At a given
            $\phi_1$, the probability density for the astrometric dimension
            $f>0$ is $\Bcmp{\fth{\pdf}}\left(\mbs{w}_{n,f} |
            \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right)$. We treat each background
            component as independent, allowing us to write the likelihood as a
            product over the astrometric dimensions:
            \begin{equation} \label{eq:astrometric_model_off_stream_probability}
                \pdf^{(B)}\left(\nth{\mbs{w}} | \mbs{\theta}(\phi_1) \right) = \prod_{f\in \mbs{I}_{\tilde{F}}} \pdf_f^{(B)}\left(\mbs{w}_{n,f} | \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right).
            \end{equation}
            Where $\mbs{I}_{\tilde{F}}$ is the set of indices of the features,
            omitting missing phase space dimensions.

        % ssub: off_stream (end)

    % sub: astrometric_model (end)

    \subsection{Photometric Model} \label{sub:method:photometric_model}

        \subsubsection{On-Stream} \label{sub:method:photometric_model:on_stream}

            In photometric coordinates we model the stream as a
            single-population isochrone with the possibility of a non-zero
            distance gradient. We accomplish this by modeling the distance of
            the isochrone as a neural network output, parametrized as a function
            of $\phi_1$. The result is a distance track along the stream,
            estimated from the variation in its isochrone with $\phi_1$.

            At any given $\phi_1$, the isochrone is modeled in absolute
            magnitudes and parameterized by the normalized stellar mass, labeled
            $\gamma \in (0, 1]$, in the modeled mass range. For example,
            increasing $\gamma$ corresponds to movement up the main sequence
            towards the red giant branch. We denote the isochrone as
            $\mcal{I(\gamma)}$. The intrinsic dipsersion around the isochrone is
            $\Sigma_\mcal{I}(\gamma)$. Since the isochrone is in absolute
            magnitudes, but the data are in apparent magnitudes, it is necessary
            to shift the isochrone by a distance modulus to the predicted
            distance of the stream, labeled  $\mu(\phi_1)$, with intrinsic
            stream distance variation $\sigma_\mu(\gamma)$.

            For the $n$-th star, the photometric model is:
            \begin{multline} \label{eq:photometric_probability_point}
                \cmp{S,m}{\pdf}(\mbs{m_i}, \gamma | \mbs{\theta}(\phi_1)) 
                \\ \sim \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1} \sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \\ \times \prior(\gamma, \nth{\mbs{m}}, \phi_1),
            \end{multline}

            where $\cmp{m}{\nth{\mbs{\Sigma}}}$ is the data covariance and
            $\prior(\gamma, \nth{\mbs{m}}, \phi_1)$ encodes both the present-day
            mass function of the stream, as well as any observational
            constraints (e.g. $g < 20$ mag).

            We are interested in the total probability (at some $\phi_1$) and so
            integrate over $\gamma$ to find:
            \begin{small}
            \begin{align} \label{eq:photometric_probability}
                &\cmp{S,m}{\pdf}(\nth{\mbs{m}} | \mbs{\theta}(\phi_1)) = \\
                &\ \int\limits_{\gamma=0}^{1} \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1}\sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \nonumber
                \\
                &\qquad \times \prior(\gamma, \nth{\mbs{m}} | \phi_1) \rm{d}\gamma. \nonumber
            \end{align}
            \end{small}
            Then, for a given $\phi_1$ we can obtain a distribution over the
            distance modulus using 
            \begin{align}
                \mu_{\rm sample}\left(\phi_1\right) \sim \mathcal{N}\left(\mu(\phi_1), \sigma^2_\mu(\phi_1) \right) & \quad\unit{mag}.
            \end{align}
            For each sample, we can convert the distance-modulus distribution
            (and error) to the distance track using 
            \begin{align}
                \rm{dist}\left(\phi_1\right) &= 10^{\frac{1}{5}\mu_{\rm sample}(\phi_1)+1} & \quad\unit{pc}, \\
                \delta \rm{dist}(\phi_1) &\simeq \frac{\ln{10}}{5} \rm{dist}(\phi_1) \ \sigma_\mu(\phi_1) & \quad\unit{pc}.
            \end{align}

            Let $\cmp{S,m}{\mbs{\theta}}(\phi_1)$ represent the feed-forward
            neural network which outputs a vector of parameters characterizing
            the stream's photometric distribution at a given $\phi_1$. The
            network has 2 outputs -- $\mu, \sigma_\mu$ -- as we only consider
            the distance track of the isochrone, holding fixed stellar
            population parameters internal to modeling the isochrone, such as
            its age and metallicity.  In principle we might include parameters
            of the isochrone in the model, now $\mcal{I}(\mbs{\theta}, \gamma)$,
            where $\theta = {a, Z, ...}$ and any other generating parameters of
            the isochrone.
            % FIXME! DO this Likewise, the dust map may be treated as a prior and it's errors incorporated into the data errors.
            However, stellar streams are generally poorly fit by theoretic
            isochrones, at least ones with realistic ages and metallicities.
            Consequently including the age and metallicity as model parameters
            does not meaningfully contribute to constraining results on the
            stream's stellar population. Moreover, age and metallicity are
            partially degenerate with the distance modulus, working counter to
            the goal of fitting a distance track to the stream. An alternate
            approach is to use data-driven isochrones instead of theoretic ones.
            the age and metallicity are not meaningful quantities. In others,
            e.g. with a suite of models or a trained emulator, the data-driven
            isochrones might have realistic ages and metallicities and thus
            meaningfully contribute to the model.
            % Is this true? re-read Uniform Modeling of 13 streams
            We defer this to a future work.

            The photometric probability, \autoref{eq:photometric_probability}
            encodes the photometric distribution -- the ``track" -- of the
            isochrone model at each point along the track, as well as the
            amplitudes, the relative probability of points along the track. This
            is the term $\prior(\gamma, \nth{\mbs{m}} | \phi_1)$, which
            describes all priors on both the isochrone as well as observational
            constraints. We assume that the properties of the isochrone are
            conditionally independent from observational constraints; namely
            \begin{equation} \label{eq:isochrone_prior_components}
                \prior(\gamma, \nth{\mbs{m}} | \phi_1) = \prior_{\cal{I}}(\gamma | \phi_1) \prior_{obs}(\nth{\mbs{m}} | \phi_1)
            \end{equation}
            where $\prior_{\mcal{I}}$ and $\prior_{obs}$ are the isochrone and
            observational priors, respectively.

            The second term in \autoref{eq:isochrone_prior_components} --
            $\prior_{obs}(\nth{\mbs{m}} | \phi_1)$ -- encodes all observational
            effects in the photometric dataset. Some effects are easily modeled,
            e.g. instruments have well-characterized magnitude limits, both
            faint and bright. Other effects are not easily modeled, e.g, complex
            location-dependent completeness issues \citep{GaiaCompleteness}.
            Accounting for the completeness is very challenging. In practice we
            avoid the contribution of completeness to $\prior_{obs}$, masking
            data for which it is relevant. \Gaia, for instance, is complete from
            $12 < G < 17$ and has high completeness up to $G ~20$, thus we mask
            $G > 20$.  The masks are easily modeled as magnitude limits and
            replace the complex completeness model in $\prior_{obs}$.
            \autoref{sub:method:missing_data} explains how masking is
            implemented in the models, though in this case every feature is
            masked, not only a few feature dimensions. With every featured
            masked, the photometric model in
            \autoref{eq:photometric_probability} reduces from an
            $F_m$-dimensional Gaussian to a 0-D Gaussian, which is a null
            distribution and does not enter into the likelihood.  In photometric
            coordinates completeness depends on a stars apparent magnitude; vied
            in in astrometric coordinates, the completeness cannot be described
            so simply. \Gaia, for instance has scanning law residuals in the
            density field. When photometric coordinates are masked due to
            completeness, we do not mask the corresponding  astrometric
            features, so that the stars  membership probability is determined
            solely by the astrometric model. This practice allows us to include
            stars with good astrometric measurements, not penalizing those stars
            for which the photometric model is insufficient.  We examine the
            results to confirm that the scan pattern does not impact the
            membership distribution.

            The first term in \autoref{eq:isochrone_prior_components} --
            $\prior_{\mcal{I}}$ -- encodes the point-wise amplitude of the
            isochrone.  Physically, this amplitude is the the present-day mass
            function (PDMF). The PDMF $\prior_{\cal{I}}(\gamma, \phi_1)$ is a
            joint distribution, depending on $\gamma$ since it is the
            $\gamma$-dependent distribution amplitude, and also on $\phi_1$.
            This latter dependence is confusingly termed the ``mass function of
            the stream", which is the expected variation in the stellar mass
            distrubion along the stream, arising due to e.g. mass segregation in
            the progenitor \TODO{[cite]}. To date this variation along the
            stream has not been observed \TODO{[cite]}, so in practice we drop
            the $\phi_1$ dependence -- $\prior_{\cal{I}}(\gamma, \phi_1) =
            \prior_{\cal{I}}(\gamma)$ -- assuming that the PDMF is constant
            along the stream.

            % link to the code!
            In the code we offer a few ansatz for the PDMF, though any
            user-defined function may be used instead. The default assumption,
            known to be non-physical, is a uniform distribution over $\gamma$.
            Similarly non-physical are truncated uniform distributions,
            considering only a segment of the isochrone, e.g. the main sequence
            turnoff (MSTO). We include also a Kroupa initial mass function (IMF)
            \citep{2001MNRAS.322..231K}. The Kroupa IMF, and all
            physically-motivated IMFs, have an much larger fraction of low-mass
            stars then high-mass stars, the latter being less likely to form.
            This is observed in streams, particularly their progenitors. However
            the Kroupa IMF, and all IMF, are poor fits to the observed PDMF as
            stream progenitors like globular clusters are extremely old
            ($\sim10$ Gyr) while an IMF is the zero-age distribution.  How then
            should the PDMF $\prior_{\cal{I}}$ be treated?
            % ⬆ rephrase previous sentence
            One approach to the PDMF is to use a more informed distribution, for
            example time-evolving an IMF, either analytically or by simulation,
            to the age of the stream's stellar population.  As discussed
            previously, globular clusters are known to be poorly fit by
            theoretic isochrones with realistic inputs (e.g. age, $Z$). Thus
            fitting an isochrone's track to the data means its age, metallicity
            and therefore PDMF are not well fit \TODO{[cite]}. Moreover, for
            many streams, e.g. those lacking known progenitors, the PDMF is
            observationally poorly constrained.  In short, using any fixed
            distribution is simple in implementation, however the choice of
            distribution is not.

            Rather than using a fixed \textit{a priori} ansatz, another approach
            to the PDMF is to incorporate it into the model, parameterized by
            some distribution and fit simultaneously to the distance modulus and
            other $\cmp{m}{\mbs{\theta}}$.  We find this, like incorporating $a,
            Z$ into the isochrone model, to be interesting avenues by which to
            develop the model. However the resulting significant increase in
            model flexibility necessitates careful regulation, and we leave this
            to a future work.

            % TODO? expand this section.
            The last approach to the issue of the PDMF, and the one we take in
            this work, is to render it unimportant.  With fixed age,
            metallicity, and other isochrone model parameters, the primary
            purpose of the isochrone is to allow the model to determine the
            distance modulus of the stream. For this, only a portion of the
            stream need be modeled.  For streams with observable main sequence
            turn-offs, like \stream{GD-1} in \gaia, this portion of the
            isochrone has both many stars and a relatively small range in
            masses. Thus, we restrict the mass range of the isochrone such that
            we are modeling only a region of interest over which the PDMF does
            not contribute to the final result.
            % TODO? work in something to the effect of: The PDMF doesn't have an
            % enormous impact, regardless as the isochrone model is very thin
            % and stars close to the isochrone photometrically are generally
            % quite far astrometrically. This is an advantange of simultaneous
            % modeling, where errors in one model are balanced by other parts of
            % the model.

        % ssub: on_stream (end)

        \subsubsection{Linking to Astrometrics: the Distance Modulus and the Parallax} \label{ssub:method:linking_to_astrometrics}

            % TODO: rephrase
            By modeling jointly the astrometrics and photometrics we may tie
            together components of the models.  In particular, the distance
            modulus in the photometric model and the parallax in the astrometric
            model are both measures of the distance track of the stream.  We
            introduce a prior to connect these components of the model:
            \begin{align}
                \prior(\{\mu_I, \mu_\parallax \} | \phi_1) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1), \label{eq:distance_track_prior}
                % \prior(\{\sigma_I, \sigma_\parallax \}(\phi_1)) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1) \label{eq:distance_width_prior}
            \end{align}
            % TODO: should we show that these are different distance functions? or is it obvious
            where $\delta$ is the Dirac delta function
            \citep{1947pqm..book.....D}, and enforces that the distance modulus
            track and parallax track must be equal (when converted to the actual
            distance) as a function of $\phi_1$.  The widths are similarly
            constrained. The distance and first-order error conversion are given
            by
            \begin{align}
                \mu_I(\parallax) &= 10 - 5\log_{10}(\parallax) \\
                \sigma_I(\sigma_\parallax)  &\approxeq \frac{5}{\ln{10}} \mu_I \lvert\sigma_\parallax\rvert.
            \end{align}

            In practice, at the distance of many streams the errors in the
            parallaxes are of order 1 and we exclude the feature from the model.
            % TODO!

        % ssub: linking_to_astrometrics (end)

        \subsubsection{Off-stream}
        \label{sub:method:photometric_model:off_stream}

            In this section we introduce our model for the distribution of
            backgrounds in photometric coordinates.  The distribution of stars
            in color (i.e., $g-r$) and magnitude (i.e., $g$) is a complicated
            function of location in the galaxy \TODO{(e.g. [cite])}. Even within
            a field centered on the stream of interest, the color-magnitude
            diagram (CMD) is a combination of stellar populations distributed
            over a range in distances. Because of this, we do not expect an
            analytic density to provide a realistic description of the data. In
            this work we take a data-driven approach, and represent the CMD
            using a normalizing flow model.

            Normalizing flows provide a flexible description of a probability
            density, by transforming samples generated from a simple base
            distribution (i.e., a Gaussian) to the more complicated target
            distribution. The transformations are typically non-linear outputs
            of a neural neural network, whose parameters are optimized until the
            target density accurately characterizes the data. The neural network
            transformation is differentiable, so that jacobian factors can be
            combined to ensure that the the target distribution remains a valid
            probability density (i.e., integrates to one) \TODO{[cite]}.

            Our implementation is as follows. For a given stream, we restrict
            the data to some field characterized by a polygon drawn around the
            stream. Because our density modeling is aimed at stream
            characterization rather than discovery, we assume that the stream is
            sufficiently well known so that it can be masked out as a simple cut
            in (e.g.) $\phi1-\phi_2$ coordinates. With the stream masked out, we
            then fit a normalizing flow to the photometric data. Because the
            background CMD can evolve with $\phi_1$, we fit a \emph{conditional}
            normalizing flow, $q_{\rm flow}(\mathbf{m}|\phi_1)$, so that the
            color-magnitude diagram is conditionally dependent on position along
            the stream. In practice, we model magntiudes and not colors (i.e.,
            $\mathbf{m} = (g,r)$ rather than $g \ \mathrm{and} \ g-r$.
            Otherwise, the two arguments will always be covariate). Our central
            assumption with this model is that the distribution of colors and
            magnitudes for background stars in the masked region are
            quantitatively similar to those in the unmasked region. For a
            sufficiently narrow mask we expect this to be the case, and find
            that this assumption does not introduce bias for the streams
            considered in this work.

            Once the normalizing flow, $q_{\rm flow}\left(\mathbf{m} |
            \phi_1\right)$, is fit to the data, its parameters are fixed while
            the other components of the model begin training. Therefore, the
            normalizing flow can be rapidly trained and computed as a
            pre-processing step.

            % The specific architecture of the photometric normalizing flow are
            % discussed in \autoref{sub:methods:model_implementation}. But
            It is easy to evaluate the log-likelihood a normalizing flow at a
            point $\mbs{m}$ when all feature dimensions are present. When one
            (or more) feature dimension is missing, however, evaluating the
            marginal log-likelihood involves integrals over the flow. While the
            procedure is straightforward it is computationally expensive. As an
            alternative, we conservatively set the log-likelihood to $0$, the
            maximum value, overemphasizing the flow's contribution to the
            background model. Conceptually, up-weighting the background model
            decreases the false-positive rate in stream membership
            identification, at the cost of increasing the false-negative rate.
            However, the intrinsic width of stellar streams tends to be quite
            small, with correspondingly small $\sigma_\mu$ (see
            \autoref{eq:photometric_probability}); thus the effective region of
            the isochrone model is well determined and we do not find the
            marginal log-likelihood approach to be impactful. Another approach,
            which we leave to a future work, is to construct normalizing flows
            of each marginal distribution, e.g. $q_{\rm flow}\left(\{m_i\}_{i
            \in \mbs{I}_F, i \neq {j, k}} | m_j, m_k, \phi_1\right)$, and select
            for each $\nth{\mbs{m}}$ the correct flow given the present
            features.

        % ssub: off_stream (end)

    % sub: photometric_model (end)

    \subsection{Priors} \label{sub:methods:priors}

        In \autoref{sub:method:likelihood_setup} we introduced the likelihood
        setup. In \autoref{sub:method:photometric_model:on_stream} we added
        priors in the form of the isochrone's PDMF and photometric observational
        constraints. In this section we introduce and develop more priors that
        are important for training the BMDN's parameters' neural network(s).

        \subsubsection{Parameter Bounds} \label{ssub:methods:priors:parameter_bounds}

            For each parameter in $\mbs{\theta}$ of the model we implement
            priors on its range.  Uninformative, e.g. uniform $\mcal{U}(a, b)$,
            priors are often assumed where little is known about a parameter
            (except that $(a,b]$ contain the correct $\theta$ and its
            distribution). Where more is known about $\theta$, for instance that
            it is a categorical, more informative priors like the Dirichlet
            distribution \citep{Bishop_Christopher_M_2016-08-23} may be used.
            Our code permits any user-defined prior, and we include $\mcal{U}(a,
            b)$.  The default choice of prior and bounds $\mcal{U}$ set to the
            limits of the dataset in each feature dimension.
            % ⬆ where should this sentence be?

            The choice of prior is important, not only for the information it
            adds to the model, but also for how optimizers explore the parameter
            space. Simple optimizers, like MCMC \TODO{[cite]}, propose a
            parameter vector then compute its likelihood given the data. Priors
            with compact support  $\theta \sim \mcal{D}((a, b])$, e.g.
            $\mcal{U}(a, b)$, may be very dangerous as the optimizer can be
            initialized in a region with zero probability and has no means but
            blind luck to find the region of support -- the region with non-zero
            likelihood. More sophisticated optimizers, like gradient-based
            descent \TODO{[cite]}, are susceptible to a similar initialization
            problem, allowing zero-probability regions, but requiring at least a
            non-zero gradient to make informed updates to the optimizers state.
            Outside the  support the gradient can be zero.  With MDNs, model
            parameters are neural network outputs. Unlike for conventional
            models where an informed initial parameter state may be chosen by
            hand, it is impractical and undesirable to initialize by hand a
            neural network's output by setting all its internal weights. One
            might imagine accomplishing a desired initial output by pre-training
            the network with a toy model that forces convergence on a set output
            vector. However, neural networks are highly non-linear in their
            response to internal state changes (a desirable property). It is
            quite possible that when switching from the toy to real model's loss
            function the pre-trained network is in a local probability island
            from which the global optimum is unreachable.  
            Therefore, with the MDN neural networks, initialization is mostly
            for checking the dependence of the results on the initial state --
            though even this dependence may instead be done using parameter
            dropout. Rather than focusing on initializing to a particular
            parameter vector, for neural networks it is instead important that
            the initial state, whatever it is, has non-zero probability (or
            non-zero gradient, depending on the optimizer).  Neural networks
            have some intrinsic phase-space range that depends on the
            architecture of the network. Therefore the network architecture must
            be chosen such that the outputs match well with the priors. For
            example, with a uniform prior $\theta \sim \mcal{U}(a,b)$, the
            network should be restricted such that $\theta \in (a, b]$, e.g. by
            ending the network's $\theta$ channel with a scaled logistic
            function.
            \begin{equation}\label{eq:scaled_logistic}
                  \sigma(x, \{a,b\}) = a + (b - a) * \sigma(x)
            \end{equation}
            Therefore, we pair every parameter bound with a matching bound in
            the network architecture.

            % ⬇ where should this sentence be?
            The parameter bounds may be independent of $\phi_1$, as is implicit
            to the above discussion, or conditionally dependent and vary as any
            user-defined function of $\phi_1$.  Alternatively, priors parameter
            may be separated into pieces, dealing with different modeling needs,
            e.g. defining broad parameter bounds as was done in this section.
            We introduce a different piece subsequently.

        % ssub: parameter_bounds (end)

        \subsubsection{Region Prior} \label{sub:methods:priors:track_region_prior}

            In \autoref{ssub:method:astrometric_model:on_stream} the stream
            model's parameters were unconstrained. Bounds on the parameters and
            the neural network architecture were introduced in
            \autoref{ssub:methods:priors:parameter_bounds} to ensure the initial
            state of the model and network were compatible with the optimizer.
            Now we provide known information about physical systems, particular
            the track of the stream.  Our density modeling is aimed at
            characterization of \textit{known} streams, not stream discovery.
            From previous studies, the broad track of a stream is known for
            significant portions of its length \citep[e.g. see the atlas
            in][]{Mateu2022}. It makes sense, then, to provide this information
            to the model. Care must be taken that this prior information does
            not dominate the model.
            % A good balance between these two goals is to have a prior that
            % only influences the track on large scales.
            We introduce a region prior to encourage the stream model to pass
            through a user-defined control region, while also remaining
            compatible with the data.  We refer to the prior as a
            ``split"-Normal prior: a Gaussian split and separated at the peak
            and pieced back together with a flat region connecting the halves.
            The un-normalized PDF is given by:
            \begin{small}
            \begin{equation}
                \pdf(\theta,\mu,\sigma,w) \propto \begin{cases} 
                   \Exp{-\frac{1}{2}\left(\frac{x-(\mu-w)}{\sigma}\right)^2} & \phantom{\mu - w <}\ x < \mu - w \\
                    1 & \mu - w \leq x < \mu + w \\
                    \Exp{-\frac{1}{2}\left(\frac{x-(\mu+w)}{\sigma}\right)^2} & \mu + w < x
                \end{cases}.
            \end{equation}\end{small}
            The piecewise PDF is smooth up to the first derivative and can be
            used with gradient-based optimizers.

            The Gaussian portions of the prior encourage the network towards the
            central region $\mu - w < x < \mu + w$. Since the prior is
            equi-probable in the region, the stream may lie anywhere as
            preferred by the data and the stream model.  It is convenient to
            reparametrize, with $\tau = \frac{1}{\sigma}$. Namely,
            \begin{small}
            \begin{equation}
                \ln \pdf(\theta,\mu,w) \propto \tau \begin{cases} 
                    -\left(x-(\mu-w)\right)^2 & \phantom{\mu - w <}\ x < \mu - w \\
                    0 & \mu - w \leq x < \mu + w \\
                    -\left(x-(\mu+w)\right)^2 & \mu + w \leq x
                \end{cases},
            \end{equation}\end{small}
            where the additive normalization is
            $-\ln\left(\sqrt{\frac{\pi}{\tau}}
            \left(\frac{1}{2}\text{erf}\left(2 \sqrt{\tau }
            w\right)+1\right)\right)$. In this form we see that $\tau$ may be
            increased to strengthen the prior and encourage $\theta$ to lie
            within the desired region. It is important that the region be set
            large enough that we may be confident the prior is not driving the
            final result on small scales. As an added measure, it is often
            useful to turn off this prior after the stream lies within the
            desired regions.

        % ssub: region_prior (end)

    % sub: priors (end)

    \vspace{5pt}
    \subsection{Model Implementation, aka the code} \label{sub:methods:model_implementation}

        In prior sections we discussed the mathematical implementation of each
        component of the Mixture Density Network. Here we make brief remarks on
        the code implementation and architecture of the neural networks.
        % Rephrase: I don't like this topic sentence.

        The neural networks are built with the \package{PyTorch} stack
        \TODO{[cite]}. We wrap the network \texttt{pytorch.Module} in a custom
        framework, allowing the Bayesian likelihood, prior, evidence, etc. to be
        bundled with the network's forward step. This framework allows for a
        high degree of modularity. Component models, e.g. the stream model, of
        the mixture are easily added to (and moved from) as items in the larger
        mixture model. Probability methods understand how to traverse the
        mixture model, calling and combining the appropriate methods of the
        constituent models, so it unnecessary to write custom loss functions.
        Another benefit of the modular framework is the ease of building and
        experimenting with different mixture models. Furthermore, each model
        component may be saved and serialized separately from the other
        components, allowing for portability of the components between models.
        For example, if a stream has a prominent extra-tidal feature, e.g. the
        spur of \stream{GD-1}, a fiducial model of just the thin stream may be
        trained, then the stream (and background) model components may be saved
        and used as the initialization to the stream (and background) of a more
        complete model that includes the spur. This modularity also has a cost:
        by allowing each model component (including the categorical mixture
        model itself) to contain its own network we decrease the cross-talk
        between nodes in separated networks. The networks communicate only
        through the backwards propagation of the loss function, increasing the
        required training time of the models. We note that the code framework,
        being built on \package{PyTorch}, easily permits a single model with
        single network. Having found a good model by experimenting with the
        modular layout, the model may be reproduced and re-trained as a single
        network. Alternatively, though we leave this to a future version of the
        code, it is possible to ``fuse" the mixture, providing one network whose
        outputs are apportioned appropriately to each constituent model.

        The models permit user-defined priors to be passed at model
        instantiation.  Priors specific to a single model are contained on that
        model, while priors connecting two (or more) models are held on the
        mixture model object. As an example of the latter, see
        \autoref{ssub:method:linking_to_astrometrics}, which connects the
        distance component of the astrometric and photometric stream models.

        The custom model framework combines for each model a neural network with
        Bayesian probability functions. The latter are set by the model type,
        e.g. the Gaussian model has a Gaussian likelihood function, but the
        former, the neural network, is set and determined by the user. So long
        as the network has the correct number of inputs (e.g. one for $\phi_1$)
        and outputs (one per parameter), any feed-forward network may be used.
        Our default choice, and for which we provide a helper function, is an
        MLP with a sequence of \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
        \!\!\textrightarrow \texttt{Dropout} blocks. This is illustrated in
        \autoref{table:mlp_structure}.
        % FIXME! reference not being picked up correctly.
        Dropout layers are used to prevent network co-adaptation and
        over-fitting. Dropout works by randomly zeroing out nodes in the
        network, decreasing the importance of any single node, encouraging
        instead emergent patterns across the network.  One model for which the
        lin-tanh-dropout network cannot be used is the normalizing flow used in
        the photometric background model
        (\autoref{sub:method:photometric_model:off_stream}). Here we use the
        \package{nflows} package, with a conditional diagonal normal base
        distribution and composite transform network of a reverse permutation
        and masked affine auto-regressive transform layers. As noted in our data
        availability statement, all the codes used in this paper are bundled by
        \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}}
        in a public repository at $\GitHubURL{}$. We will describe for each
        stream model specifics relevant to that stream but refer readers to the
        publicly-available code itself for code implementation details.
        % TODO: Maybe more details in the Appendix?

        \begin{table}
            \centering
            \caption{% 
            Example neural network structure consisting of 1 input block mapping
            a $\phi_1$ scalar to the $H$ nodes of the hidden blocks, the number
            of which determines the depth of the network. The hidden blocks map
            to the $F$ outputs in the output block. Each block consists of a
            \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
            \!\!\textrightarrow \texttt{Dropout} sequence. Vector $\phi_1$
            inputs are supported as an element-wise application of the network.
            }
            \label{table:mlp_structure}
            \begin{tabular}{@{}rrcl@{}}
            \toprule
            & Layer & Shape & Parameters \\
            \midrule
            & Linear & (1, H) & 2H  \\
            & Tanh & (H, H) & 0  \\
            & Dropout & (H, H) & 0  \\
            \hline
            & Linear & (H, H) & $2H^2$  \\
            $N_{\rm hidden} \times$ & Tanh & (H, H) & 0  \\
            & Dropout & (H, H) & 0  \\
            \hline
            & Linear & (H, F) & $2HF$  \\
            & Tanh & (F, F) & 0 \\
            & Dropout & (F, F) & 0  \\
            \bottomrule
            \end{tabular}
        \end{table}

% section method (end)

%\newpage
\section{Results: Mock} \label{sec:results_mock}

    In this section we present an application of our model to a synthetic stream
    with known ground truths, and demonstrate the model's ability to
    characterize stream density variations and membership probabilities in an
    unsupervised manner.

    In this section we apply our model to a synethic stream in order to
    demonstrate an accurate recovery of stream members and density variations.

    We generate synthetic observations by sampling points from a third degree
    polynomial \JN{check this} with a constant scatter. Density variations are
    then simulated by evaluating the stream density with two Gaussian
    distributions centered at different values of $\phi_1$ with random
    amplitudes (which sum to $1$) and widths. Stream stars are then removed
    using random sampling, weighted by their likelihood of belonging to the
    density. The background astrometric  points in are generated from
    exponential distributions at each $\phi_1$. The number of stream stars is
    $\variable{output/mock/nstream_variable.txt}$, while the number of
    background stars is $\variable{output/mock/nbackground_variable.txt}$.

    \begin{figure}
        \script{mock/plot/photometric_background_selection.py}
        \centering
        \includegraphics[width=1\linewidth]{figures/mock/photometric_background_selection.pdf}
        \caption{
            % 
            \textbf{Top panel:} Astrometric coordinates ($\phi_1$, $\phi_2$) for
            the mock data set.  Light blue points are the stream and background
            within an ``on"-stream region. Black points are in the ``off"-stream
            region used to train the background photometric model.
            % 
            \textbf{Bottom panel:} Astrometric coordinates ($\phi_1$, $\phi_2$)
            for the mock data set.  Black points are in the ``off"-stream region
            from the region defined in the astrometric coordinates.  Light blue
            points are the stream and background from the on-stream region. The
            stream's isochrone is apparent as an overdensity around (g-r,
            g)$\approx (0, 20)$.  }
        \label{fig:mock_data_photometric_background_selection}
    \end{figure}

    The synthetic stream is also simulated in magnitudes using a 12~Gyr
    isochrone with [Fe/H] = -1.35~dex. The isochrone truncates just short of the
    giant branch, and sampled from uniformly (i.e., the initial mass function,
    $P(\gamma)$, is uniform). The background photometric coordinates are
    generated as a 2-dimensional Gaussian with non-zero covariance. The distance
    modulus is derived from a polynomial parallax track with uniform thickness.
    Thus, the synthetic data scenario described here generates a mock stream in
    astrometric coordinates and photometric coordinates, with density
    variations, a distance gradient which is consistent in parallax and distance
    modulus (i.e., magnitudes), all superimposed over a noisy field of
    background points in each astrometric and photometric dimension.

    \begin{figure*}
        \script{mock/plot/results.py}
        \centering
        \includegraphics[width=1\linewidth]{figures/mock/results.pdf}
        \caption{
            %
            \textbf{Top panel:}%
                The stream mixture coefficient $f(\phi_1)$ (black dotted line)
                predicted by the model, overplotted on the true stream fraction
                (black-outlined histogram), calculated as the ratio of stream
                stars to total stars in 55 bins over the $\phi_1$ range of the
                isochrone.  \textbf{Middle panel: }%
                $\phi_2(\phi_1)$ over the full range of $\phi_1$. The data color
                and transparency is set by the model's MLE stream probability
                (with dropout turned off). Plotted behind the data are the true
                labels for the stream stars (large black circles).  The true
                labels exactly limn the model-identified stream stars.  The MLE
                stream track (black dashed line) is over-plotted on the
                probability-colored data.
            %
            \textbf{Lower panel 3: }%
                $\varpi_2(\phi_1)$ over the full range of $\phi_1$. Like for
                $\phi_2$, the data color and transparency is set by the MLE
                probability, with the true labels (large black circles)
                under-plotted and MLE track (black dashed line) over-plotted.
            %
            \textbf{Upper Histograms: }%
                $\phi_2$ histograms for each model component (blue background,
                red stream), in 4 $\phi_1$ ranges across the data set.  The
                height of each bin is the probability-weighted density.  The
                background is a uniform distribution, while the stream is a
                Gaussian. In projection across the $\phi_1$ range it can appear
                multi-modal (panel 2) if there are large gaps or `smeared' out
                if the Gaussian's mean changes significantly over the range.
            %
            \textbf{Upper Middle Histograms: }%
                Parallax $\varpi$ histograms for each model component (blue
                background, red stream), in 4 $\phi_1$ ranges across the data
                set.  The height of each bin is the probability-weighted
                density.
            %
            \textbf{Lower Middle and Lower Histograms:}%
                Photometric coordinate $(g, r)$ plots for each model component
                in 4 $\phi_1$ ranges across the data set. The top row plots the
                stream stars (red) over the background. The lower row plots the
                background stars (blue) over the stream stars to show that the
                how the two distributions mix.  } TODO:
        \begin{enumerate}
            \item the linear density in the stream-fraction plot
        \end{enumerate}
        \label{fig:mock_data_result}
    \end{figure*}

    The result of applying our density model to the synthetic data is
    illustrated in \autoref{fig:mock_data_result}. The top panel shows the
    weight parameter that controls membership probability as a function of
    $\phi_1$. The dashed curve represents the output of our data-driven model,
    while the solid curve is the ground truth. The second panel shows the stream
    and background in the $(\phi_1,\phi_2)$ plane, and the third panel shows the
    stream and background in the $(\phi_1,\parallax)$ plane. Points are
    color-coded according to their membership probability
    \autoref{eq:membership_prob}. The weight parameter clearly shows that the
    model has captured the two prominent density variations along the stream.
    The are additional variations  mostly around the edges of the stream, though
    re-sampling the neural network weights with dropout reveals that these
    variations represent regions of higher model uncertainty.

    We find that the stream is successfully recovered in both astrometric and
    photometric coordinates, with $x\%$ stream stars being recovered with
    membership probability greater than $y$. The false positive rate (i.e.,
    stars misidentified as stream members) for this test is found to be $z\%$
    when cutting on this threshold.

    In the bottom two rows of \autoref{fig:mock_data_result} we illustrate the
    synthetic data color-coded by stream membership probability (top) and
    background membership probability (background) for different $\phi_1$
    slices. Importantly, the photometric portion of the mixture model
    successfully identifies the stream in magnitude space, showing a high stream
    membership probability where the stream is present, and a low probability
    ($\sim 0$) where the stream is absent.

% sec: test_on_a_synthetic_stream (end)

\section{Results: GD-1} \label{sec:results_gd1}

    We now apply our density modeling approach to the stellar stream GD-1. We
    discuss the data selection in \autoref{sub:results_gd1_data} and results in
    \autoref{ssub:trained_gd1}.

    \subsection{Data Selection} \label{sub:results_gd1_data}

        \begin{enumerate}
            \item Gaia data.
                \begin{enumerate}
                    \item box size
                    \item selection criteria
                \end{enumerate}
            \item crossmatched with panstarrs1
            \item extinction corrected with bayestar2019
        \end{enumerate}

        We utilize data from {\it Gaia} DR3 for the astrometric coordinates of
        each star, and Pan-STARRS1 to obtain more accurate color information. We
        adopt broad cuts on the cross-matched data to limit the field to a
        region known to contain GD-1. We perform these rough cuts by requiring
        \JN{fill in what the cuts are and reference proper motion figure and CMD
        figure.}

        The result is a field populated by \JN{x} stars, from which we would
        like to identify stream members.

        \todo[inline]{Dust}
            We use \package{dustmaps} Bayestar 19 for the dust modeling. We treat it as known. Include as a source of error in $\Sigma_n$
            % TODO: actually implement that in the model

    \subsection{Model Specification}\label{sec:GD_1_ModelSpecification}

        We model GD-1 as a three-component mixture consisting of a main stream
        component, a spur, and background. In order to guide our model towards
        GD-1, we set down 4 ``control points" along GD-1's main track, and 3
        control points along the spur (illustrated in \JN{x}). These control
        points act as Gaussian priors on the model. We set the Gaussian error on
        these control points to large values, so that their influence on the
        model is not dominant. Indeed, we find that the value of the control
        points is to help reduce training time, since otherwise it will take the
        flexible density model longer time to converge to the actual stream.

        The weight parameter is also set to truncate to $0$ at $\phi_1 < X$ and $\phi_1 > Y$. This ensures that we do not model stream stars that fall

    \subsection{Trained Density Model}\label{ssub:trained_gd1}

        In this section we present results from our model applied to real {\it
        Gaia} and Pan-STARRS data of GD-1.

        We visualize the performance of our model in identifying stream stars
        and characterizing the astrometric background with
        \autoref{fig:gd1-results-full} and \autoref{fig:gd1-results-panels}. In
        \autoref{fig:gd1-results-full} The top panel shows the stream in
        $\phi_1$, $\phi_2$ coordinates, color-coded by membership probability
        (derived from \autoref{eq:membership_prob}). The following three rows
        represent background distributions in $\phi_2, \mu_{\phi_1}, \rm{\ and}
        \ \mu_{\rm phi_2}$, respectively taken at different $\phi_1$ increments
        centered on the red lines. The $\phi_1$ increments are roughly $\sim X$
        in size.

        The blue lines over-plotted represent the best fit model at the given
        $\phi_1$ slice. We find that the backgrounds are typically well
        characterized by the distributions adopted (described in \S-x). We can
        also see from \autoref{fig:gd1-results-panels} that the stream and spur
        components appear to be well characterized.

        Our ability to recover the stream is demonstrated further in
        \autoref{fig:gd1-results-full}. In this figure, we only plot stars if
        their membership probability is greater than \JN{X}. The result is a
        ``gold" catalog of probable stream and spur members, extending over
        $X$-degrees in $\phi_1$. We find the expected gap around $\phi_1 \sim
        -20~\rm{deg}$, which has been proposed to be the location of the
        dissolved progenitor (CITE).

        Our model also fits the distribution of stream stars in the CMD. These
        fits are illustrated in \autoref{fig:gd1-results-panels}, where the CMD
        for the population in a given $\phi_1$ band is plotted, color-coded by
        stream membership probability. For the on-stream regions, the isochrone
        appears to provide a decent description of the stream, though there is
        room for further improvements (discussed in \S-x). However, we note that
        the stream's presence and absence in the isochrone is well captured by
        the mixture weight, which is a neural network output.

        Additionally, the distance modulus is also a neural network output,
        conditioned on $\phi_1$. This allows us to obtain a distance modulus at
        each $\phi_1$ along the stream. The distance modulus track and
        (converted) distance track are illustrated in
        \autoref{fig:gd1-results-full}. The left panel shows the distance
        modulus (in magnitudes), while the right shows the corresponding
        distance track. Importantly, the prior on the distance track is that the
        stream must fall between $[4,40]~\rm{kpc}$. Thus, our ability to recover
        the distance to GD-1 without substantial bias provides a useful
        indicator that our color-magnitude model combined with the astrometric
        model provide an adequate description of the data. \JN{more on this...}

        \begin{figure*}[h]
            \script{gd1/plot/data_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/gd1/data_selection.pdf}
            \caption{Caption}
            \label{fig:gd1-data_selection}
        \end{figure*}

        \begin{figure}
            \script{gd1/plot/photometric_background_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/gd1/photometric_background_selection.pdf}
            \caption{Caption}
            \label{fig:gd1-photometric_background_selection}
        \end{figure}

        \begin{figure*}[h]
            \script{gd1/plot/results_full.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/gd1/results_full.pdf}
            \caption{CAPTION}
            TODO:
            \begin{itemize}
                \item Add plot of the stream linear density
                \item Color palette for the spur
                \item more subtle stream MLE path
                \item rerun model with larger pmphi2 sigma
                \item add the isochrones to the cmd plots
            \end{itemize}
            \label{fig:gd1-results-full}
        \end{figure*}

        \begin{figure*}[h]
            \script{gd1/plot/results_panels.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/gd1/results_panels.pdf}
            \caption{CAPTION}
            TODO:
            \begin{itemize}
                \item Add plot of the stream linear density
                \item Color palette for the spur
                \item add the isochrones to the cmd plots
            \end{itemize}
            \label{fig:gd1-results-panels}
        \end{figure*}

        \begin{figure*}
            \script{gd1/plot/smooth_likelihood.py}
            \centering
            \includegraphics[scale=0.8]{figures/gd1/smooth_likelihood.pdf}
            \caption{Caption}
            \label{fig:gd1-heatmap}
        \end{figure*}

        \variable{output/gd1/member_table_select.tex}

% sub: gd1 (end)

\section{Palomar 5} \label{sec:results_pal5}

    Content

% section pal5 (end)

% section results (end)

\section{Discussion}

    \subsection{Comparison to Other Methods} \label{sec:comparison}

        \subsubsection[STREAMFINDER]{\cite{2018MNRAS.477.4063M}}

            Assumes a potential

        % sub: streamfinder (end)

        \subsubsection{VIA MACHINAE}{\cite{2022MNRAS.509.5992S}}

            Only astrometric.

        % sub: via_machinae (end)

        \subsubsection[Uniform Modeling of 13 Stellar Streams]{\cite{PatrickEtAl2022}}

            Doesn't do simultaneous modeling.

        % sub: uniform_modeling_of_13_stellar_streams (end)

\section{Conclusions} \label{sec:conclusions}

% section comparison_to_other_methods (end)

\section{acknowledgements}

    We would like to thank the Academy.

    The data availability statement is modified from one provided to ShowYourWork by Mathieu Renzo.

    \paragraph{Software (alphabetical)}

        \package{asdf} \citep{Greenfield2015},
        \package{astropy} \citep{Astropy2013, Astropy2018, Astropy2022},
        \package{astroquery} \citep{Astroquery2019},
        \package{Matplotlib} \citep{Hunter2007},
        \package{NumPy} \citep{Harris2020},
        \package{Pytorch} \citep{Pytorch2019},
        \package{SciPy} \citep{Scipy2020},
        \package{ShowYourWork} \citep{}

\section*{Data Availability} \label{sec:data_availability}

    This study was carried out using the reproducibility software
    \href{https://github.com/showyourwork/showyourwork}{\showyourwork}
    \citep{Luger2021}, which uses continuous integration to
    programmatically download the data, perform the analyses, create the
    figures, and compile the manuscript. Each figure caption contains two links:
    one to the dataset used in the corresponding figure, and the other to the
    script used to make the figure. The datasets are stored at
    \todo[inline]{zenodo url}. The git repository associated
    with this study is publicly available at $\GitHubURL{}$.

% section data_availability (end)

\bibliography{bib}

\appendix

\section{Notation} \label{app:notation}

    In this work we use multiple data sets and a variety of mathematical tools
    to analyze the data, including machine learning algorithms and statistical
    models.  For clarity and to connect related components of the models used in
    this work, a consistent mathematical notation is crucial. Here we establish
    the notation and symbols used throughout this work.
    \autoref{table:symbol_table} enumerates the symbols used for the statistical
    models and the different datasets, e.g. distinguishing astrometric from
    photometric data.

    \TODO{position the table here}

    \begin{table}
    \centering
    \caption{Symbol Table}
    \label{table:symbol_table}
    \begin{tabular}{p{0.35\linewidth} p{0.6\linewidth}}
    Symbol                         & Explanation \\
    \toprule
    % -----------------------------------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Datasets}:} \\
    \midrule
    % -------------------------------------------------
    \multicolumn{2}{l}{\textbf{Full Data}:}  \vspace{5pt} \\
    $\mbf{X} \in \mbb{R}^{(N, F+1)}$ & The full dataset, including all measurements (rows) and features (columns).  \\
    $\mbs{\delta}\mbf{X} \in \mbb{R}^{(N, F+1)}$ & The full set of uncertainties in $\mbf{X}$, e.g. Gaussian $\sigma_{n,f}$ and correlation terms $\sigma_{n,(f_i, f_j)}$.\\
    $N$ & The number of independent measurements (aka stars) in the dataset.  \\
    $F$ & The number of measured features, e.g. $(\alpha, \delta, \parallax, \mu_{\alpha^*}, \mu_\delta, v_r, g, r)$,  not including the the `independent' feature over which the others are conditioned. \\
    $\mrm{n} \in \mbs{I_N} \equiv \{1, ..., N\} \in \mbb{Z^+}$ & Index into the set of measurements $\mbf{X}$ (and $\mbs{\delta}\mbf{X}$), i.e. the $n$-th star. \\
    $\mrm{f} \in \mbs{I_F}  \equiv \{1, ..., F\} \in \mbb{Z^+}$ & Index into the set of features of $\mbf{X}$  (and $\mbs{\delta}\mbf{X}$). $ f=0$ is the `independent' feature.\\
    $(\delta)x_{\mrm{n,f}}$ & $n$-th star with $f$-th feature in $(\mbs{\delta})\mbf{X}$. \\
    $(\mbs{\delta})\nth{\mbs{x}}$ & $n$-th star as a row vector of all features of $(\mbs{\delta})\mbf{X}$. \\

    % -------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Astrometric Data}:}  \vspace{5pt} \\
    $\mbf{W} \in \mathbb{R}^{(N, F_w + 1)}$ & Full astrometric dataset. \\
    $\mrm{f_w} \in \mbs{I_{F_w}} \equiv \{1, ..., F_w\} \in \mathbb{Z^+}$ & Index into the set of astrometric features. $f_w = f = 0$ is the `independent' feature.\\
    $w_{\mrm{n,f_w}}$ & $f_w$-th astrometric feature of the $n$-th star in $\mbf{W}.$\\
    $\nth{\mbs{w}}$ & $n$-th star as a row-vector of all astrometric features. \\
    $\phi_1 \equiv w_{\cdot,0}$ & The `independent' (index 0) feature column on which the models are conditioned. For convenience $\phi_1$ is used instead of $\nth{\phi_1}$ (or $w_{n,0}$) except where the $n$ subscript cannot be inferred from context.\\
    % -------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Photometric Data}:}  \vspace{5pt} \\
    $\mbf{M} \in \mathbb{R}^{(N, F_m)}$ & Full photometric dataset. \\
    $\mrm{f_m} \in \mbs{I_{F_m}} \equiv \{1, ..., F_m\} \in \mathbb{Z^+}$ & Index into the set of photometric features. \\
    $m_{\mrm{n,f_m}}$ & $n$-th measurement of the $f_m$-th feature of $\mbf{M}$. \\
    $\nth{\mbs{m}}$ & $n$-th row-vector of all features. \\
    % -----------------------------------------------------------------------------
    \\
    \multicolumn{2}{l}{\textbf{Statistics}:} \\
    \midrule
    $\pdf(...)$ & The PDF of the distribution. \\
    $\cdf(...)$ & The CDF corresponding to $\pdf(...)$. \\
    $\mbs{\theta} = \{\theta_f\}_f$ & The vector of a distribution's parameters. \\
    $\mcal{L}(\nth{\mbf{x}}, \mbs{\theta}(\phi_1))$ & The likelihood of the data given the parameters. \\
    $Q$ & The number of models in a mixture model. \\
    $\mrm{q} \in \mbs{I_Q} \equiv \{1, ..., Q \} \in \mbb{Z^+}$ & Index into the set of $Q$ models. \\
    $\mrm{b} \in \mbs{I_Q}$ & The index of the background model in the set of $Q$ models.
    % -------------------------------------------------
    \\
    \end{tabular}
    \end{table}

    We use $\tilde{F}$ to indicate that there a missing features, e.g. radial velocities.
    The index set $\mbs{I}_F \mapsto \mbs{I}_{\tilde{F}} = \{1, ..., k-1, k+1, ... \}$ where $k$ is the index of the missing feature.

% section notation (end)

% \section{Theory of Mixture Density Networks (MDNs)} \label{app:theory_of_mixture_density_networks}

%     Mixture models (MMs) are a statistical method to represent a population as a set of a sub-populations specified by probability distributions. These distributions are characterized by their parameters and mixture by a corresponding set of mixture weights. Mixture Density Networks (MDNs) 
%     are a machine learning method to combine a MM with neural networks, increasing the flexibility of the model. An MDN replaces the scalar distribution parameters of the MM with neural networks, allowing the distribution parameters and mixture weights to vary as smooth functions \citep{Bishop1994}. Functional parameters allows the model to capture variation over those parameters that is not possible with scalar-valued models. Moreover, by using neural networks the inferred functional forms are ``model-free" and driven entirely by the data.

%     % Quote from https://mikedusenberry.com/mixture-density-networks
%     % A mixture density network (MDN) is an interesting model formalism built within the general framework of neural networks and probability theory for working on supervised learning problems in which the target variable cannot be easily approximated by a single standard probability distribution. Essentially an MDN allows one to model a conditional probability distribution as a mixture of distributions, in which the individual distributions and the corresponding mixture coefficients are parameterized by functions of the inputs

%     For the MDN we take $x_{\cdot, 0} \equiv x_{\cdot}$ (indexed from 0), which is the first feature column of $\mbf{X}$, as the `independent' coordinate over which the MDN parameters are conditioned. In the context of streams $x_{\cdot, 0} \equiv \phi_1$ is the longitude coordinate in a reference frame rotated such that the stream is aligned along $\phi_1$.

%     The PDF of the MDN is

%     \begin{equation} \label{eq:general_mixture_network}
%         \pdf(\nth{\mbs{x}} | \mbs{\theta}(\phi_1))
%         = \sum_{\mrm{q} \in \mbs{I_Q}} \ \qth{f}(\phi_1) \qth{P}(\nth{x}|\qth{\mbs{\theta}}(\phi_1),q\!=\!\mrm{q})
%     \end{equation}
%     The mixture coefficients $\qth{f}$ are normalized s.t.
%     \begin{equation}
%         \sum_{\mrm{q}} \qth{f}(\phi_1) = 1 \qquad \forall \phi_1
%     \end{equation}
%     In practice this is enforced by defining
%     \begin{equation}
%         f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash b} \qth{f}(\phi_1)
%     \end{equation}
%     for some ``background" index $b \in \mbs{I_Q}$.

%     Assuming the measurement of each datum is independent, the likelihood is a sum over all data:
%     \begin{equation}
%         \mcal{L}\left(\mbf{x}_n | \mbs{\theta}(\phi_1)\right) = \sum_{q\in\mbs{I_Q}} \qth{P}\left(\nth{\mbs{x}} | \qth{\mbs{\theta}}(\phi_1)\right).
%     \end{equation}

%     The full posterior distribution includes the priors.

%     \begin{equation}
%         \ln\posterior\left(\mbf{X} | \mbs{\theta}(\mbs{x})\right) = \ln\mcal{L}\left(\mbf{X} | \mbs{\theta}(\mbs{x})\right) +  \sum_{n \in \mbs{I_N}} \ln \left[ \sum_{q\in\mbs{I_Q}} \qth{\prior}\left(\qth{\mbs{\theta}}(\mbf{x}_n)\right) \right].
%     \end{equation}

\section{Probability Distributions} \label{app:distributions}

    A variety of probability distributions are used when modeling the stream and Galactic background. This appendix presents relevant mathematical details of these distributions.
    In particular, observational errors are modeled as Gaussian distributions, which requires modifying each distribution to account for the Gaussian noise.

    \vspace{10pt}

    Let $x \sim X$ be distributed as some specified distribution
    and $\delta \sim \Delta \equiv \mcal{N}(0, \sigma_*)$ as an independent, centered Gaussian.
    An observation $Y$ is distributed as
    \begin{equation}
        Z = X + \Delta,
    \end{equation}
    Which is a convolution of the PDFs.
    \begin{align}  \label{eq:general_convolution}
        \pdf_Z(x; \mbs{\theta}, \sigma)
            & \equiv (\pdf_X(x; \mbs{\theta}) * \mcal{N}(x; 0, \sigma)) \\
            &= \int_{-\infty}^{\infty} \rm{d}\tau \, \pdf(x; \mbs{\theta})\mcal{N}(x-\tau; 0, \sigma)
    \end{align}

    In the following subsections we work through the distributions $\pdf(x,\mbs{\theta})$ and the convolutions with $\mcal{N}(x,0,\sigobs)$.

    \vspace{5pt}
    \subsection{Flat Distribution} \label{sub:flat_distribution}

        The simplest distribution is that of the univariate uniform distribution.
        For a domain $x \in [a, b]$ the PDF is given by
        \begin{equation}\label{eq:pdf_flat_univariate}
            P_{\mcal{U}}(x|a,b) = \begin{cases}
                (b-a)^{-1} & a < x \leq b \\
                0 & \text{else}
            \end{cases}
        \end{equation}
        where $a,b$ are the bounds.

        In practice, the bounds $a,b$ are the bounds of the observation window,
        creating a subset of a larger field.  Suppose the larger field is
        uniformly distributed in a region larger than the bounds $a, b$, ie.
        $P_{\mcal{U}}(x|\alpha,\beta)$.  Let each observed datum $x_n$ in the
        field have Gaussian error $\sigobs_n$, then the PDF of the
        uniform-Gaussian-convolved distribution is given by:

        \begin{equation}
            \pdf_{(\mcal{U}*\mcal{N})}(x; \alpha, \beta, \sigobs) = \begin{cases}
                \pdf_\mcal{U}(x; \alpha, \beta) \left( \cdf_{\mcal{N}}(x; \alpha, \sigobs) - \cdf_{\mcal{N}}(x; \beta, \sigobs) \right) & \alpha \leq x \leq \beta \\
                0 & \text{else}
            \end{cases}
        \end{equation}

        We then truncate this distribution to the observed field $(a, b]$
        \begin{equation} \label{eq:pdf_flat_univariate_convolved_error_full}
            \pdf_{(\mcal{U}*\mcal{N})}(x | a < X \leq b; \alpha, \beta, \sigobs)
        \end{equation}
        in the normal fashion: setting $\pdf_{(\mcal{U}*\mcal{N})} = 0$ when $x
        < a, x > b$ and normalizing $\pdf_{(\mcal{U}*\mcal{N})}$ within $(a,b]$.
        The truncated distribution is tractable but unwieldy. The distribution
        can be simplified enormously if two conditions hold: first, $\alpha \ll
        a, b \ll \beta$, the observed field is a small subset of the full field;
        and second, $\sigobs \ll |\alpha - a|$ and $\sigobs \ll |\beta - b|$,
        the errors are smaller than the observation window size compared to the
        full field.  In this case, the distribution reduces back to the uniform
        distribution!

        Therefore, for all cases considered in this work
        \eqref{eq:pdf_flat_univariate_convolved_error_full} becomes:

        \begin{equation}\label{eq:pdf_flat_univariate_convolved_error}
            \pdf_{(\mcal{U}*\mcal{N})}(x | a < X \leq b; \alpha, \beta, \sigobs) \cong P_{\mcal{U}}(x|a,b)
        \end{equation}

    \vspace{10pt}
    \subsection{Truncated Exponential Distribution} \label{sub:exponential_distribution}

        The univariate exponential distribution is:
        \begin{equation} \label{eq:pdf_exp_univariate}
            \pdf_{\mcal{E}}(x; \lambda) = \begin{cases}
                m \Exp{-\lambda x} & x \in [0, \infty) \\
                0 & x < 0
            \end{cases}
        \end{equation}

        The support of this distribution, $[0, \infty)$, rarely matches the
        support of the data. We generalize to the unit-normalized, truncated,
        univariate exponential distribution in the domain $(a, b]$:
        \begin{equation} \label{eq:pdf_truncexp_univariate}
            \pdf_{\mcal{E}_T}(x; \lambda, a, b) = \begin{cases}
                \frac{\lambda \Exp{-\lambda \, (x - a))}}{1 - \Exp{-\lambda(b - a)}} & a < x \leq b \\
                0 & \text{else},
            \end{cases}
        \end{equation}
        for bounds $x \in (a,b]$. However, as $\lambda \rightarrow 0$, which
        describes the flat distribution (\autoref{sub:flat_distribution}), this
        functional form can be numerically unstable. It is therefore practical
        to define $\pdf_{\mcal{E}}(x; |\lambda| < \epsilon, a, b) \cong
        P_{\mcal{U}}(x; a,b)$.

        Let each observed datum $x_n$ in the field have Gaussian error
        $\sigobs_n$, then the PDF of the Exponential-convolved-with-a-Gaussian
        distribution is given by:

        \begin{equation}
            \pdf_{\mcal{E}\cdot\mcal{N}}(x;m,a,\sigma_*) = \frac{m  e^{\frac{1}{2} \lambda  \left(2 b+\lambda  \sigma _*^2-2 x\right)}}{e^{\lambda  (b-a)}-1} = \pdf_{\mcal{E}}(x;\lambda,a) \ \Exp{\frac{1}{2}\lambda^2 \sigma_*^2},
        \end{equation}
        where the function is no longer truncated, since a Gaussian is not
        compact.

        We impose the same observation field $(a, b]$ and truncate
        $\pdf_{\mcal{E}\cdot\mcal{N}}(x | a < X \leq b;m,a,\sigma_*)$.  We note
        the normalization, which is the CDF evaluated at $a, b$, cancels the
        factor $\Exp{\frac{1}{2}\lambda^2 \sigma_*^2}$. Following thtough, the
        truncated convolved distribution is identical to the original
        distribution.

        \begin{equation}
            \pdf_{(\mcal{E}\cdot\mcal{N})_T}x | a < X \leq b;m,a,\sigma_*) \equiv \pdf_{\mcal{E}}(x; \lambda, a, b)
        \end{equation}

    \vspace{10pt}
    \subsection{Normal Distribution} \label{app:sub:normal_distribution}

        The properties of the Normal distribution are well known. For
        completeness we state

        \begin{equation}
            P_X(x; \mu, \sigma) = \mcal{N}(\mu, \sigma).
        \end{equation}

        With the inclusion of observational errors the distribution becomes

        \begin{equation}
            P_{\mcal{N}\cdot\mcal{N}}(x; \mu, \sigma, \sigobs) = \mcal{N}(\mu, \sqrt{\sigma^2 + \sigobs^2}).
        \end{equation}

        The Normal distribution is not compact, with non-zero value over all
        $x$. In an observational window $(a,b]$ the PDF is truncated, giving

        \begin{equation}
            P_{\mcal{N}_T}(x; \mu, \sigma, a, b) = \begin{cases}
                \frac{\mcal{N}(x, \mu, \sigma)}{\Phi_\mcal{N}(b; \mu, \sigma) - \Phi_\mcal{N}(a; \mu, \sigma)} & a < x \leq b \\
                0 & \text{else}
            \end{cases}
        \end{equation}

        Note that the observation-window-truncation is performed on the full
        distribution, which includes the observational errors, if present.
        Therefore, let $\tilde{\sigma}^2 \equiv \sigma^2+\sigma_*^2$

        \begin{equation}
            P_{(\mcal{N}\cdot\mcal{N})_T}(x; \mu, \tilde{\sigma}, a, b) = P_{\mcal{N}_T}(x; \mu, \tilde{\sigma}, a, b)
        \end{equation}

    \vspace{10pt}
    \subsection{Skew-Normal Distribution}

        The PDF of the skew-normal distribution is given by

        \begin{equation}
            \pdf_{SN}(x;\mu,\sigma,\alpha) = 2\pdf_{\mcal{N}}(x;\mu,\sigma)\cdf_{\mcal{N}}(\alpha x;\mu,\sigma).
        \end{equation}

        Despite the popularity of this distribution we are not aware of any work
        showing the convolution of the skew-normal distribution with a centered
        Gaussian. Moreover, our Mathematica (v13.2) does not successfully
        perform this convolution. Consequently we present a step-by-step
        derivation of $P_{X\cdot\mcal{N}}$.

        \autoref{eq:general_convolution} establishes that convolutions of PDFs
        are an integral operation. Integrals are often intractable, but the
        problem is solvable if recast in different terms. Rather that working
        with PDFs of distributions we can instead consider the moment-generating
        function (MGF) corresponding to the PDF's distribution. Like the PDF,
        the MGF uniquely determines a distribution, so a MFG uniquely maps to a
        PDF. For the MGF of two distributions the convolutions is a
        multiplication, not an integral.  Therefore, for the convolution of two
        PDFs their MGFs are multiplied and if the product is the MGF of a known
        distribution, the corresponding PDF is the convolution of the two
        original PDFs.

        The MGF of the Normal distribution is

        \begin{equation} \label{eq:mgf_normal}
            M_{\mcal{N}}(t) = \Exp{\frac{\sigma_*^2 t^2}{2}}
        \end{equation}

        And the MGF of the skew-normal is

        \begin{equation} \label{eq:mgf_skewnormal}
            M_{S\mcal{N}}(t) = 2 * \Exp{\mu t + \frac{\sigma^2 t^2}{2}} \cdf(\sigma \delta t), \quad \delta = \frac{\alpha}{\sqrt{1 + \alpha^2}}.
        \end{equation}

        The product of the two MGF is

        \begin{equation}
            M_{S\mcal{N}*\mcal{N}}(t) = M_{S\mcal{N}}(t) * M_{\mcal{N}}(t) = 2 * \Exp{\mu t + \frac{(\sigma^2 + \sigma_*^2) t^2}{2}} \cdf(\sigma \delta t).
        \end{equation}
        We rewrite $\tilde\sigma^2 = \sigma^2 + \sigma_*^2$, and $\tilde{\sigma} \tilde{\delta} = \sigma \delta$.

        \begin{equation}
            M_{S\mcal{N}*\mcal{N}}(t) = 2 * \Exp{\mu t + \frac{\tilde{\sigma}^2 t^2}{2}} \cdf_{\mcal{N}}(\tilde{\sigma} \tilde{\delta} t)
        \end{equation}
        This has the same form as the skew-normal, only with $\sigma \rightarrow
        \tilde\sigma$ and $\delta \rightarrow \tilde\delta$.

        Therefore the PDF of the convolved distribution is

        \begin{equation}
            \pdf_{S\mcal{N}\cdot\mcal{N}}(x;\mu,\sigma,\alpha) = 2\pdf_{\mcal{N}}(x;\mu,\tilde{\sigma})\cdf_{\mcal{N}}(\tilde{\alpha} x;\mu,\tilde{\sigma})
        \end{equation}
        where $\tilde\sigma^2 = \sigma^2 + \sigma_*^2$ and $\tilde{\alpha} = \frac{\tilde{\delta}}{\sqrt{1-\tilde{\delta}^2}}$, $\tilde{\delta} = \frac{\sigma}{\tilde{\sigma}}\delta$.

        The truncation of this distribution to the range $(a, b]$ is

       \begin{equation}
            \pdf_{S\mcal{N}_T}(x; \mu, \sigma, \alpha, a, b)
                = \begin{cases}
                    \frac{\pdf_{S\mcal{N}}(x; \mu, \tilde\sigma, \tilde\alpha)}{\cdf_{S\mcal{N}}(b; \mu, \tilde\sigma, \tilde\alpha) - \cdf_{S\mcal{N}}(a; \mu, \tilde\sigma, \tilde\alpha)} & a < x \leq b \\
                    0 & \text{else}
                \end{cases}
        \end{equation}
        where
        \begin{equation}
            \cdf_{S\mcal{N}}(x; \mu, \sigma) = \frac{1}{2} \text{erfc}\left(\frac{\mu -x}{\sqrt{2} \sigma }\right)-2 T\left(\frac{x-\mu }{\sigma },\alpha \right)
        \end{equation}
        where $T$ is Owen's T distribution \citep{Owen_1956}.

% \section{Plots} \label{app:developing_plots}

%     \begin{figure*}
%         \centering
%         \includegraphics[scale=0.7]{figures-not-syw/gd1/tot_stream_and_spur_prob.png}
%         \caption{Caption}
%         \label{fig:gd1-astrometrics}
%     \end{figure*}

%     \begin{figure*}
%         \centering
%         \includegraphics[scale=0.8]{figures-not-syw/gd1/Unknown.png}
%         \caption{Caption}
%         \label{fig:gd1-distance_modulus}
%     \end{figure*}

%     \begin{figure}
%         \centering
%         \includegraphics[scale=0.6]{figures-not-syw/gd1/Unknown-2.png}
%         \caption{Caption}
%         \label{fig:gd1-cmd}
%     \end{figure}

%     \begin{figure*}
%         \centering
%         \includegraphics[scale=0.6]{figures-not-syw/gd1/Unknown-4.png}
%         \caption{Caption}
%         \label{fig:gd1-cmds}
%     \end{figure*}

\end{document}