### 2023-11-22 src/tex/ms.tex
[showyourwork] overleaf sync
Signed-off-by: nstarman <nstarman@users.noreply.github.com>

% Define document class

\PassOptionsToPackage{table}{xcolor}

\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsmath, amssymb}
\usepackage{gensymb}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage[inline]{enumitem}
\usepackage{cleveref}
\usepackage{transparent}
\usepackage{todonotes}
\usepackage{array}
\usepackage{float}

% Change the section names
\renewcommand{\sectionautorefname}{\S\!}%
\renewcommand{\subsectionautorefname}{\S\!\!}%
\renewcommand{\subsubsectionautorefname}{\S\!}%
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\tableautorefname}{Table}
\renewcommand{\figureautorefname}{Fig.}

\newcommand{\code}[1]{\textsc{#1}}
\newcommand{\package}[1]{\code{#1}}
\newcommand{\stream}[1]{#1}
\newcommand{\dataarchive}[1]{\textit{#1}}
\newcommand{\Gaia}{\dataarchive{Gaia}}
\newcommand{\PanStarrs}{\dataarchive{Pan-STARRS1}}

% Math fonts
\newcommand{\mrm}[1]{\mathrm{#1}}
\newcommand{\mbs}[1]{\boldsymbol{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mfk}[1]{\mathfrak{#1}}
\newcommand{\mcal}[1]{\mathcal{#1}}

% Math symbols
\newcommand{\Exp}[1]{e^{#1}}
\newcommand{\Erf}[1]{\mathrm{erf}\left(#1\right)}
\NewDocumentCommand \dif {o m} {
	\IfNoValueTF{#1}
		{ \mathrm{d}{#2} }
		{ \mathrm{d}^{#1}{\!#2} }
}

% Statistics
\newcommand{\pdf}{p}
\newcommand{\cdf}{P}
\newcommand{\likelihood}{\mcal{L}}
\newcommand{\prior}{\mcal{\pi}}
\newcommand{\posterior}{p}

% Labels
\newcommand{\nth}[1]{{#1}_{\mrm{n}}}  % measurement
\newcommand{\fth}[1]{{#1}_{\mrm{f}}}  % feature
\newcommand{\qth}[1]{{#1}_{\mrm{q}}}  % model index
\newcommand{\unit}[1]{[\text{#1}]}

\newcommand{\smallcomponent}[2]{#2^{\scriptscriptstyle (#1)}}
\newcommand{\cmp}[2]{\smallcomponent{#1}{#2}}
\newcommand{\Scmp}[1]{\cmp{S}{#1}}
\newcommand{\Bcmp}[1]{\cmp{B}{#1}}
\newcommand{\astroM}[1]{{#1}_w}
\newcommand{\photoM}[1]{\smallcomponent{m}{#1}}
\newcommand{\chemoM}[1]{\smallcomponent{z}{#1}}

\newcommand{\Spdf}{\Scmp{\pdf}}
\newcommand{\Bpdf}{\Bcmp{\pdf}}

\newcommand{\parallax}{\varpi}
\newcommand{\sigobs}{{\sigma_*}}

\newcommand{\galaxy}[1]{
    \ifthenelse{\equal{#1}{}}{
        \includegraphics[width=9pt]{static/galaxy.png}
    }{
        \includegraphics[width={#1}pt]{static/galaxy.png}
    }%
}

\defcitealias{Ibata+2021}{I21}

\newcommand{\TODO}[1]{{\textcolor{red}{#1}}}
\newcommand{\JN}[1]{\TODO{#1}}

% Begin!
\begin{document}

% Title
\title{Does this Star Belong? Data-Driven Characterization of Stellar Streams \\ with Mixture Density Networks}

% Author list
\author[0000-0003-3954-3291]{Nathaniel Starkman}
\affiliation{David A Dunlap Department of Astronomy and Astrophysics, University of Toronto}

\author[0000-0001-8042-5794]{Jacob Nibauer}
\altaffiliation{Co-first author.}
\affiliation{Department of Astrophysical Sciences, Princeton University}

\author[0000-0001-6855-442X]{Jo Bovy}
\affiliation{David A Dunlap Department of Astronomy and Astrophysics, University of Toronto}

\author[0000-0003-3613-0854]{Jeremy J. Webb}
\affiliation{Division of Natural Science, Department of Science, Technology and Society, York University}

% \author[0000-0002-7846-9787]{Ana Bonaca}
% \affiliation{Carnegie Observatories: Pasadena, California, US}

\correspondingauthor{Nathaniel Starkman}
\email{n.starkman@mail.utoronto.ca}

\shortauthors{Starkman and Nibauer et al.}

% Abstract with filler text
\begin{abstract}
    Stellar streams are sensitive probes of the Milky Way's gravitational
    potential. The mean track of a stream constrains global properties of the
    potential, while its fine-grained surface density contains information about
    galactic substructure. A precise characterization of streams from potentially noisy data marks a crucial step in constraining galactic structure across orders of magnitude in mass scales. Here we present a new method for constructing a smooth probability
    density model of stellar streams using all of the available astrometric and
    photometric data. To characterize the stream's morphology and kinematics, we
    utilize mixture density networks to construct a flexible representation of
    the stream and its on-sky track, width, stellar number density, and
    kinematic distribution. In photometric coordinates we model each stream as a
    single-stellar population, with a distance track that is simultaneously
    estimated from the stream's inferred distance modulus (using photometry) and
    parallax distribution (using astrometry). Normalizing flows are used to characterize the distribution of background stars. We apply the method to the stream
    \stream{GD-1}, and the tidal tails of \stream{Palomar 5}. For both streams we obtain a catalog
    of stellar membership probabilities that are made publicly available.
    Importantly, our model is capable of handling data with incomplete
    phase-space observations, making our method applicable to the growing census
    of Milky Way stellar streams. When applied to a population of streams, the
    resulting membership probabilities from our model form the required input to
    infer the Milky Way's dark matter distribution from the scale of the stellar
    halo down to subhalos. 
\end{abstract}

\section{Introduction} \label{sec:intro}

    Stellar streams are the disrupted remnants of globular clusters and
    satellite galaxies, and their census has grown extensively in the Milky Way
    thanks to all-sky astrometric missions like \Gaia{}
    \citep{GaiaCollaboration+2016, GaiaCollaboration+2023} and low-surface
    brightness photometric surveys (e.g., \citealt{Ibata+1998,
    GrillmairDionatos2006, Shipp+2018, Ibata+2021}).  The formation of streams
    can be traced back to a progenitor (a globular cluster or satellite galaxy)
    whose stars are gradually lost to the tidal field of the host galaxy. The
    stars extend along a series of similar orbits, called tidal tails, which can
    sometimes span several tens of degrees across the sky. The orbital proximity
    of stars belonging to a stream makes them sensitive tracers of the mass
    distribution of galaxies, sourced by both the baryonic and dark matter
    components.

    The average properties of a stream (e.g., the mean position and velocity
    track of the tidal tails) can be used to reconstruct the global dark matter
    distribution of the stellar halo (e.g., \citealt{Johnston+1999, Bonaca+2014,
    Bovy2014, Bovy+2016, Nibauer+2022, Koposov+2023}), while local variations in
    the stream track and surface density encode information about the stream's
    encounters with baryonic structures such as the stellar bar (e.g.,
    \citealt{Erkal+2017, Pearson+2017}), and non-luminous structures like dark
    matter subhalos (e.g., \citealt{Ibata+2002, Johnston+2002,
    Carlberg+2012-74820C, Bovy+2017, Bonaca+2019, Hermans+2021}).

    While several methods have been developed to model stream properties as a
    function of global halo characteristics (e.g., the mass distribution,
    flattening; \citealt{Johnston+1999,2008MNRAS.386L..47B,Koposov+2010,2013MNRAS.433.1826S,Bovy2014,Nibauer+2022}) and local substructures (e.g., the bar's pattern
    speed and the mass of subhalo perturbers; \citealt{2014A&A...563A..60A,Pearson+2017,2016MNRAS.460..497H,Bovy+2017,Bonaca+2019}), these methods often
    rely on the existence of a homogenized stream-tracer population that
    characterizes a given stream across each of the observable phase-space
    dimensions. Additionally, because even the fine-grained properties of
    streams are influenced by the detailed mass distribution of its host galaxy,
    it is important to propagate errors both in measurement uncertainty and
    membership uncertainty (i.e., which stars belong to the stream versus the
    background) when attempting to model a stream. Otherwise, a poorly
    characterized stream could bias constraints on, e.g., the total mass of the
    galaxy, the dark matter halo shape, and the mass of subhalo perturbers. In
    order to generate more precise constraints on the structure of the galaxy,
    it is therefore imperative that a statistically sound and homogenized
    catalog of stellar streams is produced.

    While streams are well described in action-angle coordinates
    \citep[e.g.][]{Bovy2014} or (restricted) $N$-body simulations
    \citep[e.g.][]{Dehnen+2004}, these methods rely on prior assumptions about
    the galaxy and its underlying gravitational potential. Therefore, it is
    necessary to devise data-driven methods to characterize stellar streams
    without appealing to strong prior assumptions about the galaxy's structure.

    There are a few previous works which have modeled stellar streams in this
    context. \citet{Patrick+2022} developed the method introduced by
    \citet{Erkal+2017}, which is a data-driven, spline-based method for modeling
    the photometric properties and on-sky positions of previously identified
    streams. Their method works by fitting the stream's stellar surface density
    with a series of splines, capturing both surface brightness variations along
    the stream and changes in the position of the stream track and width. Their
    method also utilizes an isochrone fit to a given stream, from which a
    distance modulus can be derived. While this method is very successful at
    characterizing the photometric properties of a stream, it does not consider
    the kinematic dimensions which have become increasingly well-measured with
    successive data releases from \Gaia{}, as well as targeted observations.
    From a population of RR Lyrae stars, \citet{Price-Whelan+2019} utilized a
    mixture model to fit the globular cluster stream Pal-5 in the space of
    proper motions and heliocentric distances. The mixture modeling approach
    enables a simultaneous fit to both the stream and background density,
    allowing for better separation of the stream from the background
    distribution. However, the track of the stream in on-sky coordinates is not
    fit. The method identified 27 RR Lyrae stars consistent with being members
    of the stream. While useful for determining the average properties of a
    stream (e.g., its mean distance track), a limited population of RR Lyrae
    stars are not sufficient to capture the small-scale properties of streams
    expected to inform constraints on, e.g., a population of perturbers in the
    galaxy.

    Other methods for extracting stellar streams from the data include
    matched-filter algorithms, where over-densities in color-magnitude space are
    inspected as possible clusters undergoing tidal stripping
    \citep{Grillmair+1995, GrillmairJohnson2006, Shipp+2018}. While the
    matched-filter technique has been extremely successful in stream discovery,
    it is not equipped to characterize streams at the level of detail necessary
    to inform precise constraints on the potential of the galaxy and its
    substructure.

    In order to address the need for a flexible model of stellar streams both in
    photometric and kinematic spaces, in this work we develop a new method for
    characterizing streams and quantifying the membership probability of
    possible stream stars. Our approach provides several key advantage over
    previous stream modeling efforts, specifically in its ability to jointly
    model the kinematic and photometric properties of streams simultaneously
    using a flexible model.  Furthermore, our method is well-suited to fitting
    streams with only partial or incomplete phase-space or photometric
    observations, enabling a full characterization of streams insofar as the
    available data allows.

    \paragraph{Paper organization}

        The paper is organized as follows.  In \autoref{sec:method} we introduce
        our method for characterizing stream tracks and computing stellar
        membership.  We discuss initial data processing in
        \autoref{sub:method:framing_the_data} and build a probabilistic
        framework to model the stream and background in
        \autoref{sub:method:likelihood_setup}-\autoref{sub:methods:priors}.  In
        \autoref{sec:results_mock} we apply our method to simulated data to
        demonstrate the model with a known ground truth dataset, then to the
        stellar streams \stream{GD-1} (\autoref{sec:results_gd1}) and
        \stream{Pal\,5} (\autoref{sec:results_pal5}).  We compare our approach
        to characterizing stellar streams to other methods in
        \autoref{sec:discussion}.  We conclude in \autoref{sec:conclusions},
        discussing the results of this work and future directions.

% section introduction (end)

\section{Method} \label{sec:method}

    To characterize stellar stream tracks and for computing membership
    probabilities from kinematic and photometric measurements of stars, we
    develop a flexible Bayesian modeling framework that incorporates all of the
    available kinematic information for a field of stars, including those with
    only partial phase-space measurements. Below, we offer a concise overview of
    our approach, followed by a more in-depth technical explanation in
    subsequent sections.  We start, in \autoref{sub:method:framing_the_data}, by
    transforming the astrometric data into a stream-oriented on-sky reference
    frame ($\phi_1, \phi_2, \parallax, \mu_{\phi_1}^*, \mu_{\phi_2}, v_r)$, for
    each stream. In this frame, the stream track lies along $\phi_1$, and the
    stream density may be described by distributions in the other parameters as
    injective functions of $\phi_1$.  Likewise, the background of the Galactic
    field may be described with $\phi_1$-conditioned distributions.  We will
    also model the stream in photometric coordinates. Additional features, e.g.
    metallicity, can also be modeled along the stream and background as
    functions of $\phi_1$, though we leave this to a future work.

    The distributions for the stream and background are collected in a single
    mixture model that describes the entire field.  We take the parameters of
    the mixture model -- the mixture weights and component distribution
    parameters -- to be general (continuous) functions of the conditioning
    variable, $\phi_1$. This is done using a feed-forward neural network that
    takes the conditioning variable as its input and outputs the parameter
    value(s) that characterize the distribution of stars at a $\phi_1$ ``slice".
    Mixture models of conditional probability distributions using neural
    networks are called Mixture Density Networks (MDNs) \citep{Bishop+1994}.
    Since the model coefficients may evolve with $\phi_1$, MNDs are well suited
    to characterize streams (and the background) with all their variations in
    position, width, and linear density.

    % sub: framing_the_data (end)

    \subsection{Framing the Data} \label{sub:method:framing_the_data}

        As a stream progenitor orbits the Galactic center of mass, its path can
        be approximated locally as an elliptical segment. The associated arms of
        the stream generally align closely with its orbital plane.  As viewed
        from a Galactocentric reference frame, the stream approximately assumes
        the form of a great circle enabling a 1:1 mapping from some phase
        parameter to a unique position along the stream. Thus in a `stream'
        oriented frame the rotated longitudinal and latitudinal coordinates
        ($\phi_1, \phi_2$) align with the stream such that the stream's phase is
        $\phi_1$ and $\phi_2 \simeq 0$.

        While it would be ideal to characterize streams in this Galactocentirc
        frame, transformation to the frame requires knowledge of the distance to
        the Galactic center  \citep{GRAVITY2018, BennettBovy2019, Leung+2022},
        the solar motion around the Galactic center, and accurate distances to
        field stars (among other uncertain quantities). In a heliocentric frame,
        e.g. the standard ICRS\footnote{International Celestial Reference
        System} \citep{ICRS1997}, projection effects can obscure the shape of a
        stream leading to a non-trivial stream track. Therefore in a
        heliocentric frame, there is no assurance of effecting a rotation like
        in the Galactocentric frame that would similarly align the stream in an
        observer's coordinates such that $\phi_2(\phi_1) \approx 0$.
        Thankfully, most streams are adequately distant from both the Galactic
        center and the observer, or exhibit orientations such that a rotation a)
        can be established or b) holds valid. For most of the best-studied
        globular cluster streams -- \stream{GD-1}, \stream{Palomar 5},
        \stream{Jhelum}, etc. -- this holds true. However for some streams of
        dwarf galaxies, like the Magellanic stream which wraps multiple times
        around the Milky Way \citep{WannierWrixon1972}, no such transformation
        is possible. More generally, a stream's ``sky"-projected path may be
        kinked, e.g. by subhalo interactions, such that it is a many-to-one
        function in $\phi_1$.

        In this work we treat with the vast majority of streams, which have a
        ``sky" transformation such that the stream is an injective function in
        $\phi_2(\phi_1)$.  Since we are examining known streams this
        transformation is known \textit{a priori}, or may be constructed during
        pre-processing \citep[see][\S2.1]{Starkman+2023}.  We note that, with
        sufficient prior knowledge, many-to-one functions may be broken into
        injective segments and their connection constrained by a prior. Thus the
        methods developed subsequently may be extended and applied even to
        wrapped and kinked streams. However, we leave extending the model
        framework to a future work.

    \subsection{Likelihood Setup}\label{sub:method:likelihood_setup}

        \begin{figure*}
            \script{pgm.py}
            \centering
            \includegraphics[scale=0.6]{figures/pgm.pdf}
            \caption{%
                Probabilistic Graphical Models (PGMs) of the Mixture Density
                Networks used in this work.
                %
                \textbf{Left}: A PGM of the stream model. The full dataset
                $x_n^{obs}$ is made of the observed astrometry $w_n^{obs}$ and
                photometry $m_n^{obs}$, along with the corresponding
                observational errors. We model the data as a mixture of
                Gaussians (in astrometry + distance modulus), with the Gaussians
                indexed by $q$. The weight $\alpha$ determines the weighted
                contribution of each Gaussian to each datum. All models are
                conditioned on $\phi_1$.
                %
                \textbf{Right}: a PGM of the background model. The data is
                identical to that of the stream, however the model is not a
                mixture of Gaussians, but of a variety of distributions. We
                further distinguish between trainable distribution parameters
                (circles) and fixed distributions (points), like a pre-trained
                normalizing flow background model, explained in further detail
                in \autoref{sub:method:pre-training_distributions}. Alike to the
                stream model, all background models are conditioned on $\phi_1$.
                % 
            }
            \label{fig:PGM}
        \end{figure*}

        Mixture models (MMs) are a statistical method to represent a population
        as a set of a sub-populations specified by probability distributions.
        These distributions are characterized by their parameters and the
        mixture by a corresponding set of mixture coefficients. Scalar mixture
        models may be extended to model completely general conditional
        distributions by making the distribution parameters and mixture
        coefficients into functions of a conditional input
        \citep{McLachlanBasford1989}.  Mixture Density Networks (MDNs) are a
        machine learning method to implement these general MM, taking the
        parameters and coefficients to be the outputs of feed-forward neural
        networks and as such, general continuous functions \citep{Bishop1994}.
        Functional parameters allow the model to capture variation over those
        parameters that is not possible with scalar-valued models. Moreover, by
        using neural networks the inferred functional forms are ``model-free"
        and driven entirely by the data.

        For the MDN we take $x_{\cdot, 0}$, which is the first feature column of
        the data $\mbf{X}$, as the `independent' coordinate over which the MDN
        parameters are conditioned. In the context of streams $x_{\cdot, 0}
        \equiv \phi_1$ is the longitude coordinate in a reference frame rotated
        such that the stream is aligned along $\phi_1$. The parameters of the
        stream, the background field, and the mixing coefficients of the two are
        now general data-driven continuous functions along $\phi_1$.

        The PDF of a general MDN is
        \begin{equation} \label{eq:general_mixture_network}
            \!\!\! \pdf(\nth{\mbs{x}} | \mbs{\theta}(\phi_1))
            \!=\! \sum_{\mrm{q} \in \mbs{I_Q}} \! \qth{f}(\phi_1) \cmp{q}{\pdf}(\nth{\mbs{x}}|\qth{\mbs{\theta}}(\phi_1)),
            % \prior(\nth{\mbs{x}}, \phi_1)
        \end{equation}
        where $\cmp{q}{\pdf}$ is the PDF of the $q$-th model and $f_q$ its
        mixing coefficient for all $q$ models in the set of models indexed by
        $\mbs{I_Q}$. $\mbs{\theta}$ is the $\phi_1$-conditional output of the
        feed-forward neural network, and $\nth{\mbs{x}}$ the data on the $n$-th
        star in $\mbs{X}$, not including $\phi_1$. For convenience we drop the
        implied subscript $n$ in $\phi_1$, the explicit notation of
        $\mbs{\theta}$'s dependence on $\phi_1$, and the index set $\mbs{I}_X$
        in summations over its elements $x$.

        The mixture coefficients $\qth{f}$ are normalized such that
        \begin{equation} \label{eq:mixture_weight_sum}
            \sum_{\mrm{q}} \qth{f}(\phi_1) = 1, \qquad \forall \phi_1
        \end{equation}
        In practice this is enforced by defining a background
        \begin{equation} \label{eq:mixture_weight_normalization}
            f_{b}(\phi_1) = 1 - \sum_{\rm{q}\in \mbs{I_Q}\backslash \{b\}} \qth{f}(\phi_1),
        \end{equation}
        the summation over all models except for some ``background" index $b \in
        \mbs{I_Q}$.  In the context of Galactic streams the ``foreground" is the
        stream itself, while the background is the Galaxy - bulge, bar, disc,
        halo - and additional structures, like globular clusters and dwarf
        galaxies, that are not associated with the stream itself.  We discuss
        the mixture weights further in \autoref{sub:method:mixture_weight}.

        In this work we consider primarily astrometric and photometric data,
        (notated $\mbs{W}$ and $\mbs{M}$, respectively), but note that the MDN
        framework may be extended to arbitrary feature dimensions. Assuming the
        conditional independence of the astrometry and photometry\footnote{ For
        \Gaia{} data this is true in practice but not in detail: there exist
        small correlations between the astrometric and photometric data.}, we
        split each model into linearly separable astrometric and photometric
        models $\cmp{q}{\pdf} = \cmp{q,w}{\pdf} \cmp{q,m}{\pdf}$.

        The likelihood for a single star then becomes, simplifying from
        \autoref{eq:general_mixture_network},
        \begin{equation} \label{eq:general_model}
            \!\!\!\!\! \pdf(\nth{\mbs{w}}, \! \nth{\mbs{m}} | \mbs{\theta})
                \!=\! \sum_{q} \! f_q(\phi_1) \cmp{q,w}{\pdf}(\nth{\mbs{w}}|\mbs{\theta}) \cmp{q,m}{\pdf}(\nth{\mbs{m}}|\mbs{\theta})
        \end{equation}

        Assuming the measurement of each star is independent, the total
        log-likelihood is a sum over all stars:
        \begin{equation} \label{eq:general_likelihood}
            \ln\mcal{L}\left(\{\nth{\mbs{w}},\nth{\mbs{m}}\} | \mbs{\theta}\right) = \sum_n \ln \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta}).
        \end{equation}

        It will prove convenient to distinguish the models for $q$ as the
        \textit{background} and \textit{stream} models, respectively. In this
        notation:
        \begin{align} \label{eq:stream_and_bkg_prob}
            \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})
            =& \phantom{+} \qquad f \phantom{+} \Spdf(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) \\
            & + (1-f) \Bpdf(\nth{\mbs{w}}|\mbs{\theta}) \Bpdf(\nth{\mbs{m}}|\mbs{\theta}), \nonumber
        \end{align}
        where we have dropped the $(w), (m)$ superscript as the model types are
        evident from their inputs. We applied
        \autoref{eq:mixture_weight_normalization} to require only a single
        mixture coefficient.  Note that the stream and background PDFs can
        consist of further components. This will become important for our
        characterization of the stream \stream{GD-1}, which appears spatially
        bifurcated. We note this as an implementation detail since all
        stream-related components (cold, extra-tidal, perturbed) and all other
        components (stellar halo, globular clusters, etc) may be grouped into
        multi-component \autoref{eq:stream_and_bkg_prob} \textit{stream} and
        \textit{background} models.  The stream and background models are
        included graphically in \autoref{fig:PGM}, where the left plot shows the
        stream model, and the right the background.

        The probability of a being a stream member is then simply
        \begin{equation}\label{eq:membership_prob}
            \pdf\!\left(S | \nth{\mbs{w}}, \nth{\mbs{m}}, \mbs{\theta} \right)
            = \frac{f \Spdf(\nth{\mbs{w}}|\mbs{\theta}) \Spdf(\nth{\mbs{m}}|\mbs{\theta}) }{ \pdf(\nth{\mbs{w}}, \nth{\mbs{m}} | \mbs{\theta})}.
        \end{equation}

        Hereafter, when we refer to ``membership probability,"
        \autoref{eq:membership_prob} provides the formal definition. We nuance
        this definition by considering missing feature dimensions in the
        following subsection.

    %\newpage
    \subsection{Missing Phase-Space Observations}
    \label{sub:method:missing_data}

        Streams have a diversity of measurement coverage: several have full 6-D
        astrometrics as well as photometry and chemical abundances available
        \citep[e.g.,][]{Koposov+2019, Antoja+2020, Li+2022}, while most have
        partial coverage, lacking some feature dimensions.  Most streams in
        \Gaia \citep{Gaia2016, Gaia2023}, for instance, have proper motions but
        not radial velocities nor high signal-to-noise parallax measurements.
        Follow-up surveys, like \dataarchive{S5} \citep{Li+2019}, add many
        feature dimensions, however only for a select set of streams. Given the
        diversity of coverage it is important that the likelihood model --
        \autoref{eq:membership_prob} -- account for missing data.

        The method of treatment of missing data depends on the cause for that
        data to be missing.  Broadly, there are two categories of causes:
        randomness and systematics.  Consider a dataset with all features
        measured but with a random subset of features masked.  One possible
        approach is to attempt to impute the missing features, then feed the
        full-featured data to the likelihood model. Alternatively, the
        likelihood model might use only the present features, ignoring those
        missing. If however, features are missing for systematic reasons, then
        this approach can introduce systematic bias. Consider a star too far
        away to measure its parallax distance or radial velocity. This is a
        systematic reason and this information should be incorporated into any
        model as a prior. Not doing so, the data imputation approach might
        predict too-close distances. Likewise, using only present features in
        the likelihood model misses that the star is likely very distant.  The
        effects of mask systematics are mitigated by not including such data,
        moving the contribution of the systematics from a prior to the evidence.
        Our models compute likelihoods under a specific dataset, and
        \autoref{eq:membership_prob} is a ratio of the probabilities, so the
        removed systematics are not expected to be impactful.

        We adopt the approach of modeling the data masks as randomly-distributed
        and not systematic. We do not impute missing data, instead the per-star
        likelihood is modelled with a distribution with the same dimensionality
        as the data vector.  For example, a star with a 6D astrometric
        measurement will be modelled with a full 6D distribution, while a star
        missing its radial velocity measurement will be modelled with a
        dimensionally-reduced (5D) form of the 6D distribution. Using
        dimensionally reduced distributions ensure that missing phase-space
        dimensions do not amplify nor suppress the likelihood.  Notwithstanding,
        we include a quality flag on our calculations indicating the number of
        missing features.  We leave incorporating priors on the data mask
        distribution to future work.

    % sub: missing_phase_space_observations (end)

    \subsection{Pre-training distributions} 
    \label{sub:method:pre-training_distributions}

        In \autoref{sub:method:likelihood_setup} we set up the general
        mathematical framework of MDNs then specified the models on two axes:
        the component and the data type. Broadly, the components are stream
        versus background and the data types are astrometric and photometric.
        For some components, e.g. the stream, the data density will be well fit
        by an analytic function.  For others, no analytic density provides a
        realistic description of the data. In this work we take a data-driven
        approach, and represent non-analytic distributions using a normalizing
        flow model.

        Normalizing flows provide a flexible description of a probability
        density, by transforming samples generated from a simple base
        distribution (i.e., a Gaussian) to the more complicated target
        distribution \citep{EstebanVanden2010, Rippel2013, RezendeMohamed2015}.
        The transformations are typically non-linear outputs of a neural neural
        network, whose parameters are optimized until the target density
        accurately characterizes the data. The neural network transformation is
        differentiable, so that jacobian factors can be combined to ensure that
        the target distribution remains a valid probability density (i.e.,
        integrates to one) \citep{Kobyzev+2019}.  For the neural networks we
        regularize the data (minus the mean, divided by the standard deviation
        per feature column); thus we correct the overall normalization of the
        flow by the product of the jacobian of the regularization operation.

        Our implementation is as follows. For a given stream, we restrict the
        data to some field characterized by a polygon drawn around the stream.
        Because our density modeling is aimed at stream characterization rather
        than discovery, we assume that the stream is sufficiently well known so
        that it can be masked out as a simple cut in (e.g.) $\phi_1-\phi_2$
        coordinates. With the stream masked out, we then fit a normalizing flow
        to the data. Because the distribution can evolve with $\phi_1$, we fit a
        \emph{conditional} normalizing flow, $q_{\rm flow}(\mathbf{x}|\phi_1)$.
        Our central assumption with this model is that the distribution for
        background stars in the masked region are quantitatively similar to
        those in the unmasked region. For a sufficiently narrow mask we expect
        this to be the case, and find that this assumption does not introduce
        bias for the streams considered in this work.

        Once the normalizing flow, $q_{\rm flow}\left(\mathbf{x} |
        \phi_1\right)$, is fit to the data, its parameters are fixed while the
        other components of the model begin training. Therefore, the normalizing
        flow can be rapidly trained and computed as a pre-processing step.

        It is easy to evaluate the log-likelihood of a normalizing flow at a
        point $\mbs{x}$ when all feature dimensions are present. When one (or
        more) feature dimension is missing, however, evaluating the marginal
        log-likelihood involves integrals over the flow. While the procedure is
        straightforward it is computationally expensive. As an alternative, we
        conservatively set the log-likelihood to $0$, the maximum value,
        overemphasizing the flow's contribution to the background model.
        Conceptually, up-weighting the background model decreases the
        false-positive rate in stream membership identification, at the cost of
        increasing the false-negative rate.  Another approach, which we leave to
        a future work, is to construct normalizing flows of each marginal
        distribution, e.g. $q_{\rm flow}\left(\{x_i\}_{i \in \mbs{I}_F, i \neq
        {j, k}} | x_j, x_k, \phi_1\right)$, and select for each $\nth{\mbs{x}}$
        the correct flow given the present features.

    % sub: pre-training_distributions (end)

    \vspace{10pt}
    \subsection{Determination of the Mixture Weight} \label{sub:method:mixture_weight}

        The mixture weight $f_q(\phi_1)$ as used in \autoref{eq:general_model}
        is characterized as the fraction of stars belonging to the mixture
        component $q$ over an increment $[\phi_1, \phi_1 + \dif{\phi_1}]$. Thus,
        the mixture weight encodes information about the linear density of the
        stream, which has been shown to depend on substructure in the galaxy
        \citep{Siegal-GaskinsValluri2009, Yoon+2011}. The dependence of the
        mixture weight on the linear density is captured by the expression
        \begin{equation}\label{eq:fraction_param}
            f_q\left(\phi_1\right) = \frac{N_q\left(\phi_1\right)}{N_{\rm tot}\left(\phi_1\right)},
        \end{equation}
        where $N_q\left(\phi_1\right)$ is the number of stars in component $q$
        within an interval $[\phi_1, \phi_1 + \dif{\phi_1}]$, and $N_{\rm tot}$
        is the total number of stars in the same band (including those in
        component $q$). The linear  density for component $q$ is then entirely
        specified by the function $N_q(\phi_1)$.

        The mixture model specified in \autoref{eq:general_model} only models
        the weight parameter $f_q$, and not the components of the ratio in
        \autoref{eq:fraction_param}. However, the linear density $N_q(\phi_1)$
        can be readily obtained from $f_q$, provided that $N_{\rm
        tot}\left(\phi_1\right)$ can be calculated. Importantly, this function
        does not depend on the parameters of the on-stream or off-stream model.
        Instead, $N_{\rm tot}(\phi_1)$ is simply the number of stars in a small
        $\phi_1$ increment. We therefore model $N_{\rm tot}\left(\phi_1\right)$
        as a post-processing step using a single normalizing flow in $\phi_1$.

        In particular, for a given field the distribution of $\phi_1$
        coordinates is specified by the set $\{\phi_1\}_f$. A normalizing flow
        is trained to this dataset, and will satisfy $\int d\phi_1
        P_{\phi_f}(\phi_1) = 1$ by construction. In order to re-weight the
        normalizing flow to obtain $N_{\rm tot}(\phi_1)$, we simply rescale
        $P_{\phi}(\phi_1)$ by $N_{\rm tot,0}$, which is the total number of
        stars in the field considered. We can then obtain the linear stream
        density with
        \begin{equation}
            N_q\left(\phi_1\right) = [f_q N_{\rm tot}] \left(\phi_1\right)  = N_{\rm tot,0}[f_q P_{\phi_1}]\left(\phi_1\right) .
        \end{equation}

    % subsection: determination_of_the_mixture_weight (end)

    \subsection{Astrometric Model} \label{sub:method:astrometric_model}

        \subsubsection{On-Stream} \label{ssub:method:astrometric_model:on_stream}

            The N-dimensional astrometric track is modeled with a single neural
            network, parameterized as a function of $\phi_1$.  The neural
            network output for a single stream component is the mean location of
            the stream track in each dimension, $\astroM{\mbs{\mu}}(\phi_1)$,
            and its associated covariance matrix
            $\astroM{\mbs{\Sigma}}(\phi_1)$. The probability for the data is
            then
            \begin{equation}
                \cmp{S,w}{\pdf}(\nth{\mbs{w}} | \mbs{\theta})
                \!= \mcal{N}_F \big(\nth{\mbs{w}} | \astroM{\mbs{\mu}}\!(\phi_1), {\cmp{w}{\nth{\mbs{\Sigma}}}}^2 \! + \!  \astroM{\mbs{\Sigma}}^2\!\left(\phi_1\right) \!\big),
            \end{equation}
            where $\astroM{\mbs{\theta}}$ is the set of neural network
            parameters that define $\astroM{\mbs{\mu}}$ and
            $\astroM{\mbs{\Sigma}}$; and $\cmp{w}{\nth{\mbs{\Sigma}}}$ the data
            uncertainties.  In the simplest case, we take
            $\cmp{w}{\nth{\mbs{\Sigma}}}, \astroM{\mbs{\Sigma}}$ to be diagonal,
            consisting of both intrinsic dispersion for each dimension of
            $\mbs{w}$ and the data uncertainties added in quadrature. The data
            uncertainties $\cmp{w}{\nth{\mbs{\Sigma}}}$ are elements of
            $\mbs{\delta}\mbf{W}$, the dataset of astrometric uncertainties.
            % (see \autoref{app:notation}).

            In practice, we work with truncated Gaussians, where the truncation
            is set by the size of the field in each astrometric dimension.
            Details of the truncated Gaussian are included in
            \autoref{app:sub:normal_distribution}. Truncating to the field
            ensures that the probability density is zero wherever there is no
            data. This has an added benefit of correctly normalizing the
            Gaussian distribution, which is not compact (has non-zero
            probability over all space) and would otherwise not integrate to one
            within the field.

            As a MDN, we implement the Gaussian distribution parameters with a
            multi-layered feed-forward neural network. The number of layers
            required depends on the maximum dimensionality of the Gaussian, and
            whether the neural network employs regularization techniques like
            dropout. We use alternating layers of linear and tanh units. Details
            of the network architecture are discussed in
            \autoref{sub:methods:model_implementation}.

            In the presence of missing phase-space coordinate $f_k$ we reduce
            the dimensionality of the Gaussian to that of the data-vector.
            Namely, for $\mbs{I}_{\tilde{F}} \equiv \mbs{I}_{F} \backslash
            \{f_k\}$
            \begin{align}
                \nth{\tilde{\mbf{w}}} &\mapsto \nth{\tilde{\mbf{w}}} = \{ w_{n,f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\mu}} &\mapsto \astroM{\tilde{\mbs{\mu}}} = \{ \mu_{f} | f\in \mbs{I}_{\tilde{F}}\}, \nonumber\\
                \astroM{\mbs{\Sigma}} &\mapsto \astroM{\tilde{\mbs{\Sigma}}} = \{ \Sigma_{i,j} | i,j \in \mbs{I}_{\tilde{F}} \}, & \nonumber
                \intertext{such that,}
                \Spdf(\nth{\mbs{w}} & | \astroM{\mbs{\theta}}(\phi_1))
                    \mapsto \Spdf(\nth{\tilde{\mbs{w}}} | \astroM{\tilde{\mbs{\theta}}} (\phi_1))\,. % = \mcal{N}_{\tilde{F}}.
            \end{align}

        % ssub: on_stream (end)

        \subsubsection{Off-stream} \label{ssub:method:astrometric_model:off_stream}

            We now outline the off-stream astrometric model. We take a similar
            approach to the on-stream astrometric model discussed in
            \autoref{ssub:method:astrometric_model:on_stream}, though we
            typically do not characterize the backgrounds with a Gaussian
            distribution. Instead, we utilize a range of distributions (i.e.,
            exponential, skew-normal) discussed in \autoref{app:distributions}.
            The parameters of the user-specified background distributions are
            themselves neural network outputs, such that any $\phi_1$ ``slice"
            of the background should be characterized by some analytic density,
            but the evolution of the backgrounds with $\phi_1$ can be complex.
            Details of the network architecture are discussed in
            \autoref{sub:methods:model_implementation}.

            Let $\Bcmp{\fth{\mbs{\theta}}}(\phi_1)$ represent the neural network
            which outputs a vector of parameters characterizing the background
            distribution at a given $\phi_1$ for astrometric dimension $f$. The
            number of outputs for the background model is equal to
            $\mathrm{dim}\left(\fth{\Bcmp{\mbs{\theta}}}\right)$ (where $f=0$
            corresponds to $\phi_1$, which we do not model). At a given
            $\phi_1$, the probability density for the astrometric dimension
            $f>0$ is $\Bcmp{\fth{\pdf}}\left(\mbs{w}_{n,f} |
            \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right)$. We treat each background
            component dimension as independent, allowing us to write the
            likelihood as a product over the astrometric dimensions:
            \begin{equation} \label{eq:astrometric_model_off_stream_probability}
                \pdf^{(B)}\left(\nth{\mbs{w}} | \mbs{\theta}(\phi_1) \right) = \prod_{f\in \mbs{I}_{\tilde{F}}} \pdf_f^{(B)}\left(\mbs{w}_{n,f} | \Bcmp{\fth{\mbs{\theta}}}(\phi_1) \right)\,,
            \end{equation}
            where $\mbs{I}_{\tilde{F}}$ is the set of indices of the features,
            omitting missing phase space dimensions.

        % ssub:method:astrometric_model:off_stream (end)

    % sub: astrometric_model (end)

    \subsection{Photometric Model} \label{sub:method:photometric_model}

        \subsubsection{On-Stream} \label{ssub:method:photometric_model:on_stream}

            In photometric coordinates we model the stream as a
            single-population isochrone with the possibility of a non-zero
            distance gradient. We accomplish this by modeling the distance of
            the isochrone as a neural network output, parametrized as a function
            of $\phi_1$. The result is a distance track along the stream,
            estimated from the variation in its isochrone with $\phi_1$.

            At any given $\phi_1$, the isochrone is modeled in absolute
            magnitudes and parameterized by the normalized stellar mass, labeled
            $\gamma \in (0, 1]$, in the modeled mass range. For example,
            increasing $\gamma$ corresponds to movement up the main sequence
            towards the red giant branch. We denote the isochrone as
            $\mcal{I(\gamma)}$. The intrinsic dipsersion around the isochrone is
            $\Sigma_\mcal{I}(\gamma)$. Since the isochrone is in absolute
            magnitudes, but the data are in apparent magnitudes, it is necessary
            to shift the isochrone by a distance modulus to the predicted
            distance of the stream, labeled  $\mu(\phi_1)$, with intrinsic
            stream distance variation $\sigma_\mu(\gamma)$.

            For the $n$-th star, the photometric model is:
            \begin{multline} \label{eq:photometric_probability_point}
                \cmp{S,m}{\pdf}(\mbs{m_i}, \gamma | \mbs{\theta}(\phi_1)) 
                \\ \sim \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1} \sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \\ \times \prior(\gamma, \nth{\mbs{m}}, \phi_1),
            \end{multline}

            where $\cmp{m}{\nth{\mbs{\Sigma}}}$ is the data covariance and
            $\prior(\gamma, \nth{\mbs{m}}, \phi_1)$ encodes both the present-day
            mass function of the stream, as well as any observational
            constraints (e.g., $g < 20$ mag).

            For a given star, we are interested in its probability of belonging
            to the stream's isochrone. Therefore, we marginalize over $\gamma$
            to find:
            \begin{small}
            \begin{align} \label{eq:photometric_probability}
                &\cmp{S,m}{\pdf}(\nth{\mbs{m}} | \mbs{\theta}(\phi_1)) = \\
                &\ \int\limits_{\gamma=0}^{1} \mcal{N}(\nth{\mbs{m}} | \{\mcal{I(\gamma)} + \mu(\phi_1), \mbs{\Sigma}_{\mcal{I}}^2(\gamma) + \mathds{1}\sigma_\mu^2(\phi_1) + {\cmp{m}{\nth{\mbs{\Sigma}}}}^2 \}) \nonumber
                \\
                &\qquad \times \prior(\gamma, \nth{\mbs{m}} | \phi_1) \rm{d}\gamma. \nonumber
            \end{align}
            \end{small}
            Then, for a given $\phi_1$ we can obtain a distribution over the
            distance modulus using 
            \begin{align}
                \mu_{\rm sample}\left(\phi_1\right) \sim \mathcal{N}\left(\mu(\phi_1), \sigma^2_\mu(\phi_1) \right). %& \quad\unit{mag$^{-1}$}.
            \end{align}
            For each sample, we can convert the distance-modulus distribution
            (and error) to the distance track using 
            \begin{align}
                \rm{dist}\left(\phi_1\right) &= 10^{\frac{1}{5}\mu_{\rm sample}(\phi_1)+1} & \quad\unit{pc}, \\
                \delta \rm{dist}(\phi_1) &\simeq \frac{\ln{10}}{5} \rm{dist}(\phi_1) \ \sigma_\mu(\phi_1) & \quad\unit{pc}.
            \end{align}

            Let $\cmp{S,m}{\mbs{\theta}}(\phi_1)$ represent the feed-forward
            neural network which outputs a vector of parameters characterizing
            the stream's photometric distribution at a given $\phi_1$. The
            network has 2 outputs -- $\mu, \sigma_\mu$ -- as we only consider
            the distance track of the isochrone, holding fixed stellar
            population parameters internal to modeling the isochrone, such as
            its age and metallicity.  In principle we might include parameters
            of the isochrone in the model, now $\mcal{I}(\mbs{\theta}, \gamma)$,
            where $\theta = {a, Z, ...}$ and any other generating parameters of
            the isochrone.  However, at a detailed level stellar streams are
            generally poorly fit by theoretical isochrones, at least ones with
            realistic ages and metallicities.  Consequently including the age
            and metallicity as model parameters does not meaningfully contribute
            to constraining results on the stream's stellar population.
            Moreover, age and metallicity are partially degenerate with the
            distance modulus, working counter to the goal of fitting a distance
            track to the stream. An alternate approach is to use data-driven
            isochrones instead of theoretic ones.  We defer this to a future
            work.

            The photometric probability, \autoref{eq:photometric_probability},
            encodes the photometric distribution -- the ``track" -- of the
            isochrone model at each point along the track, as well as the
            relative probability of points along the track. This is the term
            $\prior(\gamma, \nth{\mbs{m}} | \phi_1)$, which describes all priors
            on both the isochrone as well as survey constraints and selection
            effects. We assume that the properties of the isochrone are
            conditionally independent from observational constraints; namely
            \begin{equation} \label{eq:isochrone_prior_components}
                \prior(\gamma, \nth{\mbs{m}} | \phi_1) = \prior_{\cal{I}}(\gamma | \phi_1) \prior_{obs}(\nth{\mbs{m}} | \phi_1)
            \end{equation}
            where $\prior_{\mcal{I}}$ and $\prior_{obs}$ are the isochrone and
            observational priors, respectively.

            The second term in \autoref{eq:isochrone_prior_components} --
            $\prior_{obs}(\nth{\mbs{m}} | \phi_1)$ -- encodes all observational
            effects in the photometric dataset. Some effects are easily modeled,
            e.g. instruments have well-characterized magnitude limits, both
            faint and bright. Other effects are not easily modeled, e.g, complex
            location-dependent completeness issues \citep{GaiaCompleteness}.
            Accounting for the completeness is very challenging. In practice we
            avoid the contribution of completeness to $\prior_{obs}$, masking
            data for which it is relevant. \Gaia, for instance, is complete from
            $12 < G < 17$ and has high completeness up to $G \approx 20$, thus
            we mask $G > 20$.  The masks are easily modeled as magnitude limits
            and replace the complex completeness model in $\prior_{obs}$.
            \autoref{sub:method:missing_data} explains how masking is
            implemented in the models, though in this case every feature is
            masked, not only a few feature dimensions. With every featured
            masked, the photometric model in
            \autoref{eq:photometric_probability} reduces from an
            $F_m$-dimensional Gaussian to a 0-D Gaussian, which is a null
            distribution and does not enter into the likelihood.  In photometric
            coordinates completeness depends on a stars apparent magnitude;
            viewed in astrometric coordinates, the completeness cannot be
            described so simply. \Gaia, for instance has scanning law residuals
            in the density field. When photometric coordinates are masked due to
            completeness, we do not mask the corresponding  astrometric
            features, so that the stars membership probability is determined
            solely by the astrometric model. This practice allows us to include
            stars with good astrometric measurements, not penalizing those stars
            for which the photometric model is insufficient.  We examine the
            results to confirm that the scan pattern does not impact the
            membership distribution.

            The first term in \autoref{eq:isochrone_prior_components} --
            $\prior_{\mcal{I}}$ -- encodes the point-wise amplitude of the
            isochrone.  Physically, this amplitude is the the present-day mass
            function (PDMF). The PDMF $\prior_{\cal{I}}(\gamma, \phi_1)$ is a
            joint distribution, depending on $\gamma$ since it is the
            $\gamma$-dependent distribution amplitude, and also on $\phi_1$.
            This latter dependence is confusingly termed the ``mass function of
            the stream", which is the expected variation in the stellar mass
            distrubion along the stream, arising due to e.g. mass segregation in
            the progenitor \citep{WebbBovy2022}. To date this variation along
            the stream has not been observed, so in practice we drop the
            $\phi_1$ dependence -- $\prior_{\cal{I}}(\gamma, \phi_1) =
            \prior_{\cal{I}}(\gamma)$ -- assuming that the PDMF is constant
            along the stream.

            In the \href{https://github.com/GalOrrery/stream_ml-pytorch}{code}
            we offer a few ansatz for the PDMF, though any user-defined function
            may be used instead. The default assumption, known to be
            non-physical, is a uniform distribution over $\gamma$.  Similarly
            non-physical are truncated uniform distributions, considering only a
            segment of the isochrone, e.g. the main sequence turnoff (MSTO). We
            include also a Kroupa initial mass function (IMF)
            \citep{Kroupa2001}. The Kroupa IMF, and all physically-motivated
            IMFs, have a much larger fraction of low-mass stars than high-mass
            stars, the latter being less likely to form.  This is observed in
            streams, particularly their progenitors. However the Kroupa IMF, and
            most out-of-the-box IMFs, are poor fits to the observed PDMF as
            stream progenitors like globular clusters are extremely old
            ($\sim10$ Gyr) and may be dynamically evolved
            \citep{GrillmairSmith2001} while an IMF is the zero-age
            distribution.  How then should the PDMF $\prior_{\cal{I}}$ be
            treated?  One approach to the PDMF is to use a more informed
            distribution, for example time-evolving an IMF, either analytically
            or by simulation, to the age of the stream's stellar population. %As
            %discussed previously, globular clusters are known to be poorly fit
            %by theoretical isochrones with realistic inputs (e.g. age, $Z$).
            %Thus fitting an isochrone's track to the data means its age,
            %metallicity and therefore PDMF are %not well fit \TODO{[cite]}.
            However, for many streams, e.g. those lacking known progenitors,
            the PDMF is observationally poorly constrained.  In short, using any
            fixed distribution is simple in implementation, however the choice
            of distribution is not.

            Rather than using a fixed \textit{a priori} ansatz, another approach
            to the PDMF is to incorporate it into the model, parameterized by
            some distribution and fit simultaneously to the distance modulus and
            other $\cmp{m}{\mbs{\theta}}$.  We find this, like incorporating $a,
            Z$ into the isochrone model, to be interesting avenues by which to
            develop the model. However the resulting significant increase in
            model flexibility necessitates careful regulation, and we leave this
            to a future work.

            The last approach to the issue of the PDMF, and the one we take in
            this work, is to render it unimportant.  With fixed age,
            metallicity, and other isochrone model parameters, the primary
            purpose of the isochrone is to allow the model to determine the
            distance modulus of the stream. For this, only a portion of the
            stream need be modeled.  For streams with observable main sequence
            turn-offs, like \stream{GD-1} in \Gaia, this portion of the
            isochrone has both many stars and a relatively small range in
            masses. Thus, we restrict the mass range of the isochrone such that
            we are modeling only a region of interest over which the PDMF does
            not meaningfully contribute to the final result.

        % ssub: on_stream (end)

        \subsubsection{Linking to Astrometrics: the Distance Modulus and the Parallax} \label{ssub:method:linking_to_astrometrics}

            % TODO: rephrase
            By modeling jointly the astrometrics and photometrics we may tie
            together components of the models.  In particular, the distance
            modulus in the photometric model and the parallax in the astrometric
            model are both measures of the distance track of the stream.  We
            introduce a prior to connect these components of the model:
            \begin{align}
                \prior(\{\mu_I, \mu_\parallax \} | \phi_1) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1), \label{eq:distance_track_prior}
            \end{align}
            where $\delta$ is the Dirac delta function \citep{Dirac1947}, and
            enforces that the distance modulus track and parallax track must be
            equal (when converted to the actual distance) as a function of
            $\phi_1$.  The widths are similarly constrained. The distance and
            first-order error conversion are given by
            \begin{align}
                \mu_I(\parallax) &= 10 - 5\log_{10}(\parallax) \\
                \sigma_I(\sigma_\parallax)  &\approxeq \frac{5}{\ln{10}} \mu_I \lvert\sigma_\parallax\rvert.
            \end{align}

            In practice, at the distance of many streams the errors in the
            parallaxes are of order 1 and we exclude the feature from the model.

        % ssub: linking_to_astrometrics (end)

        \subsubsection{Off-stream}
        \label{sub:method:photometric_model:off_stream}

            The distribution of background stars in color (i.e., $g-r$) and
            magnitude (i.e., $g$) is a complicated function of location in the
            galaxy. Even within a field centered on the
            stream of interest, the color-magnitude diagram (CMD) is a
            combination of stellar populations distributed over a range in
            distances. Because of this, we do not find an analytic density that
            describes the data and use instead a normalizing flow, discussed in
            \autoref{sub:method:pre-training_distributions}.  Because the
            background CMD can evolve with $\phi_1$, we fit a \emph{conditional}
            normalizing flow, $q_{\rm flow}(\mathbf{m}|\phi_1)$, so that the
            color-magnitude diagram is conditionally dependent on position along
            the stream. In practice, we model magnitudes and not colors (i.e.,
            $\mathbf{m} = (g,r)$ rather than $g \ \mathrm{and} \ g-r$.
            Otherwise, the two arguments will always be covariate). The
            normalizing flow approach makes modeling the photometry relatively
            straightforward.

            In photometry the intrinsic width of stellar streams tends to be
            quite small, with correspondingly small $\sigma_\mu$ (see
            \autoref{eq:photometric_probability}). Therefore the effective
            region of the isochrone model is well determined and we do not find
            the marginal log-likelihood approach discussed in
            \autoref{sub:method:pre-training_distributions} to be impactful.

        % ssub: off_stream (end)

    % sub: photometric_model (end)

    %\newpage
    \subsection{Priors} \label{sub:methods:priors}

        In \autoref{sub:method:likelihood_setup} we introduced the likelihood
        setup. In \autoref{ssub:method:photometric_model:on_stream} we added
        priors in the form of the isochrone's PDMF and photometric observational
        constraints. In this section we introduce and develop more priors that
        are important for training our Bayesian Mixture Density Network's
        parameters' neural network(s).

        \subsubsection{Parameter Bounds} \label{ssub:methods:priors:parameter_bounds}

            For each parameter in $\mbs{\theta}$ of the model we implement
            priors on its range.  Uninformative, e.g. uniform $\mcal{U}(a, b)$,
            priors are often assumed where little is known about a parameter
            (except that $(a,b]$ contain the correct $\theta$ and its
            distribution). Where more is known about $\theta$, for instance that
            it is a categorical, more informative priors like the Dirichlet
            distribution \citep{Christopher2016} may be used.  Our code permits
            any user-defined prior, and we include $\mcal{U}(a, b)$.  The
            default choice of prior and bounds $\mcal{U}$ set to the limits of
            the dataset in each feature dimension.

            The choice of prior is important, not only for the information it
            adds to the model, but also for how optimizers explore the parameter
            space. Simple optimizers, like MCMC, propose a parameter vector then
            compute its likelihood given the data
            \citep{HoggForeman-Mackey2018}. Priors with compact support  $\theta
            \sim \mcal{D}((a, b])$, e.g.  $\mcal{U}(a, b)$, may be very
            dangerous as the optimizer can be initialized in a region with zero
            probability and has no means but blind luck to find the region of
            support -- the region with non-zero likelihood. More sophisticated
            optimizers, like gradient-based descent \citep{cauchy_2009}, are
            susceptible to a similar initialization problem, allowing
            zero-probability regions, but requiring at least a non-zero gradient
            to make informed updates to the optimizers state.  Outside the
            support the gradient can be zero.  With MDNs, model parameters are
            neural network outputs. Unlike for conventional models where an
            informed initial parameter state may be chosen by hand, it is
            impractical and undesirable to initialize by hand a neural network's
            output by setting all its internal weights. One might imagine
            accomplishing a desired initial output by pre-training the network
            with a toy model that forces convergence on a set output vector.
            However, neural networks are highly non-linear in their response to
            internal state changes (a desirable property). It is quite possible
            that when switching from the toy to real model's loss function the
            pre-trained network is in a local probability island from which the
            global optimum is unreachable.  
            Therefore, with the MDN neural networks, initialization is mostly
            for checking the dependence of the results on the initial state --
            though even this dependence may instead be done using parameter
            dropout. Rather than focusing on initializing to a particular
            parameter vector, for neural networks it is instead important that
            the initial state, whatever it is, has non-zero probability (or
            non-zero gradient, depending on the optimizer).  Neural networks
            have some intrinsic phase-space range that depends on the
            architecture of the network. Therefore the network architecture must
            be chosen such that the outputs match well with the priors. For
            example, with a uniform prior $\theta \sim \mcal{U}(a,b)$, the
            network should be restricted such that $\theta \in (a, b]$, e.g. by
            ending the network's $\theta$ channel with a scaled logistic
            function.
            \begin{equation}\label{eq:scaled_logistic}
                  \sigma(x, \{a,b\}) = a + (b - a) * \sigma(x)
            \end{equation}
            Therefore, we pair every parameter bound with a matching bound in
            the network architecture.

            The parameter bounds may be independent of $\phi_1$, as is implicit
            to the above discussion, or conditionally dependent and vary as any
            user-defined function of $\phi_1$.  Alternatively, prior parameters
            may be separated into pieces, dealing with different modeling needs,
            e.g. defining broad parameter bounds as was done in this section.
            We introduce a different piece subsequently.

        % ssub:methods:priors:parameter_bounds (end)

        \subsubsection{Guides Prior} \label{sub:methods:priors:track_region_prior}

            In \autoref{ssub:method:astrometric_model:on_stream} the stream
            model's parameters were unconstrained. Bounds on the parameters and
            the neural network architecture were introduced in
            \autoref{ssub:methods:priors:parameter_bounds} to ensure the initial
            state of the model and network were compatible with the optimizer.
            Now we provide known information about physical systems,
            particularly the track of the stream.  Our density modeling is aimed
            at characterization of \textit{known} streams, not stream discovery.
            From previous studies, the broad track of a stream is known for
            significant portions of its length \citep[e.g. see the atlas
            in][]{Mateu2022}. It makes sense, then, to provide this information
            to the model. Care must be taken that this prior information does
            not dominate the model.  We introduce a region prior to encourage
            the stream model to pass through a user-defined guide region, while
            also remaining compatible with the data.  We refer to the prior as a
            ``split"-Normal prior: a Gaussian split and separated at the peak
            and pieced back together with a flat region connecting the halves.
            The un-normalized PDF is given by:
            \begin{equation}
                \pdf(\theta,\mu,\sigma,w) \propto \begin{cases} 
                   \Exp{-\frac{1}{2}\left(\frac{x-(\mu-w)}{\sigma}\right)^2} & \phantom{\mu - w <}\ x < \mu - w \\
                    1 & \mu - w \leq x < \mu + w \\
                    \Exp{-\frac{1}{2}\left(\frac{x-(\mu+w)}{\sigma}\right)^2} & \mu + w < x
                \end{cases}.
            \end{equation}
            The piecewise PDF is smooth up to the first derivative and can be
            used with gradient-based optimizers.

            The Gaussian portions of the prior encourage the network towards the
            central region $\mu - w < x < \mu + w$. Since the prior is
            equi-probable in the guide, the stream may lie anywhere as preferred
            by the data and the stream model.  It is convenient to
            reparametrize, with $\tau = \frac{1}{\sigma}$. Namely,
            % \begin{small}
            \begin{equation}
                \ln \pdf(\theta,\mu,w) \propto \tau
                \begin{cases} 
                    -\left(x-(\mu-w)\right)^2 & \phantom{\mu - w <}\ x < \mu - w \\
                    0 & \mu - w \leq x < \mu + w \\
                    -\left(x-(\mu+w)\right)^2 & \mu + w \leq x
                \end{cases},
            \end{equation}
            % \end{small}
            where the additive normalization is
            \\
            $-\ln\left(\sqrt{\frac{\pi}{\tau}}
            \left(\frac{1}{2}\text{erf}\left(2 \sqrt{\tau }
            w\right)+1\right)\right)$.  In this form we see that $\tau$ may be
            increased to strengthen the prior and encourage $\theta$ to lie
            within the desired region. It is important that the region be set
            large enough that we may be confident the prior is not driving the
            final result on small scales. As an added measure, it is often
            useful to turn off this prior after the stream lies within the
            desired guides and the learning rate of the optimizer is low enough
            that the model does not wander out.

        % sub:methods:priors:track_region_prior (end)

    % sub:methods:priors (end)

    \vspace{5pt}
    \subsection{Model Implementation, aka the code} \label{sub:methods:model_implementation}

        In prior sections we discussed the mathematical implementation of each
        component of the Mixture Density Network. Here we make brief remarks on
        the code implementation and architecture of the neural networks.

        The neural networks are built with the \package{PyTorch} stack
        \citep{Pytorch2019}. We wrap the network \texttt{pytorch.Module} in a
        custom framework, allowing the Bayesian likelihood, prior, evidence,
        etc. to be bundled with the network's forward step. This framework
        allows for a high degree of modularity. Component models, e.g. the
        stream model, of the mixture are easily added to (and moved from) as
        items in the larger mixture model. Probability methods understand how to
        traverse the mixture model, calling and combining the appropriate
        methods of the constituent models, so it unnecessary to write custom
        loss functions.  Another benefit of the modular framework is the ease of
        building and experimenting with different mixture models. Furthermore,
        each model component may be saved and serialized separately from the
        other components, allowing for portability of the components between
        models.  For example, if a stream has a prominent extra-tidal feature,
        e.g. the spur of \stream{GD-1}, a fiducial model of just the thin stream
        may be trained, then the stream (and background) model components may be
        saved and used as the initialization to the stream (and background) of a
        more complete model that includes the spur. This modularity also has a
        cost: by allowing each model component (including the categorical
        mixture model itself) to contain its own network we decrease the
        cross-talk between nodes in separated networks. The networks communicate
        only through the backwards propagation of the loss function, increasing
        the required training time of the models. We note that the code
        framework, being built on \package{PyTorch}, easily permits a single
        model with single network. Having found a good model by experimenting
        with the modular layout, the model may be reproduced and re-trained as a
        single network. Alternatively, though we leave this to a future version
        of the code, it is possible to ``fuse" the mixture, providing one
        network whose outputs are apportioned appropriately to each constituent
        model.

        The models permit user-defined priors to be passed at model
        instantiation.  Priors specific to a single model are contained on that
        model, while priors connecting two (or more) models are held on the
        mixture model object. As an example of the latter, see
        \autoref{ssub:method:linking_to_astrometrics}, which connects the
        distance component of the astrometric and photometric stream models.

        The custom model framework combines for each model a neural network with
        Bayesian probability functions. The latter are set by the model type,
        e.g. the Gaussian model has a Gaussian likelihood function, but the
        former, the neural network, is set and determined by the user. So long
        as the network has the correct number of inputs (e.g. one for $\phi_1$)
        and outputs (one per parameter), any feed-forward network may be used.
        Our default choice, and for which we provide a helper function, is an
        MLP with a sequence of \texttt{Linear} \!\!\textrightarrow \texttt{Tanh}
        \!\!\textrightarrow \texttt{Dropout} (hereafter \texttt{LTD}) blocks.
        Dropout layers are used to prevent network co-adaptation and
        over-fitting. Dropout works by randomly zeroing out nodes in the
        network, decreasing the importance of any single node, encouraging
        instead emergent patterns across the network \citep{GalGhahramani2015}.
        One model for which the \texttt{LTD} network cannot be used is the
        normalizing flow used in the photometric background model
        (\autoref{sub:method:photometric_model:off_stream}). Here we use the
        \package{zuko} package, with a conditional diagonal normal base
        distribution and composite transform network of a reverse permutation
        and masked affine auto-regressive transform layers. As noted in our data
        availability statement, all the codes used in this paper are bundled by
        \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}}
        in a public repository. We will describe for each stream model specifics
        relevant to that stream but refer readers to the publicly-available code
        itself for code implementation details.

% sec:method (end)

\section{Results: Mock} \label{sec:results_mock}

    In this section we present an application of our model to a synthetic stream
    with known ground truths, and demonstrate the model's ability to
    characterize stream density variations and membership probabilities in an
    unsupervised manner.

    \subsection{Data Simulation} \label{sub:results_mock:data}

        We generate a simple and illustrative synthetic stream observations by
        sampling from simple algebraic forms.  In $\phi_1$ we sample from a
        uniform distribution. Non-Poissonian density variations, like gaps, are
        introduced by adding to the uniform distribution two Gaussian
        distributions centered at different values of $\phi_1$ with negative
        amplitudes and random widths. Stream stars are then sampled from this
        uniform-plus-gaps mixture distribution.  In $\phi_2$ we sample from a
        quadratic in $\phi_1$ with normally-distributed $0.15^{\degree}$
        scatter.  The distances are uniformly sampled from $7$ to $15$ kpc in a
        linear function in $\phi_1$ with normally-distributed $0.25 \ \rm{kpc}$
        scatter, then transformed to the parallax $\parallax$. Even though the
        stream width is constant in the distance the width in parallax is not,
        which will be used to demonstrate that we can recover $\phi_1$ dependent
        distributions.  The number of stream stars is
        $\variable{output/mock/nstream_variable.txt}$.
        $\variable{output/mock/nbackground_variable.txt}$ background astrometric
        points are generated by sampling from uniform distributions in each
        coordinate.

        The synthetic stream's photometry is also simulated using a
        \variable{output/mock/isochrone_age_variable.txt}\unskip, [Fe/H] =
        \variable{output/mock/isochrone_feh_variable.txt} MIST isochrone
        \citep{Dotter2016, Choi+2016,brutus}. The isochrone is truncated just
        short of the giant branch, and is sampled from assuming a stream mass
        function similar to \stream{Pal\,5} \citep{GrillmairSmith2001} as well
        as with an intrinsic uniform 0.2~dex scatter orthogonal to the isochrone
        track.

        The isochrone is shifted by the stream's distance modulus, which is
        derived from the astrometric distance track.  The background photometric
        coordinates are generated as a 2-dimensional Gaussian with non-zero
        covariance.  Thus, the synthetic data scenario described here, and
        presented in \autoref{fig:mock_data_photometric_background_selection},
        generates a mock stream in astrometric coordinates and photometric
        coordinates, with density variations, a distance gradient which is
        consistent in parallax and distance modulus (i.e., magnitudes), all
        superimposed over a noisy field of background points in each astrometric
        and photometric dimension.

    \subsection{Model Specification} \label{sub:results_mock:model}

        \begin{figure}
            \script{mock/plot/photometric_background_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/mock/photometric_background_selection.pdf}
            \caption{
                %
                \textbf{Top panel:} Astrometric coordinates ($\phi_1$, $\phi_2$)
                for the mock data set.  Light blue points are the stream and
                background within an ``on"-stream region. Black points are in
                the ``off"-stream region used to train the background
                photometric model.
                %
                \textbf{Bottom panel:} Photometric coordinates ($g-r$, $g$) for
                the mock data set.  Black points are in the ``off"-stream region
                from the region defined in the astrometric coordinates.  Light
                blue points are the stream and background from the on-stream
                region. The stream's isochrone is apparent as an overdensity
                around (g-r, g)$\ \approx (0, 20)$.
                %
            }
            \label{fig:mock_data_photometric_background_selection}
        \end{figure}

        \begin{figure*}[!h]
            \script{mock/plot/results.py}
            \centering
            \includegraphics[width=0.9\linewidth]{figures/mock/results.pdf}
            \caption{
                %
                \textbf{Top panel: }%
                    The stream mixture coefficient $f(\phi_1)$ (black dotted
                    line) predicted by the model, overplotted on the true stream
                    fraction (black-outlined histogram), calculated as the ratio
                    of stream stars to total stars in 55 bins over the $\phi_1$
                    range of the isochrone.
                %
                \textbf{Middle panel: }
                    $\phi_2(\phi_1)$ over the full range of $\phi_1$. The data
                    color and transparency is set by the model's MLE stream
                    probability (with dropout turned off). Plotted behind the
                    data are the true labels for the stream stars (large black
                    circles).  The true labels exactly match the
                    model-identified stream stars.  The MLE stream track (black
                    dashed line) is over-plotted on the probability-colored
                    data.
                %
                \textbf{Lower panel 3: }%
                    $\parallax(\phi_1)$ over the full range of $\phi_1$. Like for
                    $\phi_2$, the data color and transparency is set by the MLE
                    probability, with the true labels (large black circles)
                    under-plotted and MLE track (black dashed line)
                    over-plotted.
                %
                \textbf{Upper Histograms: }%
                    $\phi_2$ histograms for each model component (blue
                    background, red stream), in 4 $\phi_1$ ranges across the
                    data set.  The height of each bin is the
                    probability-weighted density.  The background is a uniform
                    distribution, while the stream is a Gaussian. In projection
                    across the $\phi_1$ range it can appear multi-modal (panel
                    2) if there are large gaps or `smeared' out (panel 3) if the
                    Gaussian's mean changes significantly over the range.
                %
                \textbf{Upper Middle Histograms: }%
                    Parallax $\parallax$ histograms for each model component (blue
                    background, red stream), in 4 $\phi_1$ ranges across the
                    data set.  The height of each bin is the
                    probability-weighted density.
                %
                \textbf{Lower Middle and Lower Histograms:}%
                    Photometric coordinate $(g, r)$ plots for each model
                    component in 4 $\phi_1$ ranges across the data set. The top
                    row plots the stream stars (red) over the background. The
                    lower row plots the background stars (blue) over the stream
                    stars to show  how the two distributions mix.
                %
            }
            \label{fig:mock_data_result}
        \end{figure*}

        We select on-stream and off-stream regions using the known $\phi_1,
        \phi_2$ range of the mock stream.  The upper panel of
        \autoref{fig:mock_data_photometric_background_selection} shows these
        selections in blue and black, respectively.  The lower panel shows the
        on and off-stream selections in photometric coordinates.  In this space
        it is apparent that the on-stream region has two populations, drawn from
        the isochrone and from background points included in the simple $\phi_1,
        \phi_2$ box.  In contrast, the off-stream region has only the the
        background population.  We use the off-stream population to train a
        conditional normalizing flow (see
        \autoref{sub:method:pre-training_distributions}) on the magnitudes $g$
        and $r$ and conditioned on $\phi_1$.  The normalizing flow is now a
        non-parametric distribution characterizing the photometric background,
        including its variation over $\phi_1$.  For this mock stream the
        flexibility of the normalizing flow greatly exceeds necessity: the
        background is drawn from a $\phi_1$-invariant multivariate normal,
        essentially the base distribution of the normalizing flow.  It is
        therefore no surprise that the trained normalizing flow provides an
        excellent characterization of the background.  We will later see that
        the flow performs similarly well on complex distributions.  Having
        trained the background photometric model we proceed to build and train
        the full model.

        Per \autoref{eq:stream_and_bkg_prob} the stream is an independent
        addition of both astrometric and photometric models.  In astrometrics
        the stream is a Gaussian track in $\phi_2$ and parallax $\parallax$,
        using the model defined in
        \autoref{ssub:method:astrometric_model:on_stream}.  We use a 4-layer
        \texttt{LTD} with overall shape ($\phi_1$ input, 4 outputs), where the 4
        outputs are $[\mu_{\phi_2}, \sigma_{\phi_2}, \mu_{\parallax},
        \sigma_{\parallax}](\phi_1)$.  The network is small, but sufficiently
        large to permit 15\% dropout and retain robust predictions.  The
        photometric model is a distance modulus-shifted isochrone, using the
        model outlined in  \autoref{ssub:method:photometric_model:on_stream}. We
        use the same isochrone parameters as were used to simulate the stream,
        including the \citep{GrillmairSmith2001} mass function.  For this mock
        stream, we demonstrate simultaneous astrometric and photometric
        modeling, linking the distance modulus and parallax track, as explained
        in \autoref{ssub:method:linking_to_astrometrics}. Thus the photometric
        model does not have its own network, using instead the outputs
        $[\mu_{\parallax}, \sigma_{\parallax}](\phi_1)$.

        The total background model has two pieces, an astrometric and a
        photometric model.  As a pre-training phase the photometric normalizing
        flow model was fitted to an off-stream selection and is now held fixed.
        The astrometric model is split in two components, the first is a uniform
        distribution in $\phi_2$, and the second is an exponential distribution
        for $\parallax$.  The uniform distribution does not have any parameters,
        and thus no neural network.  The exponential distribution has only the
        slope parameter and we use a 3-layer \texttt{LTD} with shape (1, 32, 1)
        and 15\% dropout. Background and stream are combined in a mixture model
        whose mixture parameters are a 4-layer (1, 64, 1) \texttt{LTD} network.
        We train the mixture model with an \texttt{AdamW} optimizer with
        learning rate of $5\times10^{-3}$ and a scheduler to periodically and
        temporarily increase the learning rate, encouraging the optimizer
        towards better minima.

    \subsection{Trained Density Model} \label{sub:results_mock:results}

% sec: results_gd1 ---
    \begin{figure*}[ht]
        \script{gd1/plot/data_selection.py}
        \centering
        \includegraphics[width=1\linewidth]{figures/gd1/data_selection.pdf}
        \caption{%
            Data selections for \stream{GD-1}, within the data cube described in
            \autoref{sub:results_gd1:data}.
            %
            \textbf{Row 1 Panels:}
            Proper motion and photometric selections.  \stream{GD-1} is apparent
            in the proper motions as a small overdensity. The photometric
            selection is based on a 12 Gyr, [Fe/H] = -1.2 MIST isochrone,
            buffered by 0.3 mag to select .
            %
            \textbf{Rows 2 \& 3 Panels:}
            Applying the combination of astrometric and photometric selections,
            plotted for astrometric coordinates $\phi_2, \parallax,
            \mu_{\phi_1}, \mu_{\phi_2}$.  The proper motions are plotted
            relative to \citetalias{Ibata+2021}. Using a log-density coloring,
            the stream is identifiable in all phase-space coordinates except
            $\parallax$ where the errors are significant.  }
        \label{fig:gd1-data_selection}
    \end{figure*}
% --- sec: results_gd1 (end)

        The result of applying our density model to the synthetic data is
        illustrated in \autoref{fig:mock_data_result}. The top panel shows the
        weight parameter that controls membership probability as a function of
        $\phi_1$. The dashed curve represents the output of our data-driven
        model, while the solid curve is the ground truth. The second panel shows
        the stream and background in the $(\phi_1,\phi_2)$ plane, and the third
        panel shows the stream and background in the $(\phi_1,\parallax)$ plane.
        Points are color-coded according to their membership probability
        \autoref{eq:membership_prob}. The weight parameter clearly shows that
        the model has captured the two prominent density variations along the
        stream.  There are additional variations, mostly around the edges of the
        stream, though re-sampling the neural network weights with dropout
        reveals that these variations represent regions of higher model
        uncertainty.

        We find that the stream is successfully recovered in both astrometric
        and photometric coordinates, with
        $\variable{output/mock/ncorrect_variable.txt}\%$ of stream stars being
        recovered with membership probability greater than $80\%$.  In the
        bottom two rows of \autoref{fig:mock_data_result} we illustrate the
        synthetic data color-coded by stream membership probability (top) and
        background membership probability (background) for different $\phi_1$
        slices. Importantly, the photometric portion of the mixture model
        successfully identifies the stream in magnitude space, showing a high
        stream membership probability where the stream is present, and a low
        probability ($\sim 0$) where the stream is absent.  The false positive
        rate (i.e., stars misidentified as stream members) for this test is
        found to be $\variable{output/mock/falseident_variable.txt}\%$ (3 stars)
        when cutting on the 80\% membership probability threshold.

% sec: results  (end)

\section{Results: GD-1} \label{sec:results_gd1}

    We now apply our method for stream characterization to the stellar stream
    \stream{GD-1}, initially discovered by \citet{GrillmairDionatos2006}.
    Previous work has identified several density variations along \stream{GD-1}
    \citep{deBoer+2018}, including a bifurcation that can be modeled by an
    encounter with a dense dark matter subhalo \citep{Bonaca+2019, Webb+2019}.
    Characterizing the detailed density fluctuations of this stream is therefore
    important for constraining a population of perturbers.  We discuss the data
    selection in \autoref{sub:results_gd1:data}, model specifics in
    \autoref{sub:results_gd1:model}, and the results in
    \autoref{sub:results_gd1:results}.

    \subsection{Data Selection} \label{sub:results_gd1:data}

        We utilize data from \Gaia{} DR3 for the astrometric coordinates of each
        star.  To obtain more accurate color information, we cross-match the
        \Gaia{} field on \PanStarrs{} \citep{PanSTARRS1} using \Gaia's provided
        \texttt{Best-neighbors} catalog.  To limit the field to a region known
        to contain \stream{GD-1} we adopt broad cuts on the cross-matched data.
        We perform these rough cuts by requiring the data lie within the
        phase-space cube:
        \begin{itemize}
            \setlength\itemsep{0em}
            \item $\phi_1 \in [-100, 40] \ \rm{deg}$,
            \item $\phi_2 \in [-9, 5] \ \rm{deg}$,
            \item $\parallax \in [-10, 1] \ \rm{mas}$,
            \item $G_{\rm BP} \in [-1, 3] \ \rm{mag}$
            \item $g, r \in [0, 50] \ \rm{mag}$, %, and not \texttt{null},
        \end{itemize}
        where we use \package{gala}'s \citep{gala, galav1.3} \code{GD1Koposov10}
        frame \citep{Koposov+2010} to define $\phi_1$, $\phi_2$, and associated
        proper motions.  For downloading purposes we split the query into 32
        sub-regions in even $\phi_1$ increments.
        A plot of the $\phi_1, \phi_2$ regions is included in the
        \href{\GitHubURL{}}{\package{ShowYourWork}} repository.  Stream-specific
        coordinates ($\phi_1, \phi_2$) are not included in \Gaia.  One could
        embed the coordinate transform in the ADQL \citep{ADQL2.0}, or define
        great-circle arcs in ICRS \citep{ICRS1997}. Since only rough cuts are
        required and since the sub-region boxes are sufficiently small for a
        reasonable flat-sky approximation  we simply translate the ($\phi_1,
        \phi_2$) boxes to ICRS.  We correct the \PanStarrs{} photometry for dust
        extinction using the \textit{Bayestar19} \citep{Green+2019} models in
        the \package{dustmaps} package \citep{Green2018}.  \textit{Bayestar19}
        is a three-dimensional dust map of the Milky Way for which we use
        \Gaia's position and parallax distance information. We note that the
        extinction has very little dependence on the distance past $\approx 8 \
        \rm{kpc}$. For stars with missing parallax measurements we assume a
        distance of $8.5 \ \rm{kpc}$, a typical distance to the stream, thus
        increasing potential membership probability. Adopting a slightly lower
        or higher fiducial distance does not significantly change the results of
        our analysis. We include the 1$\sigma$ errors as a source of uncertainty
        in $\mbs{\Sigma}_n$, the per-star photometric uncertainty.  We also
        correct the parallax for the zero-point using the
        \href{https://pypi.org/project/gaiadr3-zeropoint/}{\code{gaiadr3-zeropoint}}
        package, which implements the findings from \cite{Lindegren+2021}.  The
        result is a field populated by \variable{output/gd1/ndata_variable.txt}
        stars, from which we would like to identify stream members.

        We create data masks for a higher signal-to-noise selection around
        \stream{GD-1}.  The two mask types, proper motion and photometric, are
        shown in \autoref{fig:gd1-data_selection}. For the proper motions we use
        \citet{Ibata+2021} (hereafter \citetalias{Ibata+2021}) as the reference
        for the known track of the stream. We select all all stars with
        $\mu_{\phi_1}^*, \mu_{\phi_2}$ within 6 mas/yr of the track, as a
        function of $\phi_1$. In \autoref{fig:gd1-data_selection} the proper
        motion data are shown as the difference from the \citetalias{Ibata+2021}
        track.  In photometric coordinates we define a mask around a $12 \
        \rm{Gyr}$, [Fe/H] = $-1.2$ isochrone at a distance of $7.8 \ \rm{kpc}$.
        The selection region is a $0.3 \ \rm{mag}$ orthogonal buffer around the
        MIST isochrone \citep[using][]{brutus}, which is illustrated in the
        2nd-from-the-top-left plot in \autoref{fig:gd1-data_selection}. The
        buffered isochrone selection is wide enough to contain \stream{GD-1}
        over its full distance gradient.  In addition, to avoid detailed
        modeling of the completeness as a term in $\prior_{obs}$ (from
        \autoref{eq:isochrone_prior_components}) we mask out photometry with $G
        > 20$.  With combinations of these masks the stream has much larger
        signal-to-noise, though is still only a small percent of the total data.

    \subsection{Model Specification}\label{sub:results_gd1:model}

        \begin{figure}
            \script{gd1/plot/photometric_background_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/gd1/photometric_background_selection.pdf}
            \caption{%
                The on and off-stream selection.
                % 
                \textbf{Top panel:} Astrometric coordinates ($\phi_1$,
                $\phi_2$).  Light blue points are the stream and background
                within an on-stream region. Black points are in the off-stream
                region used to train the background photometric model. Note this
                is on a reduced field with a tighter proper motion cut (the
                stream visibly truncates at $-10 \ \rm{deg}$ to increase the
                visibility of the stream's CMD).  The proper motion cut is $-15
                < \mu_{\phi_1}^* < -10$, $-4.5 < \mu_{\phi_2} < -2$ [mas/yr].
                %
                \textbf{Bottom panel:} CMD pl got of the astrometric selection
                above. The stream's isochrone is apparent as an overdensity.
                %
            }
            \label{fig:gd1-photometric_background_selection}
        \end{figure}

        The modeling framework, detailed in section \autoref{sec:method}, is
        modular, allowing many model components to be combined, nested, added to
        the mixture, and linked via priors.  The model we require for
        \stream{GD-1} is similar to the model built for the mock stream in
        \autoref{sub:results_mock:model}, but has additional components for
        \stream{GD-1}'s spur.  In this section we will discuss how the total
        model is built, noting similarities to the example mocks-stream model.

        All full Bayesian stream models necessarily include a background model,
        which we model in both astrometric and photometric spaces.  The
        photometric background is too complex to capture with simple analytic
        distributions.  Instead, we train a normalizing flow on the magnitudes
        $g, r$, conditioned on $\phi_1$. We refer readers to
        \autoref{sub:method:pre-training_distributions} for discussion of the
        photometric distribution and how a normalizing flow can characterize
        this complex distribution, including its variation over $\phi_1$.  The
        flow's flexibility, while necessary in this context, also poses a
        problem: the flow captures the stream as easily as the background.
        Consequently we select on-stream and off-stream regions using the known
        $\phi_1,\phi_2$ range of the stream, as shown in the upper panel of
        \autoref{fig:gd1-photometric_background_selection}.  In the lower panel
        are plotted these selections in photometric coordinates, where an
        isochrone is apparent in the on-stream selection.  The normalizing flow
        is trained on the off-stream selection, and will be fixed (not
        trainable, nor contribute to the gradient) when incorporated into the
        larger model.

        Unlike for the photometry, the astrometric background could be (mostly)
        fit with simple analytic distributions.  This is true if the selected
        field is large enough -- the distribution in each coordinate is
        approximately a skew-normal or a similarly tailed exponential-type
        function. However by defining a higher signal-to-noise field around the
        \citet{Ibata+2021} stream track in $\mu_{\phi_1}^*, \mu_{\phi_2}$, the
        background in these coordinates is no longer well described by an
        exponential distribution. Similarly, for much of the extent of
        \stream{GD1's} field the parallax is well modeled by a truncated
        skew-normal distribution, $\parallax \sim \rm{TruncSkew}\mcal{N}(\mu,
        \Sigma; a, b)$, however for $\phi_1 \in [-100, -60]$ this does not hold.
        Rather than splitting the model at a somewhat arbitrary $\phi_1$ or
        interpolating between two distributions we instead fit the astrometry
        with a $\phi_1$-conditioned normalizing flow.

        \begin{itemize}
            \item $\phi_2, \sim \rm{TruncExp}(\lambda; a, b)$: a truncated
                exponential distribution (see
                \autoref{app:sub:exponential_distribution}).  The distribution
                has 1 slope parameters.
            \item $\mu_{\phi_1}^*, \mu_{\phi_2}, \varpi \sim q_{\rm flow}$: a 3
                feature, one context feature normalizing flow trained on an
                off-stream selection (see
                \autoref{fig:gd1-photometric_background_selection} for that
                selection). Though discussed in the context of a photometric
                background, \autoref{sub:method:pre-training_distributions}
                details how these flows are constructed and trained.
        \end{itemize}

        We model the main stream of \stream{GD-1} similarly to the mock stream
        -- as an independent addition of both astrometric and photometric
        models.  The astrometric model is a 5-dimensional Gaussian in a Gaussian
        track in $\phi_2, \parallax, \mu_{\phi_1}, \mu_{\phi_2}, \parallax$,
        using the model defined in
        \autoref{ssub:method:astrometric_model:on_stream}.  The parameter
        network's
        % is a 5-layer \texttt{LTD} with shape (1, 128, 10), where the
        10 outputs are $\mbs{\theta} = [\mu_{\phi_2}, \sigma_{\phi_2},
        \mu_\parallax, \sigma_\parallax, \mu_{\mu_{\phi_1}^*},
        \sigma_{\mu_{\phi_1}^*}, \mu_{\mu_{\phi_2}}, \sigma_{\mu_{\phi_2}},
        \mu_\parallax, \sigma_\parallax](\phi_1)$.  In testing the network is
        sufficiently large to permit 15\% dropout and retain robust predictions.

        The stream's photometric model is a distance modulus-shifted isochrone,
        using the model from \autoref{ssub:method:photometric_model:on_stream}.
        We use a $12 \ \rm{Gyr}$, [Fe/H]$ = -1.2$ MIST isochrone with no assumed
        intrinsic dispersion.  Using the method described in
        \autoref{ssub:method:linking_to_astrometrics} the distance modulus track
        is set by the parallax track in the astrometric model. Thus the
        photometric model has no independent parameters and consequently no
        neural network.

        \variable{output/gd1/control_points.tex}

        We model the spur of \stream{GD-1} as a separate component to the main
        stream. The spur has an identical model as the stream: a 5-D astrometric
        Gaussian and isochrone in the photometry. Owing to their common origin,
        we impose the prior that the spur and stream share the same isochrone.
        The parallax is also shared, motivated by initial findings, and that
        there were too few stars to robustly detect any deviations.

        \variable{output/gd1/member_table_select.tex}

        The parameter phase-space is large and a random initialization will take
        significant time to converge towards the stream track. This initial
        training is unnecessary as the rough stream track is known \textit{a
        priori}.  To guide the astrometric model towards the known
        \stream{GD-1}, we set down guides (see
        \autoref{sub:methods:priors:track_region_prior}) along \stream{GD-1}'s
        main track and on the spur.  We set the width of these guides to large
        values, so that their influence on the model is not dominant. Indeed, we
        find that the value of the guides is to help reduce training time, since
        otherwise it will take the flexible density model much longer to
        converge to the actual stream.  We include all the track regions in
        \autoref{tab:gd1_track_prior}.  We note that since the distance moduli
        between the stream and spur are linked, those stream track guides also
        apply.

        All the component models are tied together into a mixture model.  As an
        implementation detail, instead of grouping together the stream
        components in a hierarchical mixture model (background + (stream +
        spur)) we keep the structure flat, using a single mixture (background +
        stream + spur) and reducing the number of neural networks. The singular
        weight parameter network is a \texttt{LTD} network with 15\% dropout.
        The stream weights are in the range $\ln f \in [-10, 0]$, while the spur
        weights are $\ln f \in [-10, 0]$ and set to 0 $\phi_1 < -45, -15 <
        \phi_1$. By \autoref{eq:mixture_weight_normalization} the background
        brings the cumulative weight to one, normalizing the mixture.

    \subsection{Trained Density Model}\label{sub:results_gd1:results}

        \begin{figure*}[ht]
            \script{gd1/plot/results_full.py}
            \centering
            \includegraphics[width=0.83\linewidth]{figures/gd1/results_full.pdf}
            \caption{%
                The model for \stream{GD-1}, including both a thin-stream and
                spur component.  For comparison we include the tracks from
                \citet{Price-WhelanBonaca2018} and \citet{Ibata+2021}. Note how
                the model largely agrees with the latter track in all
                astrometric dimensions.
                %
                \textbf{Panel 1: }%
                    The stream mixture coefficients $f(\phi_1)$ predicted by the
                    model, colored by the stream and spur membership likelihood.
                %
                \textbf{Panel 2: }
                    $\phi_2(\phi_1)$ over the full range of $\phi_1$. The data
                    color and transparency is set by the model's 50th-percentile
                    stream and spur probability. We include the error bars for
                    stars with $>75\%$ membership probability. The predicted
                    50th-percentile track $\pm$ width is over-plotted (red
                    band), along with the 90\% confidence region (broader red
                    band).  In the adjoining top plot is shown the 5th to 95th
                    percentile variation in the width.
                %
                \textbf{Panel 3 \& 4: }%
                    $\parallax(\phi_1)$ and $d(\phi_1)$ over the full range of
                    $\phi_1$. The $\parallax$ is tied to the photometry -- see
                    \autoref{fig:gd1-results-panels}.
                %
                \textbf{Panel 5 \& 6: }%
                    The proper motions $\mu_{\phi_1^*}, \mu_{\phi_2}$. The model
                    convolves observational errors with the stream width, but
                    the large fractional errors make the determination
                    difficult. The track is consistent with 0 at $\mu_{\phi_2} <
                    -40 \ \rm{deg}$, but has noticeable variations in
                    $\mu_{\phi_1}$.  }
            \label{fig:gd1-results-full}
        \end{figure*}

        In this section we present results from our model applied to real {\it
        Gaia} and Pan-STARRS data of \stream{GD-1}.  We visualize the
        performance of our model in identifying stream stars in
        \autoref{fig:gd1-results-full}, and illustrate our photometric fit in
        \autoref{fig:gd1-results-panels}. In \autoref{tab:gd1_member_table} we
        provide a selection of \stream{GD-1} member stars. The full catalog is
        available online \TODO{add zenodo link when upload}. Lastly,
        \autoref{fig:gd1-heatmap} is a probability-weighted density heatmap of
        the \stream{GD-1} stream derived from our fit.

        In \autoref{fig:gd1-results-full}, the top panels shows the fraction of
        stars belonging to the stream (red) and spur (yellow) as a function of
        $\phi_1$. The second panel illustrates the stream in $\phi_1$, $\phi_2$
        coordinates, color-coded by the membership probability for the stream
        and spur components (derived from \autoref{eq:membership_prob}). The
        following rows represent the other astrometric dimensions for the stream
        and spur components, with the distance track obtained through our
        photometric model (visualized in \autoref{fig:gd1-results-panels} and
        discussed below). The red and yellow bands represent model predictions
        for in each panel for the main stream and spur, respectively. The points
        with thick error bars represent control points (discussed in
        \autoref{sub:results_gd1:model}) for the stream and spur components,
        which are priors on the location of the stream and spur components
        across the astrometric dimensions. The control points in the parallax
        panel represent the reliable parallax and distance modulus estimates
        from \citet{deBoer+2020}. The actual data points are color-coded by the
        maximum membership probability of belonging to any of the three model
        components (i.e., maximum of the main stream, spur, or background
        components). For stars with membership probability greater than $75\%$,
        we also plot their astrometric errorbars from \package{Gaia} as the thin
        red lines.

        \begin{figure*}[t]
            \script{gd1/plot/results_panels.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/gd1/results_panels.pdf}
            \caption{%
                Photometric coordinate $(g, r)$ plots for each model component
                in 4 $\phi_1$ ranges across the data set. Over-plotted is the
                model isochrone, shifted to the mean distance of the track
                within the bin of $\phi_1$. Importantly this means we don't
                expect the mean isochrone to exactly match the distribution of
                points. See \autoref{fig:gd1-results-full} for the match of the
                distance gradient to the data and prior literature results.
                %
            }
            \label{fig:gd1-results-panels}
        \end{figure*}

        Because the neural networks are trained with dropout, we can estimate
        the modeling uncertainty in our fits to each stream with a Monte Carlo
        procedure \citep{GalGhahramani2015}. Namely, by incorporating dropout
        during both neural network training and inference, the distribution of
        model outputs represents the posterior predictive distribution,
        marginalized over the neural network parameters. An estimate of this
        distribution is shown in the weight parameter of
        \autoref{fig:gd1-results-full}, represented as the red (main stream) and
        yellow (spur) bands. For the astrometric dimensions, our model predicts
        a mean track, and a standard deviation. Error bands are estimated by
        incorporating dropout with a rate of 15\%. Each of the panels in
        \autoref{fig:gd1-results-full} shows the track mean and width,
        calculated from the 50th percentile of the posterior predictive
        distribution for both quantities. Plotted behind that is the 90\%
        credible region of the track mean. We note that the credible region of
        the mean alone is generally larger than that of the $\mu \pm \sigma$
        50th-percentile ridge. This highlights the importance of exploring the
        full posterior of parameters, not only a Maximum Likelihood Estimation
        (MLE) or posterior ridge-line.

        Our model finds \variable{output/gd1/nstream/nstream_variable.txt}stars
        that belong to the main stream component with membership probability
        greater than
        \variable{output/gd1/nstream/minimum_membership_probability.txt} at the
        \variable{output/gd1/nstream/posterior_percentile.txt} percentile in the
        membership posterior distribution. For the spur component, there are
        \variable{output/gd1/nspur/nspur_variable.txt} stars with the same
        membership probability threshold at the same percentile of the posterior
        distribution.  A tabulation of the astrometry, photometry, and
        membership posterior distribution for each star is included in
        \autoref{tab:gd1_member_table}. The full table may be found on the
        paper's
        \href{https://github.com/nstarman/stellar_stream_density_ml_paper}{GitHub
        repository}. \autoref{tab:gd1_member_table} gives the MLE prediction,
        but also samples the 5th-95th percentile of the posterior distribution
        of the likelihood using the dropout method. The variation in the 90\%
        credible region in the track means there can be significant variation in
        a star's estimated membership probability.

        The extent of the main stream in $\phi_1$ is roughly $70~\rm{deg}$, and
        the angular extent of the spur is roughly $25~\rm{deg}$. The distance
        track (bottom panel) places the stream in the range of $6-10~\rm{kpc}$,
        with an upwards concavity as a function of $\phi_1$. This result is
        consistent with other works, which find a similar distance to
        \stream{GD-1} using other techniques \citep{deBoer+2020, Ibata+2021}.
        Importantly, our distance estimate is not hard-coded to fall in this
        range: we allow for the stream track to fall anywhere within
        $[4,40]~\rm{kpc}$, with a broad prior around previous literature
        estimates. Therefore, the similarity of our inferred distance track
        compared to other works indicates that our photometric model is
        performing satisfactorily.

        We also recover the well-known underdensity gaps along the stream around
        $\phi_1 \approx -40, -20 \ \rm{and } -5~\rm{deg}$ (seen best in
        \autoref{fig:gd1-heatmap}). The $-20\degree$ feature was first reported
        by \citet{CarlbergGrillmair2013}, and is one location hypothesized for
        the dissolved stream progenitor. \citep{WebbBovy2019} suggests the
        $-40\degree$ gap might be  the progenitor location.  The recovery of
        this density fluctuation from a joint model in astrometric and
        photometric coordinates developed here argues that the feature is indeed
        intrinsic to the stream and not an artifact of the data.

        The spur is a distinct component of our model, independent in its weight
        and all astrometric dimensions except $\parallax$ which is shared with
        the \stream{GD-1} main stream through the photometric model. The model
        finds a non-negligible spur feature, with fractional weight approaching
        the thin stream at its peak. Corroborating this finding, the proper
        motion of the spur almost exactly agrees with the thin stream. The
        alignment of these dimensions is not hard-coded in our model. The width
        of the spur in proper motion is very thin, thinner even than the stream.
        However with the large observational errors and low number density of
        the spur it is difficult to attribute this to a physical difference
        between the two systems. In contrast, the small errors in $\phi_2$
        indicate the spur's width diverges from the thin stream at $\phi_1 > -35
        \ \rm{deg}$, whereas the width is similar at lower $\phi_1$. There are
        numerous mechanisms by which a stream's width may vary \citep[e.g.
        epicyclic effects as in][]{Ibata+2020}, so differences between the spur
        and stream are both notable and expected.

        We note that the model-predicted track does not always closely align
        with the predicted member stars or our notion of a slowly-evolving
        stream. This misalignment is most evident in $\mu_{\phi_1^*}$, where the
        track has a sharp turn up at $\phi_1 \sim -65 \ \rm{deg}$ and then
        flattens out. There are a number of contributing factors. First, the
        observational uncertainties are large enough that this flattening is
        consistent with the data. Second, we do not enforce the stream track to
        be smooth with $\phi_1$ -- our fits are dictated by the data.  Last,
        examining 2D slices when fitting in high dimensional spaces can be
        misleading \citep{Aggarwal+2002}. Unsurprisingly, the raw membership
        probabilities for each model component provide a more accurate
        representation of the stream's 6d phase-space distribution than any
        phase-space slice could, as may be seen in \autoref{fig:gd1-heatmap}. We
        discuss the smoothed stream model determined solely by membership
        probabilities at the end of this section.

        We highlight the performance of our photometric model in
        \autoref{fig:gd1-results-panels}. In this figure, we color-code each
        star by its membership probability (top row; same as
        \autoref{fig:gd1-results-full}), and in the bottom row we visualize the
        performance of our photometric model in bands of $\phi_1$. Our model
        isochrone, which shifts vertically due to our estimate of a distance
        modulus, is shown in green (we plot the mean-distance isochrone in each
        $\phi_1$ band). Because our model preforms a joint fit in photometry and
        astrometry, the most probable stream members do not need to fall exactly
        along the isochrone.  Still, the likely stream members are
        overwhelmingly distributed as might be expected from a single-stellar
        population.  Even the more diffuse tails of the stream can be seen in
        the CMD distribution, where they still roughly follow a single-stellar
        population.

        With membership probabilities for each component of the stream estimated
        from both astrometry and photometry, it is possible to quickly construct
        a custom representation of the stream's phase-space density. We
        illustrate this in  \autoref{fig:gd1-heatmap}, which shows a Gaussian
        kernel density estimate (KDE) of bandwidth $0.1 \ \rm{deg}$ fit to the
        model. The primary overdensities at $\phi_1 = (-50, -30, -10) \
        \rm{deg}$ are all visible in $\phi_2$. The gap at $20 \ \rm{deg}$ is
        prominent in all features, but the gap at $-40$ is less visible in
        proper motions, since the spur shares the stream's proper motions and
        thus contributes to the density in those coordinates. This highlights
        the importance of constructing a multidimensional smooth density model
        of streams.

        \begin{figure}[h]
            \script{gd1/plot/smooth_likelihood.py}
            \centering
            \hspace{-25  pt}\includegraphics[width=1.1\linewidth]{figures/gd1/smooth_likelihood.pdf}
            \caption{%
                Smoothed KDE model of \stream{GD-1}'s member stars, weighted by
                their membership probability. The KDE has a bandwidth of $0.1 \
                \rm{deg}$.  In $\phi_2$ the spur is clearly visible in the KDE.
                The gaps at $-20\degree$ is prominent in both $\phi_2$ and
                $\mu_{\phi_1}$.  The gap at $-40\degree$ is most prominent in
                $\phi_2$.
                %
            }
            \label{fig:gd1-heatmap}
        \end{figure}

% sub: gd1 (end)

% \newpage
\section{Results: Palomar 5} \label{sec:results_pal5}

    We now apply the method to the stellar stream \stream{Palomar 5}.
    Historically lengths of the stream have been found with matched filter
    photometric techniques, like with the discovery in \citet{Odenkirchen+2001}.
    Post-\package{Gaia} the stream was also observed by filtering kinematically
    \cite[e.g.,][]{Starkman+2019, Ibata+2021}.  However at $\sim 20 \ \rm{kpc}$,
    the \stream{Pal\,5} main sequence is not observed by \package{Gaia}, making
    a high signal-to-noise astrometric detection challenging.
    %In this work we detect over $10 \ \rm{deg}$ of stream.
    We discuss the data selection in \autoref{sub:results_pal5:data}, model
    specifics in \autoref{sub:results_pal5:model}, and the results in
    \autoref{sub:results_pal5:results}.

    \subsection{Data Selection} \label{sub:results_pal5:data}

        We construct a \Gaia{} DR3 - \PanStarrs{} crossmatched field very
        similarly to the procedure outlined in \autoref{sub:results_gd1:data}
        for \stream{GD-1}.  To limit the field to a region containing
        \stream{Pal\;5} we adopt broad cuts on the data by requiring the data
        lie within the phase-space cube:
        \begin{itemize}
            \setlength\itemsep{0em}
            \item $\phi_1 \in [-20, 20] \ \rm{deg}$,
            \item $\phi_2 \in [-5, 10] \ \rm{deg}$,
            \item $\parallax \in [-10, 1] \ \rm{mas}$,
            \item $\rm{G-R} \in [-0.5, 1.2] \ \rm{mag}$
            \item $g, r \in [0, 30] \ \rm{mag}$,
        \end{itemize}
        where we use \package{gala}'s \code{Pal5PriceWhelan18} frame
        \citep{Price-Whelan+2019} to define $\phi_1$, $\phi_2$, and associated
        proper motions.  Using the \textit{Bayestar19} maps we correct the
        \PanStarrs{} photometry for dust extinction.  For stars without distance
        information we assume a distance of $20 \ \rm{kpc}$, a typical distance
        to the stream \citep{Harris1996, Odenkirchen2003, Grillmair2006,
        Ibata+2016, Price-Whelan+2019} and thus increasing the probability of
        the star being considered a member.  The 1$\sigma$ dust correction
        errors are included by quadrature in $\mbs{\Sigma}_n$, the per-star
        photometric uncertainty. We also correct the parallax for the \Gaia\
        zero-point as in \autoref{sub:results_gd1:data}.  The result is a field
        populated by \variable{output/pal5/ndata_variable.txt} stars, from which
        we would like to identify stream members.

        Within the broad selection cuts, we create more specific data masks.  In
        particular we mask concentrated over-densities, like the globular
        cluster \stream{M5}, that are not associated with \stream{Pal\;5}.  In
        \autoref{fig:pal5-data_selection} we show the proper motion and
        photometric masks, with \stream{M5} already filtered out.  The proper
        motion mask extends from $\mu_{\phi_1}$ and $\mu_{\phi_2} \in (-4, -1)$
        [mas/yr].  The photometric mask is defined by a $0.15 \ \rm{mag}$ buffer
        around a $[\rm{Fe}/{H}] = -1.3$, $11.5 \ \rm{Gyr}$ isochrone. It is the
        shape of the isochrone which is important, not the physical reality of
        the metallicity nor age, discussed in further detail in
        \autoref{ssub:method:photometric_model:on_stream}.  In addition, to
        avoid detailed modeling of the completeness as a term in $\prior_{obs}$
        (from \autoref{eq:isochrone_prior_components}) we mask out photometry
        with $G > 20$.  With combinations of these masks the stream has much
        larger signal-to-noise, though is still only a small fraction of the
        total data.

        \begin{figure*}[ht]
            \script{pal5/plot/data_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/pal5/data_selection.pdf}
            \caption{%
                Data selections for \stream{Pal\,5}, within the data cube
                described in \autoref{sub:results_pal5:data}.
                %
                \textbf{Row 1 Panels:}
                Proper motion and photometric selections.  \stream{Pal\,5} is
                apparent in the proper motions as a small overdensity. The
                photometric selection is based on a 12 Gyr, [Fe/H] = -1.2 MIST
                isochrone, buffered by 0.3 dex.
                %
                \textbf{Rows 2 \& 3 Panels:}
                Applying the combination of astrometric and photometric
                selections, plotted for astrometric coordinates $\phi_2,
                \parallax, \mu_{\phi_1}, \mu_{\phi_2}$.  The stream is an
                identifiable overdensity in only $\phi_2$ and errors dominate
                all other coordinates.  }
            \label{fig:pal5-data_selection}
        \end{figure*}

        The reduced dataset, with all specific masks applied, is included below
        the data mask plots in \autoref{fig:pal5-data_selection}. In $\phi_2$,
        $\mu_{\phi_1}^*$, and $\mu_{\phi_2}$ the \stream{Pal\,5} progenitor is
        clearly visible. Only in $\phi_2$ is the stream visible, and then only
        for a few degrees around the progenitor. The distance of \stream{Pal\,5}
        means most $\parallax$ measurements have $\sim 100\%$ errors, so neither
        the stream nor even the progenitor are evident.

    % sub data_selection (end)

    \subsection{Model Specification} \label{sub:results_pal5:model}

        In this section we discuss how the total \stream{Pal\,5} model is built,
        using the modular modeling framework from \autoref{sec:method}.  We
        model the stream of \stream{Pal\;5} as a single Gaussian component
        conditioned on $\phi_1$. As a function of $\phi_1$, the astrometric
        stream density is a 3-dimensional Gaussian in $\phi_2, \mu_{\phi_1},
        \mu_{\phi_2}$, using the model defined in
        \autoref{ssub:method:astrometric_model:on_stream}.  The parameter
        network is a 5-layer \texttt{LTD} where the 6 outputs are $\mbs{\theta}
        = [\mu_{\phi_2}, \sigma_{\phi_2}, \mu_{\mu_{\phi_1}^*},
        \sigma_{\mu_{\phi_1}^*}, \mu_{\mu_{\phi_2}},
        \sigma_{\mu_{\phi_2}}](\phi_1)$.  In testing, the network is
        sufficiently large to permit 15\% dropout and retain robust predictions.
        At $\sim 20 \ \rm{kpc}$ any \package{Gaia}-crossmatched data is not
        photometrically deep enough to contain the main sequence nor its turn
        off. Thus we do not model the stream photometrically. We only use
        photometric cuts to remove some background. Similarly, we do not include
        the parallax as the errors are too large for the parallax to contribute
        to the model.

        \begin{figure}[ht]
            \script{pal5/plot/photometric_background_selection.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/pal5/photometric_background_selection.pdf}
            \caption{%
                \textbf{Top panel:} Astrometric coordinates ($\phi_1$,
                $\phi_2$).  Light blue points are the stream and background
                within an ``on"-stream region. Black points are in the
                ``off"-stream region used to train the background photometric
                model. The \stream{M5} cluster is masked as a black circle
                cutting into the on-stream selection at $\phi_1\sim-1\degree$.
                %
                \textbf{Bottom panel:} CMD plot of the astrometric selection
                above. The stream's isochrone is somewhat apparent, but the
                background contamination is large.
                %
            }
            \label{fig:pal5-photometric_background_selection}
        \end{figure}

        To avoid a large convergence time, we use as a prior the known stream
        track, in this case \package{galstream}'s \citep{Mateu2022}
        implementation of \stream{Pal\,5} from \citet{Ibata+2021}. From the
        track, we set down guides (see
        \autoref{sub:methods:priors:track_region_prior}) regularly spaced in
        $\phi_1$. We include all the guides in \autoref{tab:pal5_track_prior}.
        We don't place guides in the other coordinates, except at the location
        of the progenitor, because the stream is not visible and the models
        disagree on the proper motions (see \autoref{fig:pal5-data_selection},
        top left plot).  The progenitor is an exception, both visible
        kinematically and also extremely well measured
        \citep{VasilievBaumgardt2021}, and we place a tight kinematic prior on
        the stream track at the location of the progenitor.  The guide priors
        are visible in \autoref{fig:pal5-results-full}.
        \variable{output/pal5/control_points.tex}

        The background is fit with an analytic distribution and a normalizing
        flow:
        \begin{itemize}
            \setlength\itemsep{0em}
            \item $\phi_2 \sim \rm{TruncExp}(\lambda; a, b)$: a truncated
                exponential distribution (see
                \autoref{app:sub:exponential_distribution}).  The distribution
                has 1 parameter and we use a \texttt{LTD} neural network with
                15\% dropout.
            \item $\mu_{\phi_1}, \mu_{\phi_2} \sim q_{\rm flow}$: a two feature,
                one context feature normalizing flow trained on an off-stream
                selection (see
                \autoref{fig:pal5-photometric_background_selection} for that
                selection). \autoref{sub:method:pre-training_distributions}
                details how these flows are constructed and trained.
        \end{itemize}

        All the component models are tied together into a mixture model.  The
        weight parameter network is the usual \texttt{LTD} with 15\% dropout.
        The stream weights are in the range $\ln{f_{\rm stream}} \in [-10, 0]$
        and set to $-\infty$ for $\phi_1 \notin [-10^\circ, 10^\circ]$. By
        \autoref{eq:mixture_weight_normalization} the background  brings the
        cumulative weight to one, normalizing the mixture.

    % sub model_specification (end)

    \subsection{Trained Density Model} \label{sub:results_pal5:results}

        \begin{figure*}[ht]
            \script{pal5/plot/results_full.py}
            \centering
            \includegraphics[width=0.85\linewidth]{figures/pal5/results_full.pdf}
            \caption{%
                The model for \stream{Pal\,5}.  For comparison we include the
                tracks from \citet{Price-Whelan+2019} and \citet{Ibata+2021}.
                Note how the model largely agrees with both tracks in all
                astrometric dimensions from $-5 < \phi_1 < 5 \ \rm{deg}$. The
                stream fraction and number density drop to near $0$ for $5 \
                \rm{deg} < \phi_1$, making the track dominated by low-number
                statistics. The progenitor is not separated from the stream as a
                distinct component, demonstrating how the neural networks can
                capture a large range of linear densities.
                %
                \textbf{Panel 1: }%
                The stream mixture coefficients $f(\phi_1)$ predicted by the
                model, colored by the stream membership likelihood.
                %
                \textbf{Panel 2: }
                $\phi_2(\phi_1)$ over the full range of $\phi_1$. The two stream
                arms visibly emerge from the Lagrange points of the stream.  The
                data color and transparency is set by the model's
                50th-percentile stream and spur probability. We include the
                error bars for stars with $>75\%$ membership probability. The
                predicted 50th-percentile track $\pm$ width is over-plotted (red
                band), along with the 90\% confidence region (broader red bend).
                In the adjoining top plot is shown the 5th to 95th percentile
                variation in the width.
                %
                \textbf{Panel 3: }%
                $\parallax(\phi_1)$ and $d(\phi_1)$ over the full range of
                $\phi_1$.
                %
                \textbf{Panel 4 \& 5: }%
                The proper motions $\mu_{\phi_1^*}, \mu_{\phi_2}$. The model
                convolves observational errors with the stream width, but the
                large fractional errors make the determination difficult. The
                track is largely consistent with 0 width driven by small number
                statistics.  }
            \label{fig:pal5-results-full}
        \end{figure*}

        \begin{figure*}[htp]
            \script{pal5/plot/results_panels.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/pal5/results_panels.pdf}
            \caption{%
                Photometric coordinate $(g, r)$ plots for each model component
                in 5 $\phi_1$ ranges across the data set, centered on the
                \stream{Pal\,5} progenitor.
                %
                On either side of the progenitor we strongly detect the stream
                for $\approx 5\degree$  and tentatively for an additional
                $\approx 5\degree$. The leading arm (positive $\phi_1$) track
                for $\phi_1 > 5\degree$ is consistent with
                \citet{Starkman+2019}. We do detect any fanning of the stream,
                but given the shallow photometry, this is expected.
                %
            }
            \label{fig:pal5-results-panels}
        \end{figure*}

        \variable{output/pal5/member_table_select.tex}

        Our fit to \stream{Pal\,5} and its tidal tails is illustrated in
        \autoref{fig:pal5-results-full}. In the top panel we plot the stream
        fraction and its uncertainty, estimated by applying dropout during
        inference and evaluation of our model. The model clearly identifies the
        progenitor cluster $(\phi_1 = 0)$.

        The membership probability of stream stars is illustrated in subsequent
        panels of \autoref{fig:pal5-results-full}, with guide points and their
        widths shown in red. The leading and trailing arms of the stream are
        clearly identified in our fits, with high membership probability on
        either side of the cluster. Our model prefers a thin on-sky morphology
        for \stream{Pal\,5}'s tidal tails, with an angular width of the stream
        (parameterized by $\ln{\sigma_{\phi_2}}$) around $\sim 0.1~\rm{deg}$.
        This is contrary to our fit of \stream{GD-1} in
        \autoref{fig:gd1-results-full}, where we find a more appreciable width
        of $\sim 0.2~\rm{deg}$ and above.  This inferred width is consistent
        with \stream{Pal\,5} being more distant than \stream{GD-1}, though we
        expect a thin inferred width for \stream{Pal\,5} since we do not model
        the stream in photometric coordinates. When photometry is available, we
        expect that our photometric model provides substantial constraining
        power in characterizing the stream, since diffuse parts of the tidal
        tails in on-sky angular coordinates remain tightly clustered in the CMD.
        This clustering is seen in our \stream{GD-1} fits
        (\autoref{fig:gd1-results-panels}), where even the more extended parts
        of the stream remain clustered in the CMD.

        The parallax, proper motion tracks, and their widths are also shown in
        \autoref{fig:pal5-results-full}. The red vertical lines represent the
        astrometric errors in each dimension, shown only for stars with
        membership probability greater than $75\%$. The errors are large
        compared to the width of the best fit stream track (shown as the red
        band). The track is a valid minimum to our loss function: because our
        model incorporates the large astrometric uncertainties, small variations
        to the estimated width of the stream in each dimension do not produce
        substantial changes to the loss function. Thus, a narrow stream in the
        kinematic dimensions provides a valid fit to the stream given the data
        at hand.

        We also note that the track of the stream shown as the red band in each
        dimension appears to have an abrupt change in direction around $\phi_1
        \approx 7^\circ$. This also represents the region with the lowest stream
        fraction, as seen in the top panel of \autoref{fig:pal5-results-full}.
        In low stream fraction regions, the model is uncertain about the track
        of the stream since only a few stars contribute to the on-stream fit.
        The track changes highlight the importance of utilizing the membership
        probabilities of individual stars when analyzing any given stream, since
        membership probabilities incorporate information on the stream fraction
        while the mean stream-track alone does not provide this information.

        In \autoref{fig:pal5-results-panels} we again show the data in
        $\phi_1-\phi_2$ coordinates, color-coded by membership probabilities in
        the top panel (same colorbar as in \autoref{fig:pal5-results-full}). In
        the bottom panels we plot the CMD of the data in bands of $\phi_1$. As
        discussed, our fit to \stream{Pal\,5} does not incorporate information
        from the CMD.  Thus, the bottom row represents a data-driven isochrone
        for \stream{Pal\,5} and its tidal tails, obtained by fitting the
        stream's astrometric overdensity. Indeed, the highest probability
        regions of the CMD appear qualitatively similar to a standard isochrone,
        with an extended and clustered sequence. The main sequence appears to be
        missing, which is expected given our cut at $G = 20~\rm{mag}$, below
        which the main sequence of \stream{Pal\,5} resides \citep{Bonaca+2020}.
        The similarity of the color-magnitude distribution of \stream{Pal\,5}'s
        tidal tails to the red end of a standard isochrone highlights the
        validity of our astrometric membership model, since a photometric fit
        was never performed.

        The stream solution, ie the per-star membership likelihood, is a
        posterior distribution. \autoref{fig:pal5-results-full} and
        \autoref{fig:pal5-results-panels} show the membership likelihood
        evaluated at the $50$th percentile of this distribution, per star. In
        \autoref{fig:pal5-heatmap} we sample over the full distribution and
        smooth by a 0.1 $deg$ kernel to produce a smooth representation of the
        stream. As a consequence of the sampling we see the model's stream
        solution extends beyond the $\approx \pm5\degree$ shown previously.
        Though the model certainty and individual membership likelihoods are
        low, the stream path matches closely with findings in deeper photometric
        catalogues, like \citet{Ibata+2017}.

        \begin{figure}[h]
            \script{pal5/plot/smooth_likelihood.py}
            \centering
            \includegraphics[width=1\linewidth]{figures/pal5/smooth_likelihood.pdf}
            \caption{%
                Smoothed KDE model of the stream's member stars, weighted by
                their membership probability. The KDE has a bandwidth of $0.1 \
                \rm{deg}$. We exclude the progenitor to limit the density range
                and better display the diffuse tidal tails. The stream extends
                from $-12$ to $10$ degrees in this plot, which appears longer
                than in \autoref{fig:pal5-results-full} and
                \autoref{fig:pal5-results-panels}. Those show the membership
                likelihood at the $50$th-percentile of its distribution. This
                figure samples over the full posterior, showing the contribution
                of less-likely members.
                %
            }
            \label{fig:pal5-heatmap}
        \end{figure}

        While \package{Gaia} alone is not photometrically capable of observing
        the main sequence of \stream{Pal\,5}'s tidal tails, the modeling
        framework we present here is capable of operating on outer-joined
        datasets. This allows us to easily combine deep photometric data with
        kinematics from, e.g., \package{Gaia}.  We leave that to a future work,
        though expect that a simultaneous model in photometry and astrometry
        would help identify the more diffuse parts of the stream, which become
        washed out in our astrometric fit due to the substantial number of
        background stars. Still, this section highlights the capability of our
        model when applied to a noisy field with a diffuse stream.

    % sub trained_density_model

% section pal5 (end)

\section{Discussion} \label{sec:discussion}

    We now provide a discussion of the methods presented in the paper, focusing
    on future directions in \autoref{sec:conclusions} and comparison to other
    works in \autoref{sec:comparison}.

    \subsection{Limitations}\label{sec:limitations}

        We first highlight the limitations of our method, and areas for
        potential improvement.

        First, our usage of neural networks ensures that our method is
        expressive, and capable of describing a diverse range of stream
        morphologies all under one unified framework. At the same time, neural
        network training is computationally expensive, and takes a long time on
        a local device. Our fits to \stream{GD-1} and \stream{Pal\,5} presented
        in this work were performed on a local computer without GPU support,
        though training time takes on the order of a few hours per stream. This
        can be improved substantially with the use of GPUs.

        Second, our fits to \stream{GD-1} in \S\ref{sec:results_gd1} rely on a
        theoretical isochrone. While the stream does appear to be decently
        characterized by a theoretical isochrone, there is room for improvement.
        Especially considering that globular clusters do not always follow
        simple single-stellar population (see, e.g.,
        \citealt{2022Univ....8..359M} for a review of this subject), it is worth
        considering whether more data-driven approaches can be used to model the
        photometry of a stream without imposing tight priors on the stellar
        population.

        Third, we have used a combination of analytic distributions and more
        flexible machine learning methods like normalizing flows to describe the
        distribution of background stars. However, the field of the Galaxy is
        complicated, in all coordinates. The analytic distributions in our
        modeling are clearly approximate. While the analytic distributions might
        be made more complex, the alternative is to substitute for
        non-parameteric flexible models. We have postponed this to future work,
        since the analytic backgrounds provide a decent description in the
        fields of \stream{GD-1} and \stream{Pal\,5}. However, this might not be
        the case for other streams. Constructing flexible background models will
        therefore provide further improvement over our method.

        Finally, our method is suited for characterizing streams rather than
        detecting them. Thus, some amount of supervised work is required to
        reduce the data down to a clean enough selection of stars so that our
        model can converge to a high likelihood solution. If the field is too
        noisy and control points cannot be reliably placed, our method might not
        be able to characterize the stream of interest. An end-to-end stream
        discovery and detailed characterization algorithm could therefore
        provide a substantial improvement over our method and existing work.

    \subsection{Comparison to Existing Methods} \label{sec:comparison}

        There are several existing methods for the detection of stellar streams,
        and some existing work on detailed stream characterization. Here we
        briefly overview the existing methods, and highlight commonalities and
        differences between the results of our work and previous studies.

        Currently, the matched filter is the most common method for detecting
        and characterizing streams. The method works by selecting a stellar
        population in the CMD (i.e., from a metal-poor isochrone) and searching
        for a stellar overdensity in the astrometric data. The approach has
        proved successful for stream detection, but to a lesser extent for
        characterization. From the matched filter approach alone, stream
        membership probabilities are not obtained and there is substantial
        contamination from background stars. While the approach can provide an
        excellent coarse-grained description of a given stream, it does not
        provide a density model in astrometry and photometry.

        To a good approximation streams are expected to roughly trace a stellar
        orbit. The Streamfinder algorithm exploits this physical expectation
        \citep{STREAMFINDER}, by integrating orbits in a trial Galactic
        potential and identifying stellar overdensities that appear to coincide
        with the trial orbits. While the method has identified many stream
        candidates, using these streams for potential reconstruction is circular
        in reasoning. That is, the detection and characterization of these
        streams is conditioned on a choice of the potential. The method we
        present here does not make assumptions about the gravitational
        potential, nor do we enforce physical assumptions about the dynamics of
        each stream.

        This work represents one machine learning-based approach for stream
        characterization. Another method is presented in \citet{Shih+2022}, VIA
        MACHINAE, which uses conditional density estimation and a linear feature
        detector to discover streams and identify probable members. While the
        method is promising for stream discovery, its intended use is not for
        detailed characterization (as stated in \citealt{Shih+2022}). Our work
        has the opposite approach: the method presented is not designed to
        detect streams, rather, it is well-suited for characterizing streams
        that fall within some predefined field. Thus, our method can be used for
        characterizing the VIA MACHINAE candidates, and obtaining more robust
        membership probabilities.

        In \citet{Patrick+2022} the stellar density of 13 tidal streams is
        modeled in on-sky positions and color-magnitude coordinates. The
        morphology of each stream is characterized by a series of splines, which
        are themselves a function of $\phi_1$. This method was originally
        developed in \citet{Erkal+2017}.  Importantly, \citet{Patrick+2022}
        performs a fit in $\phi_1-\phi_2$ and photometry, similar to our work.
        The photometric approach in \citet{Patrick+2022} relies on generating
        mock stellar populations in the CMD, selecting probable stream stars,
        modeling the $\phi_1-\phi_2$ distribution for those stars, and then
        refining the CMD fit based on the $\phi_1-\phi_2$ fit. This approach
        therefore represents a more principled way of carrying out a matched
        filter analysis. The model does not currently incorporate kinematic
        information, though the work could be extended. The improvement of our
        work is that we do not require an iterative fit to the CMD and
        astrometry: we handle both spaces through a joint likelihood without
        simulating mock-observations. Additionally, we can easily incorporate
        kinematic information and data with missing phase-space dimensions.

        Finally, our method relies on neural networks to describe the stream
        track rather than splines, and a combination of normalizing flows and
        analytic densities to characterize the distribution of background stars
        in photometry and astrometry. While splines are attractive in their
        simplicity and flexibility, the expressiveness of a spline is largely
        determined by (a) the number of spline ``knots", and (b) the location of
        each knot. This is not entirely dissimilar from a neural network, which
        is subject to some choice of hyper-parameters. However, neural networks
        are universal function approximators and are therefore readily equipped
        to describe non-trivial density distributions without imposing priors on
        the ``reach" of a model. However, when priors are needed, we have shown
        that they can be added as control points in the loss function, guiding
        our fits to prefer extended filamentary structures on the sky.
        Therefore, our neural network based approach enjoys many of the same
        benfits as splines, with the benefit of being more agnostic to choices
        like the number and location of knots, and the order of the spline
        (i.e., cubic, quartic, etc.).

        We also highlight upcoming work from Tavangar et al. 2023 (in prep.),
        which provides a spline-based characterization of streams implemented in
        the \package{Jax} library \citep{jax2018github}. The \package{Jax}
        backend ensures that fitting the model to a stream is very fast (on the
        order of minutes rather than hours), and also has the capability of
        incorporating kinematic data with missing phase-space dimensions. The
        method does not currently model the color-magnitude distribution of a
        stream, but relies on a loose photometric selection. The photometric
        model in this work could be extended to the method in Tavangar et al.
        2023 (In prep.) in a straightforward matter.

        In total, having multiple methods to characterize streams and their
        membership probabilities is advantageous. As the volume of tidal
        features continues to grow, it will become important to fit different
        models to the same stream, in order to evaluate whether certain features
        or density variations are real versus spurious artifacts from the
        adopted method. Considering the information encoded about perturbations
        in the small-scale density fluctuations along a given stream, a
        posterior distribution over models will help capture systematic errors
        in our description of streams and the field of perturbations that give
        rise to their density.

\section{Summary and Conclusion} \label{sec:conclusions}

    We have constructed and demonstrated a new method for characterizing stellar
    streams, given all of the astrometric and photometric data available. We
    model a stream in astrometric coordinates using mixture density networks,
    allowing for a flexible representation of diverse stream morphologies under
    a single modeling framework. Additionally, our astrometric fit does not rely
    on assumptions about the gravitational potential. Our simultaneous
    photometric model treats each stream as a single-stellar population,
    generated from a theoretical isochrone. The distance modulus along the
    stream is itself a function of $\phi_1$, allowing us to tie photometric
    information to parallax measurements along the stream. This tie enables a
    reconstruction of stellar stream distance tracks that are compatible with
    photometry and astrometry by construction. A joint model in astrometric and
    photometric coordinates also naturally allows for more spatially diffuse
    regions of a stream to be detected. Additionally, our background model
    relies on a combination of analytic densities and normalizing flows, each of
    which are modular in our code-base and readily adaptable to a given field.

    The combination of a flexible stream and background model allows us to
    obtain robust stream-membership probabilities for each star in a given
    field. This enables the generation of statistical samples for stream stars,
    informed by astrometry and photometry. We apply the method to the stream
    \stream{GD-1} and the tidal tails of \stream{Palomar 5}, generating
    membership catalogs for both streams that are publicly available
    \JN{reference link or footnote}. Our characterization of \stream{GD-1}
    reveals a significant detection of the stream's main component as well as
    the ``spur" component, which represents a bifurcation from the main stream.
    We also recover density variations and gaps along the stream that have been
    previously observed. Our fit to \stream{Pal\,5} demonstrates the performance
    of our model on a more diffuse stream with low signal-to-noise measurements.
    The inclusion of deeper photometric data for \stream{Pal\,5} will enable a
    more extended view of the stream.

    Our method represents a new avenue to characterize the growing census of
    stellar streams discovered in the Milky Way. While most streams have been
    roughly characterized with matched filter algorithms, a statistical and
    homogeneous determination of stream-membership probabilities for each star
    has not been performed or made publicly available. A homogeneous catalog of
    stellar streams and stellar membership probabilities will therefore
    represent a substantial step forward for galactic dynamics, since both the
    global track of a stream and its small-scale density distribution are
    sensitive to the Milky Way's gravitational potential.

    Especially when considering the small-scale perturbation history of a stream
    predicted by $\Lambda\rm{CDM}$, the inclusion or exclusion of stream stars
    can make a substantial difference in our inference of the perturbers.
    Constructing catalogs with membership probabilities as we do in this work
    provides a crucial step forward to making our inference of, e.g., dark
    matter subhalos in the Galaxy more robust. Equipped with membership
    probabilities, one is finally able to propagate the uncertainty in our
    determination of the ``on-stream data" through to the posterior distribution
    of model parameters (e.g., those characterizing the subhalo mass function).
    Thus, our works provides a new technique to generate samples of stream stars
    using all of the existing kinematic and photometric data, from which models
    for the Milky Way and its potential can be constrained in a statistically
    sound manner.

% section comparison_to_other_methods (end)

\section{acknowledgements}

    NS acknowledges support from the Natural Sciences and Engineering Research
    Council of Canada (NSERC) - Canadian Graduate Scholarships Doctorate Program
    [funding reference number 547219 - 2020].  NS received partial support from
    NSERC (funding reference number RGPIN-2020-04712) and from an Ontario Early
    Researcher Award (ER16-12-061; PI Bovy).

    JN is supported by a National Science Foundation Graduate Research
    Fellowship, Grant No. \JN{fill in}.

    Both NS and JN would like to thank the
    \href{https://stellarstreams.org/streams22/}{Community Atlas of Tidal
    Streams 2022} conference for starting the discussions that led to this work.

    This work has made use of data from the European Space Agency (ESA) mission
    \Gaia{} (\url{https://www.cosmos.esa.int/gaia}), processed by the \Gaia{}
    Data Processing and Analysis Consortium (DPAC,
    \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the
    DPAC has been provided by national institutions, in particular the
    institutions participating in the \Gaia{} Multilateral Agreement.

    The data availability statement is modified from one provided to
    ShowYourWork by Mathieu Renzo.

    \paragraph{Software (alphabetical)}

        \package{asdf} \citep{Greenfield+2015}, %
        \package{astropy} \citep{Astropy2013, Astropy2018, Astropy2022}, %
        \package{astroquery} \citep{Astroquery2019}, %
        \package{brutus} \citep{brutus}, %
        \package{Matplotlib} \citep{Hunter2007}, %
        \package{NumPy} \citep{Harris+2020}, %
        \package{Pytorch} \citep{Pytorch2019}, %
        \package{SciPy} \citep{Scipy2020}, %
        \package{ShowYourWork} \citep{Luger+2021} %
        \package{zuko} \citep{zuko}, %

\section*{Data Availability} \label{sec:data_availability}

    This study was carried out using the reproducibility software
    \href{https://github.com/showyourwork/showyourwork}{\showyourwork}
    \citep{Luger+2021}, which uses continuous integration to programmatically
    download the data, perform the analyses, create the figures, and compile the
    manuscript. Each figure caption contains two links: one to the dataset used
    in the corresponding figure, and the other to the script used to make the
    figure. The datasets are stored at \todo[inline]{zenodo url}. The git
    repository associated with this study is publicly available at
    \package{\url{https://github.com/nstarman/stellar_stream_density_ml_paper}}.

% section data_availability (end)

\bibliography{bib}

\appendix

\section{Probability Distributions} \label{app:distributions}

    A variety of probability distributions are used when modeling the stream and
    Galactic background. This appendix presents relevant mathematical details of
    these distributions.  In particular, observational errors are modeled as
    Gaussian distributions, which requires modifying each distribution to
    account for the Gaussian noise.

    \vspace{10pt}

    Let $x \sim X$ be distributed as some specified distribution and $\delta
    \sim \Delta \equiv \mcal{N}(0, \sigma_*)$ as an independent, centered
    Gaussian.  An observation $Y$ is distributed as
    \begin{equation}
        Z = X + \Delta,
    \end{equation}
    Which is a convolution of the PDFs.
    \begin{align} \label{eq:general_convolution}
        \pdf_Z(x; \mbs{\theta}, \sigma)
            & \equiv (\pdf_X(x; \mbs{\theta}) * \mcal{N}(x; 0, \sigma)) \\
            &= \int_{-\infty}^{\infty} \rm{d}\tau \, \pdf(x; \mbs{\theta})\mcal{N}(x-\tau; 0, \sigma)
    \end{align}

    In the following subsections we work through the distributions
    $\pdf(x,\mbs{\theta})$ and the convolutions with $\mcal{N}(x,0,\sigobs)$.

    \vspace{5pt}
    \subsection{Flat Distribution} \label{sub:flat_distribution}

        The simplest distribution is that of the univariate uniform
        distribution.  For a domain $x \in [a, b]$ the PDF is given by
        \begin{equation} \label{eq:pdf_flat_univariate}
            P_{\mcal{U}}(x|a,b) = \begin{cases}
                (b-a)^{-1} & a < x \leq b \\
                0 & \text{else}
            \end{cases}
        \end{equation}
        where $a,b$ are the bounds.

        In practice, the bounds $a,b$ are the bounds of the observation window,
        creating a subset of a larger field.  Suppose the larger field is
        uniformly distributed in a region larger than the bounds $a, b$, ie.
        $P_{\mcal{U}}(x|\alpha,\beta)$.  Let each observed datum $x_n$ in the
        field have Gaussian error $\sigobs_n$, then the PDF of the
        uniform-Gaussian-convolved distribution is given by:

        \begin{equation} \label{eq:flat_distribution}
            \pdf_{(\mcal{U}*\mcal{N})}(x; \alpha, \beta, \sigobs) = \begin{cases}
                \pdf_\mcal{U}(x; \alpha, \beta) \left( \cdf_{\mcal{N}}(x; \alpha, \sigobs) - \cdf_{\mcal{N}}(x; \beta, \sigobs) \right) & \alpha \leq x \leq \beta \\
                0 & \text{else}
            \end{cases}
        \end{equation}

        We then truncate this distribution to the observed field $(a, b]$
        \begin{equation} \label{eq:pdf_flat_univariate_convolved_error_full}
            \pdf_{(\mcal{U}*\mcal{N})}(x | a < X \leq b; \alpha, \beta, \sigobs)
        \end{equation}
        in the normal fashion: setting $\pdf_{(\mcal{U}*\mcal{N})} = 0$ when $x
        < a, x > b$ and normalizing $\pdf_{(\mcal{U}*\mcal{N})}$ within $(a,b]$.
        The truncated distribution is tractable but unwieldy. The distribution
        can be simplified enormously if two conditions hold: first, $\alpha \ll
        a, b \ll \beta$, the observed field is a small subset of the full field;
        and second, $\sigobs \ll |\alpha - a|$ and $\sigobs \ll |\beta - b|$,
        the errors are smaller than the observation window size compared to the
        full field.  In this case, the distribution reduces back to the uniform
        distribution!

        Therefore, for all cases considered in this work
        \eqref{eq:pdf_flat_univariate_convolved_error_full} becomes:

        \begin{equation}\label{eq:pdf_flat_univariate_convolved_error}
            \pdf_{(\mcal{U}*\mcal{N})}(x | a < X \leq b; \alpha, \beta, \sigobs) \cong P_{\mcal{U}}(x|a,b)
        \end{equation}

    \vspace{10pt}
    \subsection{Truncated Exponential Distribution} \label{app:sub:exponential_distribution}

        The univariate exponential distribution is:
        \begin{equation} \label{eq:pdf_exp_univariate}
            \pdf_{\mcal{E}}(x; \lambda) = \begin{cases}
                m \Exp{-\lambda x} & x \in [0, \infty) \\
                0 & x < 0
            \end{cases}
        \end{equation}

        The support of this distribution, $[0, \infty)$, rarely matches the
        support of the data. We generalize to the unit-normalized, truncated,
        univariate exponential distribution in the domain $(a, b]$:
        \begin{equation} \label{eq:pdf_truncexp_univariate}
            \pdf_{\mcal{E}_T}(x; \lambda, a, b) = \begin{cases}
                \frac{\lambda \Exp{-\lambda \, (x - a))}}{1 - \Exp{-\lambda(b - a)}} & a < x \leq b \\
                0 & \text{else},
            \end{cases}
        \end{equation}
        for bounds $x \in (a,b]$. However, as $\lambda \rightarrow 0$, which
        describes the flat distribution (\autoref{sub:flat_distribution}), this
        functional form can be numerically unstable. It is therefore practical
        to define $\pdf_{\mcal{E}}(x; |\lambda| < \epsilon, a, b) \cong
        P_{\mcal{U}}(x; a,b)$.

        Let each observed datum $x_n$ in the field have Gaussian error
        $\sigobs_n$, then the PDF of the Exponential-convolved-with-a-Gaussian
        distribution is given by:

        \begin{equation}
            \pdf_{\mcal{E}\cdot\mcal{N}}(x;m,a,\sigma_*) = \frac{m  e^{\frac{1}{2} \lambda  \left(2 b+\lambda  \sigma _*^2-2 x\right)}}{e^{\lambda  (b-a)}-1} = \pdf_{\mcal{E}}(x;\lambda,a) \ \Exp{\frac{1}{2}\lambda^2 \sigma_*^2},
        \end{equation}
        where the function is no longer truncated, since a Gaussian is not
        compact.

        We impose the same observation field $(a, b]$ and truncate
        $\pdf_{\mcal{E}\cdot\mcal{N}}(x | a < X \leq b;m,a,\sigma_*)$.  We note
        the normalization, which is the CDF evaluated at $a, b$, cancels the
        factor $\Exp{\frac{1}{2}\lambda^2 \sigma_*^2}$. Following thtough, the
        truncated convolved distribution is identical to the original
        distribution.

        \begin{equation}
            \pdf_{(\mcal{E}\cdot\mcal{N})_T}x | a < X \leq b;m,a,\sigma_*) \equiv \pdf_{\mcal{E}}(x; \lambda, a, b)
        \end{equation}

    \vspace{10pt}
    \subsection{Normal Distribution} \label{app:sub:normal_distribution}

        The properties of the Normal distribution are well known. For
        completeness we state

        \begin{equation}
            P_X(x; \mu, \sigma) = \mcal{N}(\mu, \sigma).
        \end{equation}

        With the inclusion of observational errors the distribution becomes

        \begin{equation}
            P_{\mcal{N}\cdot\mcal{N}}(x; \mu, \sigma, \sigobs) = \mcal{N}(\mu, \sqrt{\sigma^2 + \sigobs^2}).
        \end{equation}

        The Normal distribution is not compact, with non-zero value over all
        $x$. In an observational window $(a,b]$ the PDF is truncated, giving

        \begin{equation}
            P_{\mcal{N}_T}(x; \mu, \sigma, a, b) = \begin{cases}
                \frac{\mcal{N}(x, \mu, \sigma)}{\Phi_\mcal{N}(b; \mu, \sigma) - \Phi_\mcal{N}(a; \mu, \sigma)} & a < x \leq b \\
                0 & \text{else}
            \end{cases}
        \end{equation}

        Note that the observation-window-truncation is performed on the full
        distribution, which includes the observational errors, if present.
        Therefore, let $\tilde{\sigma}^2 \equiv \sigma^2+\sigma_*^2$

        \begin{equation}
            P_{(\mcal{N}\cdot\mcal{N})_T}(x; \mu, \tilde{\sigma}, a, b) = P_{\mcal{N}_T}(x; \mu, \tilde{\sigma}, a, b)
        \end{equation}

\end{document}